### 2018-09-17 Docs/Marami2018/Abstract/marami.tex
abstract_marami

\documentclass[12pt]{article}

\usepackage{times}

% general packages without options
\usepackage{amsmath,amssymb,bbm}
% graphics
\usepackage{graphicx}
% text formatting
\usepackage[document]{ragged2e}
\usepackage{pagecolor,color}

\newcommand{\noun}[1]{\textsc{#1}}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% geometry
\usepackage[margin=2.5cm]{geometry}

\usepackage{multicol}
\usepackage{setspace}

\usepackage{natbib}
\setlength{\bibsep}{0.0pt}

%\usepackage[french]{babel}

% layout : use fancyhdr package
%\usepackage{fancyhdr}
%\pagestyle{fancy}

% variable to include comments or not in the compilation ; set to 1 to include
\def \draft {1}

% writing utilities

% comments and responses
%  -> use this comment to ask questions on what other wrote/answer questions with optional arguments (up to 4 answers)
\usepackage{xparse}
\usepackage{ifthen}
\DeclareDocumentCommand{\comment}{m o o o o}
{\ifthenelse{\draft=1}{
    \textcolor{red}{\textbf{C : }#1}
    \IfValueT{#2}{\textcolor{blue}{\textbf{A1 : }#2}}
    \IfValueT{#3}{\textcolor{ForestGreen}{\textbf{A2 : }#3}}
    \IfValueT{#4}{\textcolor{red!50!blue}{\textbf{A3 : }#4}}
    \IfValueT{#5}{\textcolor{Aquamarine}{\textbf{A4 : }#5}}
 }{}
}

% todo
\newcommand{\todo}[1]{
\ifthenelse{\draft=1}{\textcolor{red!50!blue}{\textbf{TODO : \textit{#1}}}}{}
}

\makeatletter

\makeatother

\linespread{1.25}

\begin{document}

\title{
A multi-dimensional percolation approach to characterize sustainable mega-city regions
\medskip\\
\textit{Marami 2018\\
Communication proposal
}
}
\author{\noun{Juste Raimbault}\medskip\\
(UPS CNRS 3611 ISC-PIF et UMR CNRS 8504 Géographie-cités)\\
}
\date{}

\maketitle

\justify

\pagenumbering{gobble}

\textbf{Keywords : }\textit{Road network; multilayer percolation; mega-city region}

\medskip

The structure of road networks both translates its past growth dynamics and has a significant impact on the sustainability of territories it irrigates. A method to characterize topologies of these spatial networks is network percolation. Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling} and to the analysis of street networks for example to extract endogenous urban regions \citep{arcaute2016cities} or to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}. Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.

This communication addresses such a gap by introducing a multi-dimensional percolation heuristic, which is analog to multilayer network percolation \citep{boccaletti2014structure}. Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold. We apply the heuristic to urban morphology and road network topology measures in Europe. More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}. We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.

We systematically explore the clusters obtained for 840 parameter configurations. Maps reveal that most configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}. We use this endogenous definition of regional urban systems the percolation algorithm produced to evaluate their sustainability, in terms of conflicting objectives of economic integration and greenhouse gases emissions. Applying a gravity model to each region, we estimate transportation flows within each and extrapolate emissions by coupling with the Edgar emission database \citep{janssens2017edgar} and economic activities with a scaling law of population. We show therein that different population, network and distance thresholds yield different performances in terms of sustainability, exhibiting a Pareto front. This suggests policies in terms of regional integration to increase the sustainability of mega-city regions. Further work will consist in the use of calibration heuristics \citep{reuillon2013openmole} to find in a more robust way optimal parameter values.

%%%%%%%%%%%%%%%%%%%%
%% Biblio
%%%%%%%%%%%%%%%%%%%%
%\tiny

%\begin{multicols}{2}

%\setstretch{0.3}
%\setlength{\parskip}{-0.4em}

\bibliographystyle{apalike}
\bibliography{biblio}
%\end{multicols}

\end{document}

### 2018-09-17 Models/Percolation/percolation.R

@@ -49,6 +49,28 @@ save(res,file='directSampling.RData')
 
 # load('directSampling.RData')
 
+# get summary indicators of pop / areas distrib
+
+areasalpha = unlist(lapply(res,function(l){
+  areas = l$areas
+  return(ifelse(length(which(areas>0)),lm(data=data.frame(rank=log(1:length(areas[areas>0])),size=log(areas[areas>0])),size~rank)$coefficients[2],NA))
+}))
+
+popsalpha = unlist(lapply(res,function(l){
+  pops = l$pops
+  return(ifelse(length(which(pops>0)),lm(data=data.frame(rank=log(1:length(pops[pops>0])),size=log(pops[pops>0])),size~rank)$coefficients[2],NA))
+}))
+
+clustnum = unlist(lapply(res,function(l){length(l$pops)}))
+
+d=data.frame(areasalpha,popsalpha,clustnum)
+
+g=ggplot(d[clustnum>15,],aes(x=areasalpha,y=popsalpha,colour=clustnum))
+g+geom_point()
+
+# -> hierarchy of "endogenous city system" ?
+# : find criteria on which to optimize !
+
 
 
 

### 2018-09-17 Models/Percolation/percolationFunctions.R

@@ -74,9 +74,17 @@ conditionalPercolation <- function(d,popthq,nwthq,radius,minclustsize=5,
     ggsave(g+geom_raster()+scale_fill_discrete(name="cluster"),file=paste0(resdir,popcol,popth,'_',nwcol,nwth,'_radius',radius,'.png'),width=15,height=10,units='cm')
   }
   
-  return(computeClustersIndics(sppoints))
-  #return(sppoints)
+  #return(computeClustersIndics(sppoints))
+  return(sppoints)
 }
+
+
+#'
+#'
+computeIndics <- function(sppoints,popcol="totalPop"){
+  
+}
+
 
 
 

### 2018-10-06 Models/DataPreparation/data.R
rows

@@ -1,0 +1,33 @@
+
+setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology'))
+
+library(dplyr)
+
+source('Models/StaticCorrelations/mapFunctions.R')
+
+# network and pop indicators
+areasize=100;offset=50;factor=0.5
+indics = as.tbl(loadIndicatorData(paste0(Sys.getenv('CS_HOME'),"/UrbanMorphology/Data/StaticCorrelations/res/europecoupled_areasize",areasize,"_offset",offset,"_factor",factor,"_temp.RData")))
+
+# emissions
+# -> aggregate on different gases: for rows emissions -> indics, no need for now (same grid each gas)
+emissions <- as.tbl(read.csv('Data/EDGAR/v432_CO2_excl_short-cycle_org_C_2012/v432_CO2_excl_short-cycle_org_C_2012.txt',sep=';'))
+
+
+# by hand "join" (aggreg)
+
+library(doParallel)
+cl <- makeCluster(60,outfile='log')
+registerDoParallel(cl)
+rows <- foreach(i=1:nrow(emissions)) %dopar% {
+  #dif=abs(emissions$lon-indics$lonmin[i])+abs(emissions$lat-indics$latmin[i])
+  #dif=abs(emissions$lon[i]-indics$lonmin)+abs(emissions$lat[i]-indics$latmin)
+  dif=sqrt((emissions$lon[i]-indics$lonmin)^2+(emissions$lat[i]-indics$latmin)^2)
+  return(which(dif==min(dif))[1])
+}
+stopCluster(cl)
+save(rows,file='Models/DataPreparation/rowsemissions.RData')
+system('rm log')
+stopCluster(cl)
+
+

### 2018-10-06 Models/Percolation/percolationFunctions.R

@@ -117,15 +117,17 @@ morphoIndices <-function(sppoints,popcol="totalPop"){
 #' Compute a gravity potential between spatial points with population, of the form
 #'   (P_i P_j / P_tot)^gamma * exp(-d_{ij} / d_0)
 #'   TODO : influence of normalizing by pmax vs ptot ? -> investigate in interaction models
-#'   
-#'   - normalized such that sum = 1
-interactionPotential <- function(sppoints,gravityGamma=1,gravityDistance=1,popcol="totalPop"){
+#'   - normalized such that sum = 1 ? NO to be able to have a size effects in comparing configurations
+interactionPotential <- function(sppoints,gravityGamma=1,gravityDistance=1,weightcol="totalPop"){
   dmat = exp(-st_distance(sppoints)/gravityDistance)
   diag(dmat)<-0
   dmat = Matrix(dmat)
-  pops = sppoints[[popcol]]
-  popmat = Diagonal(x=(pops/sum(pops))^gravityGamma)
-  potentials = popmat%*%dmat%*%popmat
+  #pops = sppoints[[weightcol]]
+  #popmat = Diagonal(x=(pops/sum(pops))^gravityGamma)
+  weights=sppoints[[weightcol]]
+  # weights assumed already normalized
+  weightmat=Diagonal(x=(weights)^gravityGamma)
+  potentials = weightmat%*%dmat%*%weightmat
   return(as.matrix(potentials))
 }
 

### 2018-10-06 Models/DataPreparation/data.R
.

@@ -17,13 +17,16 @@ emissions <- as.tbl(read.csv('Data/EDGAR/v432_CO2_excl_short-cycle_org_C_2012/v4
 # by hand "join" (aggreg)
 
 library(doParallel)
-cl <- makeCluster(60,outfile='log')
+cl <- makeCluster(10,outfile='log')
 registerDoParallel(cl)
 rows <- foreach(i=1:nrow(emissions)) %dopar% {
   #dif=abs(emissions$lon-indics$lonmin[i])+abs(emissions$lat-indics$latmin[i])
   #dif=abs(emissions$lon[i]-indics$lonmin)+abs(emissions$lat[i]-indics$latmin)
+  show(paste0(100*i/nrow(emissions),'%'))
   dif=sqrt((emissions$lon[i]-indics$lonmin)^2+(emissions$lat[i]-indics$latmin)^2)
-  return(which(dif==min(dif))[1])
+  indmin = which(dif==min(dif))[1]
+  show(indmin)
+  return(indmin)
 }
 stopCluster(cl)
 save(rows,file='Models/DataPreparation/rowsemissions.RData')

### 2018-10-06 Models/DataPreparation/consolidate.R
dataprep

@@ -1,0 +1,105 @@
+
+##
+# consolidate datasets
+
+
+setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology'))
+
+library(dplyr)
+library(ggplot2)
+source(paste0(Sys.getenv('CS_HOME'),'/Organisation/Models/Utils/R/plots.R'))
+
+source('Models/StaticCorrelations/mapFunctions.R')
+
+# network and pop indicators
+areasize=100;offset=50;factor=0.5
+indics = as.tbl(loadIndicatorData(paste0(Sys.getenv('CS_HOME'),"/UrbanMorphology/Data/StaticCorrelations/res/europecoupled_areasize",areasize,"_offset",offset,"_factor",factor,"_temp.RData")))
+
+
+# different emissions gases
+emissionsCO2 <- as.tbl(read.csv('Data/EDGAR/v432_CO2_excl_short-cycle_org_C_2012/v432_CO2_excl_short-cycle_org_C_2012.txt',sep=';'))
+emissionsCH4 <- as.tbl(read.csv('Data/EDGAR/v432_CH4_2012/v432_CH4_2012.txt',sep=';'))
+emissionsN20 <- as.tbl(read.csv('Data/EDGAR/v432_N2O_2012/v432_N2O_2012.txt',sep=';'))
+emissionsCO2short <- as.tbl(read.csv('Data/EDGAR/v432_CO2_org_short-cycle_C_2012/v432_CO2_org_short-cycle_C_2012.txt',sep=';'))
+
+# summaries
+summary(emissionsCO2$emission)
+summary(emissionsCH4$emission)
+summary(emissionsN20$emission)
+summary(emissionsCO2short$emission)
+# Lashof, D. A., & Ahuja, D. R. (1990). Relative contributions of greenhouse gas emissions to global warming. Nature, 344(6266), 529.
+# => keep only CO2
+# (for more thorough study should be more precise on ghg potential ?)
+
+aggregems=data.frame(emissions=c(sum(emissionsCO2$emission),sum(emissionsCH4$emission),sum(emissionsN20$emission),sum(emissionsCO2short$emission)),
+                    Type=c("CO2","CH4","N2O","CO2 (short)")
+                     )
+g <- ggplot(aggregems, aes(x="",y=emissions, fill=Type))
[…]

### 2018-10-06 Models/MorphologyPackage/morphology.R

@@ -4,11 +4,14 @@
 # TODO creating R packages
 # https://cran.r-project.org/doc/contrib/Leisch-CreatingPackages.pdf
 
+#setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/MorphologyPackage'))
+
+
+#######
+
 library(raster)
 library(sf)
-
-
-
+library(Matrix)
 
 #########
 ## Vector functions
@@ -18,7 +21,7 @@ library(sf)
 moranPoints <-function(points,varcol,weights=spatialWeightsPoints(points)){
   v = points[[varcol]]
   v = v-mean(v)
-  num = sum(weights*(Diagonal(v)%*%Diagonal(v)))
+  num = sum(Diagonal(x = v)%*%weights%*%Diagonal(x = v))
   denom = sum(v^2)
   normalization = nrow(points) / sum(weights)
   return(normalization*num/denom)
@@ -28,13 +31,22 @@ moranPoints <-function(points,varcol,weights=spatialWeightsPoints(points)){
 #'
 spatialWeightsPoints<-function(points){
   distances <- 1/st_distance(points)
-  distances[distances==Inf]=0
-  return(distances)
+  #distances[as.matrix(distances)==matrix(rep(Inf,nrow(distances)*ncol(distances)),nrow=nrow(distances))]=0
+  # no point at the same place
+  diag(distances)<-0
+  return(Matrix(drop_units(distances)))
 }
 
-avgDistancePoints<-function(){
-  
+#'
+#'
+avgDistancePoints<-function(points,varcol){
+  v = points[[varcol]]
+  distances <- 1/st_distance(points)
+  distances[distances==Inf]=0
+  vv = Diagonal(v)%*%Diagonal(v)
+  return(distances*vv/(sum(vv)^2)/max(distances))
 }
+
 
 
 
@@ -178,6 +190,8 @@ getCoordinates<-function(file,r,c){
 
 
 
+#system('rm -rf morphology')
+#package.skeleton(name='morphology')
 
-
+#install.packages(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/MorphologyPackage/morphology/'),repos=NULL,type='source')
 

### 2018-10-06 Models/Percolation/percolation.R

@@ -7,11 +7,10 @@ library(dplyr)
 source('../StaticCorrelations/mapFunctions.R')
 source('percolationFunctions.R')
 
-areasize=100;offset=50;factor=0.5
-indics = as.tbl(loadIndicatorData(paste0(Sys.getenv('CS_HOME'),"/UrbanMorphology/Data/StaticCorrelations/res/europecoupled_areasize",areasize,"_offset",offset,"_factor",factor,"_temp.RData")))
+# assumes data has been consolidated before
+load('Data/consolidated/indics.RData')
 
-# TODO add emission loading (transportation ?) and spatial join
-
+# parameter values
 quantiles = c(0.75,0.8,0.85,0.9,0.95)
 radiuses=c(8000,10000,15000,20000,50000,75000,100000)
 nwindics= c("ecount","mu","vcount","euclPerf")
@@ -35,11 +34,12 @@ res <- foreach(i=1:nrow(params)) %dopar% {
   source('percolationFunctions.R')
   #p = conditionalPercolation(d=indics,popthq=params$popthq[i],nwthq=params$nwthq[i],radius=params$radius[i])
   return(conditionalPercolation(d=indics,
-                                popthq=params$popthq[i],
-                                nwthq=params$nwthq[i],
                                 radius=params$radius[i],
-                                nwcol=params$nwcol[i]
-         ))
+                                popthq=params$popthq[i],
+                                nwcol=params$nwcol[i],
+                                nwthq=params$nwthq[i]
+                                )
+         )
 }
 stopCluster(cl)
 
@@ -51,26 +51,26 @@ save(res,file='directSampling.RData')
 
 # get summary indicators of pop / areas distrib
 
-areasalpha = unlist(lapply(res,function(l){
-  areas = l$areas
-  return(ifelse(length(which(areas>0)),lm(data=data.frame(rank=log(1:length(areas[areas>0])),size=log(areas[areas>0])),size~rank)$coefficients[2],NA))
-}))
-
-popsalpha = unlist(lapply(res,function(l){
-  pops = l$pops
-  return(ifelse(length(which(pops>0)),lm(data=data.frame(rank=log(1:length(pops[pops>0])),size=log(pops[pops>0])),size~rank)$coefficients[2],NA))
-}))
-
[…]

### 2018-10-06 Models/Percolation/percolationFunctions.R

@@ -13,15 +13,14 @@ library(Matrix)
 
 #'
 #'
-conditionalPercolation <- function(d,popthq,nwthq,radius,minclustsize=5,
-                                   xcol="lonmin",ycol="latmin",popcol="totalPop",nwcol="euclPerf",
+conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,minclustsize=5,
+                                   xcol="lonmin",ycol="latmin",popcol="totalPop",
+                                   withMaps=F,
                                    resdir=paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/Maps/')
                                    ){
   dir.create(resdir)
-  #popth=5000000;nwth=3000;radius=50000
+  #popthq=0.95;nwthq=0.95;radius=50000;nwcol="euclPerf";popcol="totalPop";minclustsize=5
   #xcol="lonmin";ycol="latmin";popcol="totalPop";nwcol="mu";d=indics
-  #x=data[,xcol];y=data[,ycol];population=data[,popcol];network=data[,nwcol]
-  #sppoints = SpatialPoints(data.frame(lon=x,lat==y),proj4string = countries@proj4string)
   
   show(paste0("radius = ",radius))
   
@@ -68,31 +67,26 @@ conditionalPercolation <- function(d,popthq,nwthq,radius,minclustsize=5,
   #show(sppoints)
   #show(clusters)
   
-  #plot(sppoints%>%transmute(cluster=cluster))
-  if(length(which(!is.na(clusters)))>0){
-    g=ggplot(data.frame(x=indics$lonmin,y=indics$latmin,cluster=clusters),aes(x=x,y=y,fill=as.character(cluster)))
-    ggsave(g+geom_raster()+scale_fill_discrete(name="cluster"),file=paste0(resdir,popcol,popth,'_',nwcol,nwth,'_radius',radius,'.png'),width=15,height=10,units='cm')
+  if(withMaps==T){
+    #plot(sppoints%>%transmute(cluster=cluster))
+    if(length(which(!is.na(clusters)))>0){
+      g=ggplot(data.frame(x=indics$lonmin,y=indics$latmin,cluster=clusters),aes(x=x,y=y,fill=as.character(cluster)))
[…]

### 2018-10-06 Models/DataPreparation/data.R
explos

@@ -2,6 +2,7 @@
 setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology'))
 
 library(dplyr)
+library(sp)
 
 source('Models/StaticCorrelations/mapFunctions.R')
 
@@ -13,24 +14,28 @@ indics = as.tbl(loadIndicatorData(paste0(Sys.getenv('CS_HOME'),"/UrbanMorphology
 # -> aggregate on different gases: for rows emissions -> indics, no need for now (same grid each gas)
 emissions <- as.tbl(read.csv('Data/EDGAR/v432_CO2_excl_short-cycle_org_C_2012/v432_CO2_excl_short-cycle_org_C_2012.txt',sep=';'))
 
+# filter on bbox
+emissions=emissions[emissions$lon>=min(indics$lonmin)&emissions$lon<=max(indics$lonmin)&emissions$lat>=min(indics$latmin)&emissions$lat<=max(indics$latmin),]
 
 # by hand "join" (aggreg)
 
-library(doParallel)
-cl <- makeCluster(10,outfile='log')
-registerDoParallel(cl)
-rows <- foreach(i=1:nrow(emissions)) %dopar% {
+#library(doParallel)
+#cl <- makeCluster(4,outfile='log')
+#registerDoParallel(cl)
+#rows <- foreach(i=1:nrow(emissions)) %dopar% {
+rows=c()
+for(i in 1:nrow(emissions)){
   #dif=abs(emissions$lon-indics$lonmin[i])+abs(emissions$lat-indics$latmin[i])
   #dif=abs(emissions$lon[i]-indics$lonmin)+abs(emissions$lat[i]-indics$latmin)
-  show(paste0(100*i/nrow(emissions),'%'))
-  dif=sqrt((emissions$lon[i]-indics$lonmin)^2+(emissions$lat[i]-indics$latmin)^2)
-  indmin = which(dif==min(dif))[1]
-  show(indmin)
-  return(indmin)
+  if(i%%1000==0){show(paste0(100*i/nrow(emissions),'%'))}
+  #dif=sqrt((emissions$lon[i]-indics$lonmin)^2+(emissions$lat[i]-indics$latmin)^2)
+  dif = spDistsN1(as.matrix(indics[,c("lonmin","latmin")]),c(emissions$lon[i],emissions$lat[i]),longlat = T)
+  indmin = ifelse(min(dif) < 15 ,which(dif==min(dif))[1],NA)
[…]

### 2018-10-06 Models/MorphologyPackage/morphology.R

@@ -43,11 +43,24 @@ avgDistancePoints<-function(points,varcol){
   v = points[[varcol]]
   distances <- 1/st_distance(points)
   distances[distances==Inf]=0
-  vv = Diagonal(v)%*%Diagonal(v)
-  return(distances*vv/(sum(vv)^2)/max(distances))
+  vv = Diagonal(v/sum(v^2))%*%distances%*%Diagonal(v/sum(v^2))
+  return(sum(vv)/max(distances))
 }
 
+entropyPoints<-function(points,varcol){
+  v = points[[varcol]]
+  v=v[v>0]
+  return(-sum(v*log(v))/log(length(v)))
+}
 
+slopePoints<-function(points,varcol){
+  v = points[[varcol]]
+  v=v[v>0]
+  v=sort(v,decreasing=TRUE)
+  rank = log(1:length(v))
+  reg = lm(v~rank,data.frame(rank,v))
+  return(reg$coefficients[2])
+}
 
 
 #########

### 2018-10-06 Models/Percolation/percolation.R

@@ -2,9 +2,7 @@
 setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/Percolation'))
 
 library(dplyr)
-#library(viridis)
 
-source('../StaticCorrelations/mapFunctions.R')
 source('percolationFunctions.R')
 
 # assumes data has been consolidated before

### 2018-10-06 Models/Percolation/percolationFunctions.R

@@ -3,8 +3,8 @@ library(dplyr)
 #library(sp)
 #library(rgeos)
 library(sf)
-library(tidyverse)
-library(ggplot2)
+#library(tidyverse)
+#library(ggplot2)
 library(Matrix)
 
 #library(devtools)
@@ -13,7 +13,7 @@ library(Matrix)
 
 #'
 #'
-conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,minclustsize=5,
+conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,minclustsize=5,
                                    xcol="lonmin",ycol="latmin",popcol="totalPop",
                                    withMaps=F,
                                    resdir=paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/Maps/')
@@ -67,26 +67,40 @@ conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,minclustsize=5,
   #show(sppoints)
   #show(clusters)
   
-  if(withMaps==T){
-    #plot(sppoints%>%transmute(cluster=cluster))
-    if(length(which(!is.na(clusters)))>0){
-      g=ggplot(data.frame(x=indics$lonmin,y=indics$latmin,cluster=clusters),aes(x=x,y=y,fill=as.character(cluster)))
-      ggsave(g+geom_raster()+scale_fill_discrete(name="cluster"),file=paste0(resdir,popcol,popth,'_',nwcol,nwth,'_radius',radius,'.png'),width=15,height=10,units='cm')
-    }
-  }
+  #if(withMaps==T){
+  #  #plot(sppoints%>%transmute(cluster=cluster))
+  #  if(length(which(!is.na(clusters)))>0){
+  #    g=ggplot(data.frame(x=indics$lonmin,y=indics$latmin,cluster=clusters),aes(x=x,y=y,fill=as.character(cluster)))
+  #    ggsave(g+geom_raster()+scale_fill_discrete(name="cluster"),file=paste0(resdir,popcol,popth,'_',nwcol,nwth,'_radius',radius,'.png'),width=15,height=10,units='cm')
+  #  }
+  #}
     
-  return(computeIndics(sppoints))
+  return(computeIndics(sppoints,gamma=gamma,decay=decay))
   #return(sppoints)
 }
 
 
+aggregIndics <- function(res){
+  return(list(
+    area=sum(res$areas),
+    pop=sum(res$areas),
+    moran=mean(res$morans),
[…]

### 2018-10-07 Docs/Marami2018/Communication/marami2018.tex
skeleton_slides

\documentclass[english,11pt]{beamer}

\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Proba}{\mathbb{P}}

\newcommand{\Covb}[2]{\ensuremath{\Cov\!\left[#1,#2\right]}}
\newcommand{\Eb}[1]{\ensuremath{\E\!\left[#1\right]}}
\newcommand{\Pb}[1]{\ensuremath{\Proba\!\left[#1\right]}}
\newcommand{\Varb}[1]{\ensuremath{\Var\!\left[#1\right]}}

% norm
\newcommand{\norm}[1]{\| #1 \|}

\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}

\usepackage{mathptmx,amsmath,amssymb,graphicx,bibentry,bbm,babel,ragged2e}

\makeatletter

\newcommand{\noun}[1]{\textsc{#1}}
\newcommand{\jitem}[1]{\item \begin{justify} #1 \end{justify} \vfill{}}
\newcommand{\sframe}[2]{\frame{\frametitle{#1} #2}}

\newenvironment{centercolumns}{\begin{columns}[c]}{\end{columns}}
%\newenvironment{jitem}{\begin{justify}\begin{itemize}}{\end{itemize}\end{justify}}

\usetheme{Warsaw}
\setbeamertemplate{footline}[text line]{}
\setbeamercolor{structure}{fg=purple!50!blue, bg=purple!50!blue}

\setbeamersize{text margin left=15pt,text margin right=15pt}

\setbeamercovered{transparent}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{multirow}

\makeatother

\begin{document}

\title{A multi-dimensional percolation approach to characterize sustainable mega-city regions}

\author{J.~Raimbault$^{1,2,\ast}$\\
\texttt{juste.raimbault@polytechnique.edu}
}

\institute{$^{1}$Complex Systems Institute, Paris, UPS CNRS 3611 ISC-PIF\\
$^{2}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
}

\date{MARAMI 2018\\\smallskip
Avignon\\\smallskip
October 18th 2018
}

\frame{\maketitle}

\sframe{Networks and territories}{

% Keywords : }\textit{Road network; multilayer percolation; mega-city region}

% - intro : LGV ? Avignon : travaux Claire ! - truc avec les Papes ? 
% https://www.persee.fr/doc/anami_0003-4398_1985_num_97_172_2228_t1_0446_0000_3
% https://www.persee.fr/doc/mefr_0223-5110_1984_num_96_1_2762

\centering

	%\includegraphics[width=0.52\textwidth]{figures/}

	\footnotesize
%\textit{Source: Wikipedia}

}

\sframe{Characterizing Road networks}{

%The structure of road networks both translates its past growth dynamics and has a significant impact on the sustainability of territories it irrigates.

}

\sframe{Network percolation}{

% A method to characterize topologies of these spatial networks is network percolation. Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling} and to the analysis of street networks for example to extract endogenous urban regions \citep{arcaute2016cities} or to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.

% from monodimensional to multidimensional

}

\sframe{Multidimensional percolation}{

% Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.

% - why the relation between form and function, aka pop and network in our view is crucial
% - take into account in percolation
% - potential application : endogenous sustainable urban entities

\justify

$\rightarrow$

\medskip

$\rightarrow$

% research question

\bigskip
\bigskip

\textbf{Research objective : } \textit{}

}

% \section{Method}

\sframe{Formalization}{

%This communication addresses such a gap by introducing a multi-dimensional percolation heuristic, which is analog to multilayer network percolation \citep{boccaletti2014structure}. Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold.

}

\sframe{Empirical data and variables}{

%We apply the heuristic to urban morphology and road network topology measures in Europe. More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}. We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.

}

\sframe{Experience plan}{

%We systematically explore the clusters obtained for 840 parameter configurations.

% + explain openmole

}

\sframe{Results: endogenous mega-regions}{

%Maps reveal that most configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}.

}

\sframe{Characterizing sustainibility}{

% We use this endogenous definition of regional urban systems the percolation algorithm produced to evaluate their sustainability, in terms of conflicting objectives of economic integration and greenhouse gases emissions. Applying a gravity model to each region, we estimate transportation flows within each and extrapolate emissions by coupling with the Edgar emission database \citep{janssens2017edgar} and economic activities with a scaling law of population.

}

\sframe{Results: Pareto fronts}{

%  We show therein that different population, network and distance thresholds yield different performances in terms of sustainability, exhibiting a Pareto front. This suggests policies in terms of regional integration to increase the sustainability of mega-city regions.

}

\sframe{Calibration}{

% Further work will consist in the use of calibration heuristics \citep{reuillon2013openmole} to find in a more robust way optimal parameter values.

% -> depending on results

}

\sframe{Discussion}{

% This work paves the way towards ?

\justify

\vspace{-1cm}

\textbf{Implications}

$\rightarrow$ 
\smallskip

$\rightarrow$

\bigskip

\textbf{Developments}

%

$\rightarrow$

}

\sframe{Conclusion}{

\justify

\vspace{-1cm}

% rq : try in all conclusion to come back to these points, close to Banos commadments ?

$\rightarrow$ %First step towards systematic benchmarks and multi-modeling. \textbf{Need for more systematic model exploration.} % model coupling / benchmarking

\smallskip

$\rightarrow$ %Model integration and multi-scalarity ? \textbf{Need for more integrated models.}

\smallskip

$\rightarrow$ %Multiple perspectives on urban systems ? \textbf{Need for more interdisciplinarity.}

\bigskip

\footnotesize

\textbf{Related works}

Raimbault, J. (2018). Calibration of a density-based model of urban morphogenesis. PloS one, 13(9), e0203516.

\smallskip

Raimbault, J. (2018). An Urban Morphogenesis Model Capturing Interactions between Networks and Territories. \textit{Forthcoming in Mathematics or Urban Morphogenesis.} arXiv:1805.05195.

\smallskip

Raimbault, J. (2018). Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution des r{\'e}seaux de transport et des territoires (Doctoral dissertation, Université Paris 7 Denis Diderot). \url{https://halshs.archives-ouvertes.fr/tel-01857741}

\smallskip

\textbf{Open repository} at \texttt{https://github.com/JusteRaimbault/UrbanMorphology}\\\smallskip
\textbf{Acknowledgments}: thanks to the \textit{EGI} for access to the infrastructure.

}

\sframe{Reserve slides}{

\centering

\Large

\textbf{Reserve Slides}

}

%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{/Users/juste/ComplexSystems/CityNetwork/Biblio/Bibtex/CityNetwork,biblio}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2018-10-07 Models/DataPreparation/data.R

@@ -23,17 +23,20 @@ emissions=emissions[emissions$lon>=min(indics$lonmin)&emissions$lon<=max(indics$
 #cl <- makeCluster(4,outfile='log')
 #registerDoParallel(cl)
 #rows <- foreach(i=1:nrow(emissions)) %dopar% {
-rows=c()
+rows=c();mins=c()
 for(i in 1:nrow(emissions)){
   #dif=abs(emissions$lon-indics$lonmin[i])+abs(emissions$lat-indics$latmin[i])
   #dif=abs(emissions$lon[i]-indics$lonmin)+abs(emissions$lat[i]-indics$latmin)
   if(i%%1000==0){show(paste0(100*i/nrow(emissions),'%'))}
   #dif=sqrt((emissions$lon[i]-indics$lonmin)^2+(emissions$lat[i]-indics$latmin)^2)
   dif = spDistsN1(as.matrix(indics[,c("lonmin","latmin")]),c(emissions$lon[i],emissions$lat[i]),longlat = T)
-  indmin = ifelse(min(dif) < 15 ,which(dif==min(dif))[1],NA)
+  #indmin = ifelse(min(dif) < 15 ,which(dif==min(dif))[1],NA) # check distrib of distances
+  indmin = which(dif==min(dif))[1]
+  mins=append(mins,min(dif))
   rows=append(rows,indmin)
 }
 #stopCluster(cl)
+
 save(rows,file='Models/DataPreparation/rowsemissions.RData')
 #system('rm log')
 #stopCluster(cl)

### 2018-10-08 Models/DataPreparation/consolidate.R
consolidated_data

@@ -15,12 +15,18 @@ source('Models/StaticCorrelations/mapFunctions.R')
 areasize=100;offset=50;factor=0.5
 indics = as.tbl(loadIndicatorData(paste0(Sys.getenv('CS_HOME'),"/UrbanMorphology/Data/StaticCorrelations/res/europecoupled_areasize",areasize,"_offset",offset,"_factor",factor,"_temp.RData")))
 
+# filter indics bbox
+filterbbox <- function(d){d[d$lon>=min(indics$lonmin)&d$lon<=max(indics$lonmin)&d$lat>=min(indics$latmin)&d$lat<=max(indics$latmin),]}
 
 # different emissions gases
 emissionsCO2 <- as.tbl(read.csv('Data/EDGAR/v432_CO2_excl_short-cycle_org_C_2012/v432_CO2_excl_short-cycle_org_C_2012.txt',sep=';'))
+emissionsCO2=filterbbox(emissionsCO2)
 emissionsCH4 <- as.tbl(read.csv('Data/EDGAR/v432_CH4_2012/v432_CH4_2012.txt',sep=';'))
+emissionsCH4=filterbbox(emissionsCH4)
 emissionsN20 <- as.tbl(read.csv('Data/EDGAR/v432_N2O_2012/v432_N2O_2012.txt',sep=';'))
+emissionsN20=filterbbox(emissionsN20)
 emissionsCO2short <- as.tbl(read.csv('Data/EDGAR/v432_CO2_org_short-cycle_C_2012/v432_CO2_org_short-cycle_C_2012.txt',sep=';'))
+emissionsCO2short=filterbbox(emissionsCO2short)
 
 # summaries
 summary(emissionsCO2$emission)
@@ -43,24 +49,34 @@ cO2joined = left_join(emissionsCO2,emissionsCO2short,by=c('lat','lon'))
 cO2joined[is.na(cO2joined)]=0
 cO2 = data.frame(lat=cO2joined$lat,lon=cO2joined$lon,emission=cO2joined$emission.x+cO2joined$emission.y)
 # aggregate by indic cells (requires data.R to have been run)
-load('Models/DataPreparation/rowsemissions.RData')
-cO2$indicrow = rows
-cO2indics = cO2 %>% group_by(indicrow) %>% summarize(emission=sum(emission))
[…]

### 2018-10-08 Models/DataPreparation/data.R

@@ -24,12 +24,14 @@ emissions=emissions[emissions$lon>=min(indics$lonmin)&emissions$lon<=max(indics$
 #registerDoParallel(cl)
 #rows <- foreach(i=1:nrow(emissions)) %dopar% {
 rows=c();mins=c()
-for(i in 1:nrow(emissions)){
+#for(i in 1:nrow(emissions)){
+for(i in 1:nrow(indics)){
   #dif=abs(emissions$lon-indics$lonmin[i])+abs(emissions$lat-indics$latmin[i])
   #dif=abs(emissions$lon[i]-indics$lonmin)+abs(emissions$lat[i]-indics$latmin)
-  if(i%%1000==0){show(paste0(100*i/nrow(emissions),'%'))}
+  if(i%%1000==0){show(paste0(100*i/nrow(indics),'%'))}
   #dif=sqrt((emissions$lon[i]-indics$lonmin)^2+(emissions$lat[i]-indics$latmin)^2)
-  dif = spDistsN1(as.matrix(indics[,c("lonmin","latmin")]),c(emissions$lon[i],emissions$lat[i]),longlat = T)
+  #dif = spDistsN1(as.matrix(indics[,c("lonmin","latmin")]),c(emissions$lon[i],emissions$lat[i]),longlat = T)
+  dif = spDistsN1(as.matrix(emissions[,c("lon","lat")]),c(indics$lonmin[i],indics$latmin[i]),longlat = T)
   #indmin = ifelse(min(dif) < 15 ,which(dif==min(dif))[1],NA) # check distrib of distances
   indmin = which(dif==min(dif))[1]
   mins=append(mins,min(dif))
@@ -37,7 +39,19 @@ for(i in 1:nrow(emissions)){
 }
 #stopCluster(cl)
 
-save(rows,file='Models/DataPreparation/rowsemissions.RData')
+#hist(log(mins),breaks=500)
+#hist(mins[mins<20],breaks=500)
+# need to check visually
+#library(ggplot2)
+#library(viridis)
+#g=ggplot(data.frame(emissions,mins)[mins<20,],aes(x=lon,y=lat,fill=mins))
+#g+geom_raster()+scale_fill_viridis(option = "magma", direction = -1)
+#ggsave(file=paste0('Results/Maps/Test/spatialaliasing.png'),width=30,height=20,units='cm')
+#length(which(mins<20)) # -> actually the resolution of emissions is smaller than our indics : redo with rows for indics
+
[…]

### 2018-10-08 Models/MorphologyPackage/morphology.R

@@ -6,6 +6,7 @@
 
 #setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/MorphologyPackage'))
 
+library(units)
 
 #######
 
@@ -41,9 +42,9 @@ spatialWeightsPoints<-function(points){
 #'
 avgDistancePoints<-function(points,varcol){
   v = points[[varcol]]
-  distances <- 1/st_distance(points)
+  distances <- drop_units(1/st_distance(points))
   distances[distances==Inf]=0
-  vv = Diagonal(v/sum(v^2))%*%distances%*%Diagonal(v/sum(v^2))
+  vv = Diagonal(x=v/sum(v^2))%*%distances%*%Diagonal(x=v/sum(v^2))
   return(sum(vv)/max(distances))
 }
 

### 2018-10-08 Models/Percolation/graphFunctions.R

@@ -1,0 +1,2 @@
+
+# graph functions

### 2018-10-08 Models/Percolation/morphology.R

@@ -1,0 +1,211 @@
+
+# morphological indices
+
+# TODO creating R packages
+# https://cran.r-project.org/doc/contrib/Leisch-CreatingPackages.pdf
+
+#setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/MorphologyPackage'))
+
+library(units)
+
+#######
+
+library(raster)
+library(sf)
+library(Matrix)
+
+#########
+## Vector functions
+
+#'
+#'
+moranPoints <-function(points,varcol,weights=spatialWeightsPoints(points)){
+  v = points[[varcol]]
+  v = v-mean(v)
+  num = sum(Diagonal(x = v)%*%weights%*%Diagonal(x = v))
+  denom = sum(v^2)
+  normalization = nrow(points) / sum(weights)
+  return(normalization*num/denom)
+}
+
+#'
+#'
+spatialWeightsPoints<-function(points){
+  distances <- 1/st_distance(points)
+  #distances[as.matrix(distances)==matrix(rep(Inf,nrow(distances)*ncol(distances)),nrow=nrow(distances))]=0
+  # no point at the same place
+  diag(distances)<-0
+  return(Matrix(drop_units(distances)))
+}
+
+#'
+#'
+avgDistancePoints<-function(points,varcol){
+  v = points[[varcol]]
+  distances <- drop_units(1/st_distance(points))
+  distances[distances==Inf]=0
+  vv = Diagonal(x=v/sum(v^2))%*%distances%*%Diagonal(x=v/sum(v^2))
+  return(sum(vv)/max(distances))
+}
+
+entropyPoints<-function(points,varcol){
+  v = points[[varcol]]
+  v=v[v>0]
+  return(-sum(v*log(v))/log(length(v)))
+}
+
+slopePoints<-function(points,varcol){
+  v = points[[varcol]]
+  v=v[v>0]
+  v=sort(v,decreasing=TRUE)
+  rank = log(1:length(v))
+  reg = lm(v~rank,data.frame(rank,v))
+  return(reg$coefficients[2])
+}
+
+
+#########
+## Raster functions
+
+
+# weights for Moran
+spatialWeights <- function (N,P,rescaleFactor=1){
+  d_hor = (matrix(rep(cumsum(matrix(1,2*N+1,1)),2*P+1),nrow=2*N+1) - N - 1) ^ 2
[…]

### 2018-10-08 Models/Percolation/percolation.R

@@ -3,6 +3,7 @@ setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/Percolation'))
 
 library(dplyr)
 
+#source('Models/Percolation/percolationFunctions.R')
 source('percolationFunctions.R')
 
 # assumes data has been consolidated before
@@ -29,19 +30,23 @@ library(doParallel)
 cl <- makeCluster(50,outfile='log')
 registerDoParallel(cl)
 res <- foreach(i=1:nrow(params)) %dopar% {
+  #source('Models/Percolation/percolationFunctions.R')
   source('percolationFunctions.R')
   #p = conditionalPercolation(d=indics,popthq=params$popthq[i],nwthq=params$nwthq[i],radius=params$radius[i])
+  # Mandatory args : d,radius,popthq,nwcol,nwthq,gamma,decay
   return(conditionalPercolation(d=indics,
                                 radius=params$radius[i],
                                 popthq=params$popthq[i],
-                                nwcol=params$nwcol[i],
-                                nwthq=params$nwthq[i]
+                                nwcol=as.character(params$nwcol[i]),
+                                nwthq=params$nwthq[i],
+                                gamma=1,
+                                decay=500
                                 )
          )
 }
 stopCluster(cl)
 
-save(res,file='directSampling.RData')
+save(res,file=paste0('res/directSampling_',format(Sys.time(), "%Y%m%d_%H%M%S"),'.RData'))
 
 #############
 

### 2018-10-08 Models/Percolation/percolationFunctions.R

@@ -7,18 +7,24 @@ library(sf)
 #library(ggplot2)
 library(Matrix)
 
+#source(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/MorphologyPackage/morphology.R'))
+# for oml : must be in the same dir
+source('morphology.R')
+
 #library(devtools)
 #devtools::install_github("tidyverse/ggplot2")
 #require(ggplot2)
 
 #'
 #'
-conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,minclustsize=5,
+conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
+                                   minclustsize=5,
                                    xcol="lonmin",ycol="latmin",popcol="totalPop",
                                    withMaps=F,
                                    resdir=paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/Maps/')
                                    ){
-  dir.create(resdir)
+  
+  if(withMaps==T){dir.create(resdir)}
   #popthq=0.95;nwthq=0.95;radius=50000;nwcol="euclPerf";popcol="totalPop";minclustsize=5
   #xcol="lonmin";ycol="latmin";popcol="totalPop";nwcol="mu";d=indics
   
@@ -127,7 +133,7 @@ computeIndics <- function(sppoints,popcol="totalPop",emissioncol="CO2",gamma=1,d
 #'   TODO : influence of normalizing by pmax vs ptot ? -> investigate in interaction models
 #'   - normalized such that sum = 1 ? NO to be able to have a size effects in comparing configurations
 interactionPotential <- function(sppoints,gravityGamma=1,gravityDistance=1,weightcol="totalPop"){
-  dmat = exp(-st_distance(sppoints)/gravityDistance)
+  dmat = exp(-drop_units(st_distance(sppoints))/gravityDistance)
   diag(dmat)<-0
   dmat = Matrix(dmat)
   #pops = sppoints[[weightcol]]

### 2018-10-08 Models/Percolation/transportationFunctions.R

@@ -1,0 +1,21 @@
+
+library(sf)
+library(igraph)
+
+source('percolationFunctions.R')
+
+#'
+#'
+#'
+transportEmissionsMSE <- function(indics,gamma,decay,xcol="lonmin",ycol="latmin",popcol="totalPop"){
+  # generate flows
+  sppoints <- st_as_sf(d, coords = c(xcol, ycol), crs = 4326) %>% st_transform(3035)
+  flows = interactionPotential(sppoints,gamma,decay)
+
+  # assignate : must first construct the graph : aggregate at this stage ?
+  
+  #graph_from_adj_list(,mode = 'all')
+  
+  
+  
+}

### 2018-10-11 Models/Percolation/percolation.R
.

@@ -7,42 +7,42 @@ library(dplyr)
 source('percolationFunctions.R')
 
 # assumes data has been consolidated before
-load('Data/consolidated/indics.RData')
+load('../../Data/consolidated/indics.RData')
 
 # parameter values
 quantiles = c(0.75,0.8,0.85,0.9,0.95)
 radiuses=c(8000,10000,15000,20000,50000,75000,100000)
 nwindics= c("ecount","mu","vcount","euclPerf")
+gammas=c(1)
+decays=c(500)
 
-params=matrix(0,length(quantiles)*(length(quantiles)+1)*length(radiuses)*length(nwindics),4)
+params=matrix(0,length(quantiles)*(length(quantiles)+1)*length(radiuses)*length(nwindics)*length(gammas)*length(decays),6)
 i=1
-for(nwcol in nwindics){for(nwthq in c(0,quantiles)){for(popthq in quantiles){
-  for(radius in radiuses){
-    params[i,]=c(nwcol,popthq,nwthq,radius)
+for(nwcol in nwindics){for(nwthq in c(0,quantiles)){for(popthq in quantiles){for(radius in radiuses){for(gamma in gammas){for(decay in decays){
+    params[i,]=c(nwcol,popthq,nwthq,radius,gamma,decay)
     i=i+1
-  }
-}}}
+  }}}}}}
 params=as.data.frame(params)
-colnames(params)=c("nwcol","popthq","nwthq","radius")
-params$nwthq=as.numeric(as.character(params$nwthq));params$popthq=as.numeric(as.character(params$popthq));params$radius=as.numeric(as.character(params$radius))
+colnames(params)=c("nwcol","popthq","nwthq","radius","gamma","decay")
+params$nwthq=as.numeric(as.character(params$nwthq));params$popthq=as.numeric(as.character(params$popthq));params$radius=as.numeric(as.character(params$radius));params$gamma=as.numeric(as.character(params$gamma));params$decay=as.numeric(as.character(params$decay))
 
 library(doParallel)
-cl <- makeCluster(50,outfile='log')
+cl <- makeCluster(20,outfile='log')
 registerDoParallel(cl)
 res <- foreach(i=1:nrow(params)) %dopar% {
   #source('Models/Percolation/percolationFunctions.R')
   source('percolationFunctions.R')
   #p = conditionalPercolation(d=indics,popthq=params$popthq[i],nwthq=params$nwthq[i],radius=params$radius[i])
[…]

### 2018-10-11 Models/Percolation/percolationFunctions.R

@@ -28,11 +28,11 @@ conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
   #popthq=0.95;nwthq=0.95;radius=50000;nwcol="euclPerf";popcol="totalPop";minclustsize=5
   #xcol="lonmin";ycol="latmin";popcol="totalPop";nwcol="mu";d=indics
   
-  show(paste0("radius = ",radius))
+  show(paste0("Parameter : radius = ",radius,' ; popthq = ',popthq,' ; nwcol = ',nwcol,' ; nwthq = ',nwthq,' ; gamma = ',gamma,' ; decay = ',decay))
   
   popth = quantile(d[[popcol]],c(popthq));nwth=quantile(d[[nwcol]],c(nwthq))
   
-  show(paste0("popth = ",popth));show(paste0("nwth = ",nwth));
+  show(paste0("Values : popth = ",popth));show(paste0("nwth = ",nwth));
   
   sppoints <- st_as_sf(d, coords = c(xcol, ycol), crs = 4326) %>% st_transform(3035)
   sppoints$index=1:nrow(sppoints)
@@ -81,12 +81,15 @@ conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
   #  }
   #}
     
-  return(computeIndics(sppoints,gamma=gamma,decay=decay))
+  return(computeIndics(sppoints,gamma,decay))
   #return(sppoints)
 }
 
 
+#'
+#' Aggregatd indices
 aggregIndics <- function(res){
+  show(paste0('Aggregating indicators for result with ',length(res$areas),' clusters'))
   return(list(
     area=sum(res$areas),
     pop=sum(res$areas),
@@ -94,7 +97,7 @@ aggregIndics <- function(res){
     avgdist=mean(res$avgdists),
     entropy=mean(res$entropies),
     slope=mean(res$slopes),
-    efficiency=sum(efficiency),
+    efficiency=sum(efficiencies),
     emissions=sum(emissions)
   ))
 }
@@ -103,24 +106,29 @@ aggregIndics <- function(res){
 
 #'
 #'
-computeIndics <- function(sppoints,popcol="totalPop",emissioncol="CO2",gamma=1,decay=1){
+computeIndics <- function(sppoints,gamma,decay,popcol="totalPop",emissioncol="CO2"){
[…]

### 2018-10-11 Models/Percolation/percolation.R
.

@@ -7,7 +7,7 @@ library(dplyr)
 source('percolationFunctions.R')
 
 # assumes data has been consolidated before
-load('Data/consolidated/indics.RData')
+load('../../Data/consolidated/indics.RData')
 
 # parameter values
 quantiles = c(0.75,0.8,0.85,0.9,0.95)

### 2018-10-14 Models/Percolation/percolation.R
.

@@ -27,7 +27,7 @@ colnames(params)=c("nwcol","popthq","nwthq","radius","gamma","decay")
 params$nwthq=as.numeric(as.character(params$nwthq));params$popthq=as.numeric(as.character(params$popthq));params$radius=as.numeric(as.character(params$radius));params$gamma=as.numeric(as.character(params$gamma));params$decay=as.numeric(as.character(params$decay))
 
 library(doParallel)
-cl <- makeCluster(20,outfile='log')
+cl <- makeCluster(60,outfile='log')
 registerDoParallel(cl)
 res <- foreach(i=1:nrow(params)) %dopar% {
   #source('Models/Percolation/percolationFunctions.R')

### 2018-10-14 Models/Percolation/morphology.R
graph_algo

@@ -19,12 +19,13 @@ library(Matrix)
 
 #'
 #'
-moranPoints <-function(points,varcol){
-  if(nrow(points)<10000){sppoints = points}else{
-    th = quantile(points[[weightcol]],c(1 - 10000/nrow(points)))
+moranPoints <-function(points,varcol,dmat=NULL){
+  show(paste0('Moran for ',nrow(points),' points'))
+  if(nrow(points)<5000){sppoints = points}else{
+    th = quantile(points[[weightcol]],c(1 - 5000/nrow(points)))
     sppoints = points[points[[weightcol]]>th,]
   }
-  weights=spatialWeightsPoints(sppoints)
+  if(is.null(dmat)){weights=spatialWeightsPoints(sppoints)}else{weights = Matrix(1/dmat);diag(weights)<-0}
   v = sppoints[[varcol]]
   v = v-mean(v)
   num = sum(Diagonal(x = v)%*%weights%*%Diagonal(x = v))
@@ -48,13 +49,14 @@ spatialWeightsPoints<-function(points){
 
 #'
 #'
-avgDistancePoints<-function(points,varcol){
-  if(nrow(points)<10000){sppoints = points}else{
-    th = quantile(points[[weightcol]],c(1 - 10000/nrow(points)))
+avgDistancePoints<-function(points,varcol,dmat=NULL){
+  show(paste0('Avg dist for ',nrow(points),' points'))
+  if(nrow(points)<5000){sppoints = points}else{
+    th = quantile(points[[weightcol]],c(1 - 5000/nrow(points)))
     sppoints = points[points[[weightcol]]>th,]
   }
   v = sppoints[[varcol]]
-  distances <- drop_units(st_distance(sppoints))
+  if(is.null(dmat)){distances <- drop_units(st_distance(sppoints))}else{distances = dmat}
   #distances[distances==Inf]=0
   #distances[distances<0.01]=0
   #vv = (Diagonal(x=v/sum(v^2))%*%Matrix(distances,sparse=T))%*%Diagonal(x=v/sum(v^2))

### 2018-10-14 Models/Percolation/percolation.R

@@ -10,15 +10,21 @@ source('percolationFunctions.R')
 load('../../Data/consolidated/indics.RData')
 
 # parameter values
-quantiles = c(0.75,0.8,0.85,0.9,0.95)
-radiuses=c(8000,10000,15000,20000,50000,75000,100000)
+#quantiles = c(0.85,0.9,0.95)
+popquantiles=c(0.9,0.95)
+nwquantiles=c(0.0,0.4,0.8)
+radiuses=c(8000,10000,15000,20000,50000)
+#radiuses=c(8000)
 nwindics= c("ecount","mu","vcount","euclPerf")
+#nwindics= c("ecount")
 gammas=c(0.5,1,1.5)
-decays=c(10,100,500,1000)
+#gammas=c(1)
+decays=c(1000,10000,50000,100000)
+#decays=c(1000)
 
-params=matrix(0,length(quantiles)*(length(quantiles)+1)*length(radiuses)*length(nwindics)*length(gammas)*length(decays),6)
+params=matrix(0,length(popquantiles)*length(nwquantiles)*length(radiuses)*length(nwindics)*length(gammas)*length(decays),6)
 i=1
-for(nwcol in nwindics){for(nwthq in c(0,quantiles)){for(popthq in quantiles){for(radius in radiuses){for(gamma in gammas){for(decay in decays){
+for(nwcol in nwindics){for(nwthq in nwquantiles){for(popthq in popquantiles){for(radius in radiuses){for(gamma in gammas){for(decay in decays){
     params[i,]=c(nwcol,popthq,nwthq,radius,gamma,decay)
     i=i+1
   }}}}}}
@@ -27,14 +33,16 @@ colnames(params)=c("nwcol","popthq","nwthq","radius","gamma","decay")
 params$nwthq=as.numeric(as.character(params$nwthq));params$popthq=as.numeric(as.character(params$popthq));params$radius=as.numeric(as.character(params$radius));params$gamma=as.numeric(as.character(params$gamma));params$decay=as.numeric(as.character(params$decay))
 
 library(doParallel)
-cl <- makeCluster(20,outfile='log')
+cl <- makeCluster(50,outfile='log')
 registerDoParallel(cl)
 res <- foreach(i=1:nrow(params)) %dopar% {
-  #source('Models/Percolation/percolationFunctions.R')
+#res=list()
+#for(i in 1:nrow(params)){
+    #source('Models/Percolation/percolationFunctions.R')
   source('percolationFunctions.R')
[…]

### 2018-10-14 Models/Percolation/percolationFunctions.R

@@ -4,8 +4,10 @@ library(dplyr)
 #library(rgeos)
 library(sf)
 #library(tidyverse)
-#library(ggplot2)
+library(ggplot2)
 library(Matrix)
+library(igraph)
+
 
 #source(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/MorphologyPackage/morphology.R'))
 # for oml : must be in the same dir
@@ -68,6 +70,7 @@ conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
     }
     
     # this is wrong : could have a next cluster larger than minclustsize
+    # totally wrong assumption that cluster will be ordered by size..
     #remainingpoints=(nrow(cpoints)>0)&(nrow(currentcluster)>minclustsize)
     remainingpoints=(nrow(cpoints)>0)
     }else{
@@ -96,6 +99,53 @@ conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
 
 
 #'
+#' Note : determinstic graph assumes deterministic percolation.
+#'   precise assumptions of road nw percolation
+graphPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
+                             minclustsize=25,
+                             xcol="lonmin",ycol="latmin",popcol="totalPop",
+                             withMaps=T,
+                             resdir=paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/Maps2/')
+){
+
+  if(withMaps==T){dir.create(resdir)}
+  show(paste0("Parameters : radius = ",radius,' ; popthq = ',popthq,' ; nwcol = ',nwcol,' ; nwthq = ',nwthq,' ; gamma = ',gamma,' ; decay = ',decay))
+  popth = quantile(d[[popcol]],c(popthq));nwth=quantile(d[[nwcol]],c(nwthq))
+  show(paste0("Values : popth = ",popth," ; nwth = ",nwth));
+  sppoints <- st_as_sf(d, coords = c(xcol, ycol), crs = 4326) %>% st_transform(3035)
+  
+  cpoints = sppoints[sppoints[[popcol]]>popth&sppoints[[nwcol]]>nwth,]
+  cpoints$cluster = rep(NA,nrow(cpoints))
+  
+  dmat = drop_units(st_distance(cpoints))
+  adjmat = Matrix(dmat<=radius,sparse=T)
[…]

### 2018-10-15 Models/Percolation/morphology.R
fix_sampledmorpho

@@ -22,8 +22,8 @@ library(Matrix)
 moranPoints <-function(points,varcol,dmat=NULL){
   show(paste0('Moran for ',nrow(points),' points'))
   if(nrow(points)<5000){sppoints = points}else{
-    th = quantile(points[[weightcol]],c(1 - 5000/nrow(points)))
-    sppoints = points[points[[weightcol]]>th,]
+    th = quantile(points[[varcol]],c(1 - 5000/nrow(points)))
+    sppoints = points[points[[varcol]]>th,]
   }
   if(is.null(dmat)){weights=spatialWeightsPoints(sppoints)}else{weights = Matrix(1/dmat);diag(weights)<-0}
   v = sppoints[[varcol]]
@@ -52,8 +52,8 @@ spatialWeightsPoints<-function(points){
 avgDistancePoints<-function(points,varcol,dmat=NULL){
   show(paste0('Avg dist for ',nrow(points),' points'))
   if(nrow(points)<5000){sppoints = points}else{
-    th = quantile(points[[weightcol]],c(1 - 5000/nrow(points)))
-    sppoints = points[points[[weightcol]]>th,]
+    th = quantile(points[[varcol]],c(1 - 5000/nrow(points)))
+    sppoints = points[points[[varcol]]>th,]
   }
   v = sppoints[[varcol]]
   if(is.null(dmat)){distances <- drop_units(st_distance(sppoints))}else{distances = dmat}

### 2018-10-15 Models/Percolation/percolation.R
res

@@ -12,7 +12,7 @@ load('../../Data/consolidated/indics.RData')
 # parameter values
 #quantiles = c(0.85,0.9,0.95)
 popquantiles=c(0.9,0.95)
-nwquantiles=c(0.0,0.4,0.8)
+nwquantiles=c(0.0,0.8,0.9)
 radiuses=c(8000,10000,15000,20000,50000)
 #radiuses=c(8000)
 nwindics= c("ecount","mu","vcount","euclPerf")

### 2018-10-16 Models/Percolation/percolation.R
res_refined

@@ -11,15 +11,15 @@ load('../../Data/consolidated/indics.RData')
 
 # parameter values
 #quantiles = c(0.85,0.9,0.95)
-popquantiles=c(0.9,0.95)
-nwquantiles=c(0.0,0.8,0.9)
+popquantiles=c(0.85,0.9,0.95)
+nwquantiles=c(0.0,0.8,0.9,0.95)
 radiuses=c(8000,10000,15000,20000,50000)
 #radiuses=c(8000)
 nwindics= c("ecount","mu","vcount","euclPerf")
 #nwindics= c("ecount")
-gammas=c(0.5,1,1.5)
+gammas=c(0.5,1,1.5,2.0)
 #gammas=c(1)
-decays=c(1000,10000,50000,100000)
+decays=c(100,1000,10000,50000,100000)
 #decays=c(1000)
 
 params=matrix(0,length(popquantiles)*length(nwquantiles)*length(radiuses)*length(nwindics)*length(gammas)*length(decays),6)

### 2018-10-16 Models/Percolation/percolationFunctions.R
plots

@@ -148,19 +148,34 @@ graphPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
 #'
 #' Aggregatd indices
 aggregIndics <- function(res){
-  show(paste0('Aggregating indicators for result with ',length(res$areas),' clusters'))
-  return(list(
+  #show(paste0('Aggregating indicators for result with ',length(res$areas),' clusters'))
+  return(c(
     area=sum(res$areas),
-    pop=sum(res$areas),
+    pop=sum(res$pops),
+    avgpop=mean(res$pops),
+    pophierarchy=hierarchy(res$pops),
     moran=mean(res$morans),
     avgdist=mean(res$avgdists),
     entropy=mean(res$entropies),
     slope=mean(res$slopes),
     efficiency=sum(res$efficiencies),
-    emissions=sum(res$emissions)
+    avgefficiency=mean(res$efficiencies),
+    emissions=sum(res$emissions),
+    avgemissions=mean(res$emissions),
+    totalemissions=sum(res$totalemissions),
+    avgtotemissions=mean(res$totalemissions)
   ))
 }
 
+hierarchy <- function(x){
+  x=x[x>0]
+  if(length(x)>1){
+    #show(log(1:length(x)));show(log(sort(x,decreasing = T)))
+    d=data.frame(rank=log(1:length(x)),size=log(sort(x,decreasing = T)))
+  reg = lm(size~rank,data=d)
+  return(reg$coefficients[2])
+  }else{return(0)}
+}
 
 
 #'

### 2018-10-16 Models/Percolation/percolationResults.R

@@ -1,0 +1,232 @@
+
+# load results from percolation
+
+resstamp = 'directSampling_20181016_024956'
+load(paste0('res/',resstamp,'.RData'))
+resdir = paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/',resstamp,'/');dir.create(resdir)
+
+source(paste0(Sys.getenv('CS_HOME'),'/Organisation/Models/Utils/R/plots.R'))
+
+
+totalPop = sum(indics$totalPop)
+
+#####
+## full areas stats
+
+for(k in 1:length(res)){res[[k]]$paramrow=rep(k,length(res[[k]]$pops))}
+full = data.frame(matrix(unlist(lapply(res,function(l){l$areas=NULL;unlist(data.frame(t(data.frame(l))))})),byrow=T,ncol=length(res[[1]])-1))
+colnames(full)<-names(res[[1]])[-1]
+params$paramrow=1:nrow(params)
+full=left_join(full,params)
+
+full$relemissions=rep(nrow(full),0)
+full$relefficiency=rep(nrow(full),0)
+for(k in unique(full$paramrow)){
+  full$relemissions[full$paramrow==k]=(max(full$emissions[full$paramrow==k])-full$emissions[full$paramrow==k])/(max(full$emissions[full$paramrow==k])-min(full$emissions[full$paramrow==k]))
+  full$relefficiency[full$paramrow==k]= 1 - (max(full$efficiencies[full$paramrow==k])-full$efficiencies[full$paramrow==k])/(max(full$efficiencies[full$paramrow==k])-min(full$efficiencies[full$paramrow==k])) 
+}
+
+# within config pareto fronts ?
+popthq=c(0.85,0.95);nwthq=c(0.0,0.95);radius=c(8000,50000)
+g=ggplot(full[full$popthq%in%popthq&full$nwthq%in%nwthq&full$radius%in%radius,],aes(x=relemissions,y=relefficiency,color=paste0(popthq," ; ",nwthq," ; ",radius),size= pops/totalPop))
+g+geom_point(alpha=0.5)+facet_wrap(~paste0(decay," ; ",gamma),scales="free")+
+  scale_color_discrete(name=expression(theta[P]*" ; "*theta[N]*" ; "*r[0]))+scale_size_continuous(name="Relative population")+
[…]

### 2018-10-17 Docs/Marami2018/Communication/marami2018.tex
consolidated_data
data_sources
more_figs

@@
 
-\sframe{Networks and territories}{
+\sframe{Morphologies of networks and territories}{
 
@@
 
+% -> visu OSM : see Metropolitan region ; plus positioning of Avignon ; Q of sustainibility ?
+
 \centering
 
-	%\includegraphics[width=0.52\textwidth]{figures/}
+	\includegraphics[width=1.2\textwidth]{figures/roadssouth.png}
 
 	\footnotesize
-%\textit{Source: Wikipedia}
+\textit{Source: OpenStreetMap}
 
@@
 %The structure of road networks both translates its past growth dynamics and has a significant impact on the sustainability of territories it irrigates.
+
+% here extracts from Claire's work
+% Lagesse, C., Bordin, P., & Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. -> nice picture Avignon
+
+\textit{Multiple dimensions to characterize road networks}
+
+\medskip
+
+{
+\centering
+\includegraphics[width=\textwidth,height=0.7\textheight]{figures/avignon_claire.png}
+}
+
+\footnotesize
+
+Lagesse, C., Bordin, P., \& Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. \cite{lagesse2015spatial}
 
@@
 % from monodimensional to multidimensional
+
+% similar to 
+%Cottineau, C., Finance, O., Hatna, E., Arcaute, E., & Batty, M.
+(2018).
+Defining urban clusters to detect agglomeration economies.
+Environment and Planning B: Urban Analytics and City Science, 2399808318755146.
+
+%
+
+\justify
+
+\textbf{Network percolation: } \textit{progressive occupation/connection of nodes of a network} \cite{callaway2000network}
+
+\bigskip
+
+Application to the study of cities:
+
+\begin{itemize}
+	\item modeling urban growth \cite{makse1998modeling}
+	\item endeogenous determination of regions \cite{arcaute2016cities}
+	\item characterization of spatial point patterns \cite{huynh2018characterisation}
+\end{itemize}
+
+\bigskip
+
+\textit{Towards complementary dimensions to condition road network percolation}
+
+$\rightarrow$ similar to \cite{cottineau2018defining} to define urban areas
 
@@
 This suggests policies in terms of regional integration to increase the sustainability of mega-city regions.
+
+}
+
+\sframe{Extrapolating transportation flows}{
+
+% detail method to extrapolate parameters of gravity flows
+
+% interesting result : compare these emission with the synthetic ?
+(then just a refinment ?) -> yes, better.
+show new Pareto fronts with this fixed parameters for only emissions, and varying for economic ?
 

### 2018-10-17 Models/Percolation/percolationResults.R

@@ -102,25 +102,52 @@ for(gamma in gammas){for (decay in decays){
 
 
 g=ggplot(sres,aes(x=emissions,y=efficiency,col=pop/totalPop))
-g+geom_point()+facet_grid(decay~gamma)+xlab("Relative potential emissions")+ylab("1 - Relative potential efficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
+g+geom_point()+facet_grid(decay~gamma)+xlab("Normalized potential emissions")+ylab("1 - Normalized potential efficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_paretos.png'),width=30,height=25,units='cm')
+
+g=ggplot(sres,aes(x=emissions*totalPop/pop,y=efficiency*totalPop/pop,col=pop/totalPop))
+g+geom_point()+facet_grid(decay~gamma)+xlab("Relative potential emissions")+ylab("Relative potential unefficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
+ggsave(file=paste0(resdir,'aggreg_paretos_relative.png'),width=30,height=25,units='cm')
+
+
 
 # test : which params produce pareto fronts ?
 g=ggplot(sres,aes(x=emissions,y=efficiency,col=nwcol))
-g+geom_point(alpha=0.5)+facet_grid(decay~gamma)+xlab("Relative potential emissions")+ylab("1 - Relative potential efficiency")+scale_color_discrete(name="Percolation\nIndicator")+stdtheme
+g+geom_point(alpha=0.5)+facet_grid(decay~gamma)+xlab("Normalized potential emissions")+ylab("1 - Normalized potential efficiency")+scale_color_discrete(name="Percolation\nIndicator")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_paretos_nwcol.png'),width=30,height=25,units='cm')
 
+g=ggplot(sres,aes(x=emissions*totalPop/pop,y=efficiency*totalPop/pop,col=nwcol))
+g+geom_point(alpha=0.5)+facet_grid(decay~gamma)+xlab("Relative potential emissions")+ylab("Relative potential unefficiency")+scale_color_discrete(name="Percolation\nIndicator")+stdtheme
[…]

### 2018-10-17 Docs/Marami2018/Communication/marami2018.tex
slides

@@
 
-$\rightarrow$
+$\rightarrow$ Need to combine morphological and functional dimensions of cities \cite{burger2012form}
 
@@
 
-$\rightarrow$
+$\rightarrow$ Interaction between networks and territories to capture the form-function link \cite{raimbault2018caracterisation}; potential application to sustainability of urban systems
 
@@
 
-\textbf{Research objective : } \textit{}
+\textbf{Research objective : } \textit{Investigate a multi-dimensional percolation of territorial networks taking into account urban morphology and road network topology; endogenous characterization of urban regions.}
 
@@
 
-\sframe{Formalization}{
+\sframe{Multilayer percolation}{
 
@@
 Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold.
+
+\textit{Multi-dimensional network percolation heuristic, similar to multilayer percolation} \cite{boccaletti2014structure}
+
+\bigskip
+
+\centering
+\includegraphics[width=0.8\textwidth]{figures/principle.pdf}
+
+\bigskip
+
+\raggedright
+
+\textbf{Parameters: } percolation radius $r_0$, percolation thresholds $\theta_i$ for each layer
 
@@
 More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}.
-We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.
+
+\justify
+
+\textit{Territorial indicators computed for Europe by \cite{raimbault2018urban}}
+
+% here add maps to better visualize the spatial fields
+
+\medskip
+
+\includegraphics[width=0.49\textwidth]{figures/indics_morpho_discrquantiles.png}
+\includegraphics[width=0.49\textwidth]{figures/indics_network_en_areasize100_offset50_factor0_5.png}
+
+\medskip
+
+\textbf{Population distribution morphology} and \textbf{Network topology} (betweenness, closeness, clustering, efficiency, \ldots) computed on 50km spatial windows (Eurostat density grid and OpenStreetMap)
 
@@
 
-\sframe{Experience plan}{
+\sframe{Network construction}{
+
+%We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.
+
+\justify
+
+\textbf{Two layers: } population density (threshold $\theta_P$) and network characteristics (threshold $\theta_N$) taken among \{Number of edges, Number of vertices, Cyclomatic number $\mu$, Euclidian efficiency $v$ \}; percolated with a radius $r_0$
+
+% banos2012towards
+
+\bigskip
+
+\textbf{Rationale: } \textit{two locations will be in relation if they are close, have a high population density and given network characteristics.}
+
+% note : euclperf should be minimized whereas other maximized : issue ?
+
+\bigskip
+
+\textbf{Implementation: } construction of a single layer spatial network given the condition on the two layers and distances; extraction of connected components.
+
+\bigskip
+
+\textbf{Experience plan: } grid sampling for $r_0,\theta_P,\theta_N$ and network variables; additional gravity potential parameters $\gamma,d_0$ (detailed after)
+
+$\rightarrow$ 4800 parameter points
 
@@
 
-% + explain openmole
+% + explain openmole -> later in potential developments
 
@@
 %Maps reveal that most configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}.
+
+\justify
+
+\vspace{-0.4cm}
+\textit{Extraction of endogenous polycentric mega-city regions \cite{hall2006polycentric}}
+
+\medskip
+
+\centering
+
+% note : forgot r_0 in figs
+\includegraphics[width=\textwidth,height=0.9\textheight]{figures/totalPop4183694_00056402_ecount850_radius8000.png}
+
+}
+
+\sframe{Different endogenous morphologies}{
+
+% TODO for the paper : check effect of adding network threshold : methodo info on multilayer percol in itself
+
+\hspace{-0.5cm}
+\includegraphics[width=0.52\textwidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}
+\includegraphics[width=0.52\textwidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}
 
@@
 
+% explain indicators : convex hull.
+
+\justify
+
+\textbf{Application: } sustainability indicators for the endogenous urban regions; proxys for two conflicting dimensions: GHG emissions and economic integration \cite{viguie2012trade}.
+
+\bigskip
+
+\textbf{Data: } EDGAR database for GHG emissions (v4.3.2)
+
+\cite{janssens2017edgar}
+
+\bigskip
+
+\textbf{Estimation: } Abstract flows approximated with a gravity model
+
+\[
+\phi_{ij} = \left(\frac{v_i v_j}{(\sum_k v_k)^2}\right)^\gamma \cdot \exp\left(\frac{-d_{ij}}{d_0}\right)
+\]
+
+where $v_k$ are either effective local GHG emissions or population (economic activity scaling law of population \cite{bettencourt2007growth})
+
+\medskip
+
+$\rightarrow$ sum of flows within the geographical span of the cluster (convex hull) approximate potential emissions and economic activity
+
 }
 
-\sframe{Results: Pareto fronts}{
+\sframe{Pareto fronts for sustainability}{
 
@@
 
-}
+\textit{Superposing Pareto front for observed population and emissions, on all clusters.}
 
-\sframe{Extrapolating transportation flows}{
+\centering
 
-% detail method to extrapolate parameters of gravity flows
+\medskip
 
-% interesting result : compare these emission with the synthetic ?
-(then just a refinment ?) -> yes, better.
-show new Pareto fronts with this fixed parameters for only emissions, and varying for economic ?
+\includegraphics[height=0.8\textheight]{figures/full_effective_pareto.png}
+
+% -> full fronts in reserve slide
 
@@
 
-\sframe{Calibration}{
+\sframe{Pareto fronts for aggregated indicators}{
+
+\textit{Aggregated sustainability indicators suggest some configurations are more Pareto efficient (high $\gamma$ regime, activities with high added value).}
+
+\centering
+
+\medskip
+
+\includegraphics[height=0.8\textheight]{figures/aggreg_paretos_radiuspopthq.png}
+
+}
+
+\sframe{An optimal morphology ?}{
+
+%# Cumulative Proportion  0.7296 0.9650 0.99761 1.00000
+%                PC1       PC2         PC3         PC4
+%moran   -0.3088585 0.9493848 -0.04444327  0.03605266
+%avgdist  0.5417362 0.1415668 -0.82239570 -0.10072759
+%entropy  0.5108424 0.2140447  0.45847647 -0.69499942
+%slope    0.5917502 0.1811415  0.33390034  0.71100630
+
+\textit{Optimal degree of monocentricity of the system for emissions.}
+
+\centering
+
+\medskip
+
+\includegraphics[height=0.85\textheight]{figures/aggreg_morpho_pc1-emissions.png}
+
+}
+
+\sframe{Morphological trade-offs}{
+
+\vspace{-0.3cm}
+
+\textit{No ``optimal'' cities, different forms yield different compromises in terms of relative indicators.}
+
+\centering
+
+\medskip
+
+\includegraphics[height=0.85\textheight]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale.png}
+
+}
+
+\sframe{Towards a systematic calibration}{
 
@@
 % -> depending on results
+
+Grid sampling to explore regions rapidly limited
+
+$\rightarrow$ towards the use of genetic algorithms on grid, made smooth with the OpenMOLE software \url{https://next.openmole.org/}
+
+\centering
+
+\smallskip
+
+\includegraphics[height=0.35\textheight]{figures/openmole.png}
+
+\smallskip
+
+\raggedright\justify
+
+\footnotesize \textit{OpenMOLE: (i) embed any model as a black box; (ii) transparent access to main High Performance Computing environments; (iii) model exploration and calibration methods.}
+
+\bigskip
+
+\textbf{Apply to the summer school !} \texttt{https://exmodelo.org/}
 
@@
 
-$\rightarrow$ 
+$\rightarrow$ Transfer to policy making recommandations: Pareto-optimal configuration can be used for the planning of regional transportation networks, subsidies policies, etc.
 \smallskip
@@
 
-$\rightarrow$
+% detail method to extrapolate parameters of gravity flows
+
+% interesting result : compare these emission with the synthetic ?
+(then just a refinment ?) -> yes, better.
+show new Pareto fronts with this fixed parameters for only emissions, and varying for economic ?
+
+$\rightarrow$ Extrapolation of transportation flows
 
@@
 
+\sframe{Results: effective emissions}{
+
+\textit{Effective emissions exhibit a supralinear scaling of population}
+
+\medskip
+
+\centering
+
+\includegraphics[height=0.9\textheight]{figures/aggreg_effective.png}
+
+}
+
+\sframe{Results: all clusters Pareto fronts}{
+
+\textit{Variation of Pareto front patterns when potential parameter $\gamma,d_0$ vary.}
+
+\medskip
+
+\centering
+
+\includegraphics[width=0.9\textwidth]{figures/full_paretos.png}
+
+}
+
+\sframe{Results: an optimal morphology}{
+
+\textit{More monocentric areas are more optimal in terms of relative emissions and efficiency ?}
+
+\medskip
+
+\centering
+
+\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relefficiency.png}
+\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relemissions.png}
+
+% TODO : curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.
+
+}
+
 %%%%%%%%%%%%%%%%%%%%%

### 2018-10-17 Models/Percolation/percolationResults.R

@@ -220,9 +220,20 @@ g=ggplot(sres,aes(x=PC1,y=efficiency,col=pop/totpop))
 g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC1")+ylab("1 - Normalized potential efficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_morpho_pc1-efficiency.png'),width=30,height=25,units='cm')
 
+g=ggplot(sres,aes(x=PC1,y=efficiency,col=pop/totpop))
+g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC1")+ylab("Relative potential unefficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
+ggsave(file=paste0(resdir,'aggreg_morpho_pc1-relefficiency.png'),width=30,height=25,units='cm')
+
+
 g=ggplot(sres,aes(x=PC2,y=efficiency,col=pop/totpop))
 g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC2")+ylab("1 - Normalized potential efficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_morpho_pc2-efficiency.png'),width=30,height=25,units='cm')
+
+g=ggplot(sres,aes(x=PC2,y=efficiency,col=pop/totpop))
+g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC2")+ylab("Relative potential unefficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
+ggsave(file=paste0(resdir,'aggreg_morpho_pc2-relefficiency.png'),width=30,height=25,units='cm')
+
+
 
 g=ggplot(sres,aes(x=emissions,y=efficiency,col=PC1,size=pop/totalPop))
 g+geom_point(alpha=0.5)+facet_grid(decay~gamma)+scale_color_viridis_c(name="PC1")+xlab("Normalized potential emissions")+ylab("1 - Normalized potential efficiency")+scale_size_continuous(name="Relative\nPopulation")+stdtheme

### 2018-10-17 Docs/Marami2018/Communication/marami2018.tex
slides_v1

@@
 
-\institute{$^{1}$Complex Systems Institute, Paris, UPS CNRS 3611 ISC-PIF\\
+\institute{$^{1}$UPS CNRS 3611 ISC-PIF\\
 $^{2}$UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
@@
 
-$\rightarrow$ Interaction between networks and territories to capture the form-function link \cite{raimbault2018caracterisation}; potential application to sustainability of urban systems
+$\rightarrow$ Interactions between networks and territories to capture the link between form and function \cite{raimbault2018caracterisation}; potential application to sustainability of urban systems
 
@@
 
-\textbf{Implementation: } construction of a single layer spatial network given the condition on the two layers and distances; extraction of connected components.
+\textbf{Implementation: } construction of a single layer spatial network given the condition on the two layers and distances, from the 5km resolution indicators spatial field; extraction of connected components.
 
@@
 
-\includegraphics[height=0.85\textheight]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale.png}
+\includegraphics[height=0.83\textheight]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale.png}
 
@@
 
-\textbf{Apply to the summer school !} \texttt{https://exmodelo.org/}
+\textbf{Apply to the summer school !} \url{https://exmodelo.org/}
 
@@
 
-$\rightarrow$ Transfer to policy making recommandations: Pareto-optimal configuration can be used for the planning of regional transportation networks, subsidies policies, etc.
+\medskip
+
+$\rightarrow$ Multi-dimensionality of urban systems and a link between form and function captured through multilayer percolation.
 \smallskip
 
-$\rightarrow$
+$\rightarrow$ Possible transfer to policy-making recommandations: Pareto-optimal configuration can be used for the planning of regional transportation networks, policies for subsidies, etc.
 
@@
 
-$\rightarrow$ Extrapolation of transportation flows
+\medskip
+
+$\rightarrow$ Systematic calibration of parameters to unveil more exhaustive Pareto fronts.
+
+\smallskip
+
+$\rightarrow$ Extrapolation of transportation flows to estimate potential emissions linked to transportation: calibration of gravity model on actual transportation emissions; use of the extrapolated parameters in potentials.
+
+\smallskip
+
+$\rightarrow$ More refined indicators for sustainability (socio-economic integration, accessibilities, different scaling exponents).
+% ex : at least different scaling exponent.
 
@@
 
-\vspace{-1cm}
+\vspace{-0.8cm}
 
@@
 
-$\rightarrow$ %First step towards systematic benchmarks and multi-modeling. \textbf{Need for more systematic model exploration.} % model coupling / benchmarking
+$\rightarrow$ Empirical and theoretical research directed towards concrete policy-making applications. \textbf{Need for more data-driven approaches.}
+%First step towards systematic benchmarks and multi-modeling. \textbf{Need for more systematic model exploration.} % model coupling / benchmarking
 
@@
 
-$\rightarrow$ %Model integration and multi-scalarity ? \textbf{Need for more integrated models.}
+$\rightarrow$ Towards multi-scalar approaches ? \textbf{Need for more integrated models.}
+ %Model integration and multi-scalarity ? \textbf{Need for more integrated models.}
 
@@
 
-$\rightarrow$ %Multiple perspectives on urban systems ? \textbf{Need for more interdisciplinarity.}
+$\rightarrow$ Multidimensionality of urban systems ? \textbf{Need for more interdisciplinarity.}
 
@@
 
-\smallskip
+\medskip
 
-\textbf{Open repository} at \texttt{https://github.com/JusteRaimbault/UrbanMorphology}\\\smallskip
-\textbf{Acknowledgments}: thanks to the \textit{EGI} for access to the infrastructure.
+\textbf{Open repository} at \texttt{https://github.com/JusteRaimbault/UrbanMorphology} (code, data and results)%\\\smallskip
+
+% no grid used
+%\textbf{Acknowledgments}: thanks to the \textit{EGI} for access to the infrastructure.
 
@@
 
-\includegraphics[height=0.9\textheight]{figures/aggreg_effective.png}
+\includegraphics[height=0.83\textheight]{figures/aggreg_effective.png}
 

### 2018-10-17 Models/Percolation/percolationResults.R

@@ -220,7 +220,7 @@ g=ggplot(sres,aes(x=PC1,y=efficiency,col=pop/totpop))
 g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC1")+ylab("1 - Normalized potential efficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_morpho_pc1-efficiency.png'),width=30,height=25,units='cm')
 
-g=ggplot(sres,aes(x=PC1,y=efficiency,col=pop/totpop))
+g=ggplot(sres,aes(x=PC1,y=efficiency*totalPop/pop,col=pop/totpop))
 g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC1")+ylab("Relative potential unefficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_morpho_pc1-relefficiency.png'),width=30,height=25,units='cm')
 
@@ -229,7 +229,7 @@ g=ggplot(sres,aes(x=PC2,y=efficiency,col=pop/totpop))
 g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC2")+ylab("1 - Normalized potential efficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_morpho_pc2-efficiency.png'),width=30,height=25,units='cm')
 
-g=ggplot(sres,aes(x=PC2,y=efficiency,col=pop/totpop))
+g=ggplot(sres,aes(x=PC2,y=efficiency*totalPop/pop,col=pop/totpop))
 g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC2")+ylab("Relative potential unefficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_morpho_pc2-relefficiency.png'),width=30,height=25,units='cm')
 

### 2019-03-10 Docs/Papers/Percolation/paper.tex
skeleton_paper

\documentclass{jimis-en}
\pdfoutput=1 %forces arXiv/HAL to compile with PDFLaTeX

% Packages
\usepackage{array}
\usepackage{pgfplots}
\usepackage{graphicx}

% header
\title{Identifying European Sustainable Mega-city Regions with Multi-dimensional Network Percolation}
% Endogenous urban regions as an optimal compromise between economic integration and greenhouse gases emissions
% A multi-dimensional percolation approach to characterize sustainable mega-city regions

\author[1,2,3,*]{Juste RAIMBAULT}
\affil[1]{UPS CNRS 3611 ISC-PIF, France} 
\affil[2]{CASA, UCL, UK}
\affil[3]{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s}

% corresponding author with his/her email
\corrauthor{juste.raimbault@polytechnique.edu}

%% Information provided by JIMIS:
%ISSN
% DOI 
\doi{10.18713/JIMIS-ddmmyy-v-a}{}
% reviewing process: first date = submission, second date = acceptance 
\review{Month-in-letters Day Year}{Month-in-letters Day Year}
% Volume and Year of the issue
\publication{N}{YYYY}

%% Information provided by the Editors:
% Issue Title 
\issue{Title of the interdisciplinary issue}
% List of k Editors
\editors{First-name-$1$ Family-name-$1$, First-name-$2$ Family-name-$2$, First-name-$k$ Family-name-$k$...}

\begin{document}

\maketitle

\abstract{Territorial structure significantly conditions the sustainability of human settlements. Recent forms of urbanization, in particular integrated mega-city regions, may imply different patterns of economic and transportation flows and thus exhibit various performances regarding different indicators of sustainability. This paper proposes to reconstruct endogenous urban regions from the bottom-up using network percolation, in a multi-dimensional way to take into account both urban form (spatial distribution of population) and urban function (properties of transportation networks). Variable parametrizations allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions). This suggests a customizable spatial design of policies to develop sustainable territories.}

\keywords{Road network; multilayer percolation; mega-city region}

\section{Introduction}

\strut
\vspace{-4ex}

% Morphologies of networks and territories
% - intro : LGV ? Avignon : travaux Claire ! - truc avec les Papes ? 
% https://www.persee.fr/doc/anami_0003-4398_1985_num_97_172_2228_t1_0446_0000_3
% https://www.persee.fr/doc/mefr_0223-5110_1984_num_96_1_2762

% -> visu OSM : see Metropolitan region ; plus positioning of Avignon ; Q of sustainibility ?

% les formes des établissements humains peuvent se percevoir a differentes echelles et selon differentes dimensions. Elle font partie de notre vie courante, a ces mutiples echelles, et pourtant nous sont si etrangeres. Afin d'introduire notre sujet, je propose ici une image familière pour bien d'entre vous qui fréquentent cette région.
% La metropole etendue de la region marseilleise ne se limite pas a l'aspect administratif d'aix marseille, mais englobe bien avignon et la ciotat, voire manosque, des lors que l'on prend un certain recul geographique.
% il n'aura pas manque aux eoils attentifs que ce recul est d'autant plus une abstraction, puisque nous ne represnetons ici uniquement les reseaux routiers, qui s'averent deja etre un proxy suffisant pour estrapoler une densite d'activités humaines.
% Cette presentation s'axera ainsi sur la morphologie des reseaux et des territories et sur leurs relations.
%\includegraphics[width=1.2\textwidth]{figures/roadssouth.png}
%\textit{Source: OpenStreetMap}

The structure of road networks both translates its past growth dynamics and has a significant impact on the sustainability of territories it irrigates. Diverse methods to characterize the structure of spatial networks, and more particularly road networks, have been developed in that context.

% here extracts from Claire's work
% Lagesse, C., Bordin, P., & Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. -> nice picture Avignon

% La caracterisation des reseaux routiers, par exemples en termes de classification et dindicateurs topoliques, est liee a differents enjeux qui incluent la comprehnsion des dynamiques ppasses par la contingence de ces artefacts, amis qussi des questions futues liees a la soutenabilité des territoires qui'ils irriguent.
% les developpements recents, issus d'efforts interdispclinaires, incluent par exmeple ce travail de Claire Lagesse applique a la ville que v ous reconnaissaez bien, pour la definition de nouveau object a la pertineence thematique plus grande que les seuls entites des bases de donnes sans ontologie propre. (ex perso ballade avignon hier : vraie coherence thematique de l'objet "Way")

%\textit{Multiple dimensions to characterize road networks}
%\includegraphics[width=\textwidth,height=0.7\textheight]{figures/avignon_claire.png}

%Lagesse, C., Bordin, P., \& Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. \cite{lagesse2015spatial}

A method to characterize topologies of these spatial networks is network percolation. Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling} and to the analysis of street networks for example to extract endogenous urban regions \citep{arcaute2016cities} or to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.

%Network percolation
% A method to characterize topologies of these spatial networks is network percolation. Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling} and to the analysis of street networks for example to extract endogenous urban regions \citep{arcaute2016cities} or to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.

% from monodimensional to multidimensional

% similar to 
%Cottineau, C., Finance, O., Hatna, E., Arcaute, E., & Batty, M. (2018). Defining urban clusters to detect agglomeration economies. Environment and Planning B: Urban Analytics and City Science, 2399808318755146.

%

% parmi ces outils, la percolation a ete applique recemment aux reseaux routiers.
% en phyique, ce phenomene designe la traversee d'un milieu plus ou omins poreux par un fluide, et par exetnsion les processus pour l'attendre. En pariculier, nous l'entendrons au sens d'une occupation progressice des sites / connections des noeus d'un reseau, jusqu'a un etat critique ou l'ensemble de ceux-ci sont connectés (percolation threshold). [probailite de perco p_c]
% en particulier ppour letude des viulles
% des travaux relaticement anciens ont propose un phenomene de percolation locale correlation pour modeliser l'etalement urbain.
% plus recemment, et plsu proches des travaux des physiciens, une determination endoganee des regions urbaines a etet proposee par l'equipe du acvs en appliquant une heuristqiue de percolation aux reseaux routiers.
% ces methodes peuvent egalement servier a effecteur des statistqieus spatiales.
% nous porposons ici d;'etendre ces methodes en prenant en compte differentes dimensions urbaines, de la meme maniere que cottineau et al combine densite de population et flux de commuting pour definir les regoins urbaines.

%\textbf{Network percolation: } \textit{progressive occupation/connection of nodes of a network} \cite{callaway2000network}

%Application to the study of cities:

%\begin{itemize}
%	\item modeling urban growth \cite{makse1998modeling}
%	\item endeogenous determination of regions \cite{arcaute2016cities}
%	\item characterization of spatial point patterns \cite{huynh2018characterisation}
%\end{itemize}
%\textit{Towards complementary dimensions to condition road network percolation}
%$\rightarrow$ similar to \cite{cottineau2018defining} to define urban areas

% necessity of multi-dimensional percolation

Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.

This paper addresses such a gap by introducing a multi-dimensional percolation heuristic.

%Multidimensional percolation

% Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.

% - why the relation between form and function, aka pop and network in our view is crucial
% - take into account in percolation
% - potential application : endogenous sustainable urban entities

% notre approceh reponde a la question ouverte du lien entre ofrme et fonction da`ns les systemes urbaine ;  et par ailleurs se base sur les interatcions entre reseaux te territoires poour fournire un proy de ce lien.
% nous propospns ainsi d'explorer une heuristique multi-dimensionnelle de la percolation qui prend en compte morplholohie urbaine et yopologie dy reseau routier ; nous l'appliquon pour caracterisere endogenement les regions urbaines.
% [rq : deux niveau de reseuax dans notree apporche : (mais en. fait comme souvent en geo ; ) le physique, et le abstraot que l'on construite a partir d'entites spatiales.

%$\rightarrow$ Need to combine morphological and functional dimensions of cities \cite{burger2012form}

%$\rightarrow$ Interactions between networks and territories to capture the link between form and function \cite{raimbault2018caracterisation}; potential application to sustainability of urban systems

% research question

%\textbf{Research objective : } \textit{Investigate a multi-dimensional percolation of territorial networks taking into account urban morphology and road network topology; endogenous characterization of urban regions.}

We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.

\section{Methods}

\subsection{Multi-dimensional percolation}

Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold. This heuristic is similar to multilayer network percolation \cite{boccaletti2014structure}.

% decrivons a presons de neniere stylisee l'algorithme utilise.
% le processus de peroclation initrial est deterministe, base sur la distance, met est prochess des random eulidian networks evoque hier par Alain Franc : si deux neouds sont a une distance d'un rayon critique r_0, ceux ci sont connectes par un lien.
% nous ajoutons alors sur chaque couche du reaseu un seul de peroclation qui determine si le neod peut etre prus en compte dqns le reseau de la couhce.
% par conjonction des contraintes, nous obtenus un reseau a unique couche dans laquelle simplement els composantes connexes donnent les clusters perfcolés.

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/principle.pdf}}
  \centering
  \label{fig:method}
  \caption{Schematic representation of the multi-dimensional network percolation heuristic.}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%\textbf{Parameters: } percolation radius $r_0$, percolation thresholds $\theta_i$ for each layer

\subsection{Empirical data}

We apply the heuristic to urban morphology and road network topology measures in Europe. More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}.

% Precisons a ppresent les objets geographiques a partir  desqueles nous allons construire le reseau.
% nous rappelns que nous cherchons a combiner prorpiete des territoires, que nous approximerons tres simplement par la districbution spatiale de la population, et propriete des reseaux routiers, que nous approximerons par des indicateirs topollogiques locaux.
% ces differentes caracteristqiues ont ete calucles, dans le cadre de ce travail de modelisation a l'echelle mesocscopique de la coevolution entre reseaux et cilles, sur les feentres glissantes de taille 50km poru l'ensemble de l'europe.
% nous en donnons ici l'illustration dans le cas de la France
% [several infos on different regimes etc]

%\textit{Territorial indicators computed for Europe by \cite{raimbault2018urban}}

% here add maps to better visualize the spatial fields

%\medskip

%\includegraphics[width=0.49\textwidth]{figures/indics_morpho_discrquantiles.png}
%\includegraphics[width=0.49\textwidth]{figures/indics_network_en_areasize100_offset50_factor0_5.png}

%\textbf{Population distribution morphology} and \textbf{Network topology} (betweenness, closeness, clustering, efficiency, \ldots) computed on 50km spatial windows (Eurostat density grid and OpenStreetMap)

\subsection{Sustainability indicators}

We use this endogenous definition of regional urban systems produced by the percolation algorithm to evaluate their sustainability, in terms of conflicting objectives of economic integration and greenhouse gases emissions. The definition of sustainability, or sustainable development, is by essence multi-dimensional \citep{viguie2012trade}. Its characterization as quantitative indicators is even more subject to numerous degrees of freedom. We work here with two stylized indicators for two conflicting dimensions, as a proof-of-concept.

The EDGAR database \citep{janssens2017edgar} (version 4.3.2) is used for local estimates of greenhouse gases emissions.

% explain indicators : convex hull.

Applying a gravity model to each region, we estimate abstract transportation flows within each and extrapolate emissions by coupling with the Edgar emission database \citep{janssens2017edgar} and economic activities with a scaling law of population.  More precisely, greenhouse gases emissions derived from economic and transportation flows are estimated with the following expression

\begin{equation}
\phi_{ij} = \left(\frac{v_i v_j}{(\sum_k v_k)^2}\right)^\gamma \cdot \exp\left(\frac{-d_{ij}}{d_0}\right)
\end{equation}

where $v_k$ are either effective local GHG emissions or population. Indeed, the economic activity follows relatively well scaling laws of populations \cite{bettencourt2007growth}, the exponent being dependant on the activity and the definition of areas on which it is estimated~\citep{cottineau2017diverse}.

The sum of all flows within the geographical span of the cluster (that we approximate as the convex Hull envelope of its points), allows us to approximate the cumulated potential emissions and economic activity.

\section{Results}

\subsection{Implementation}

%Network construction

%We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.

% a partir de ces deux couches nous considerosn d'un [art la densite de population, d'autre part des proxus pour la densite ou perdormance du reseau ; avce les paramteres associes.
% la logique geographique derriere cela est liee a la premiere loi de Tobler : deux lieus proches etc  ; ajoute a la logique de la masse (loi de gravite), mais selon different aspects du reseau (car difficile de savoir quelle dimenion a priori jouera le role "equivalent" d'une densite de population : autant laisser libre et explorer plutot que de fixer arbitrairemenbt - maxime d'un computational scientist converti, bien sur.
% en pratqiuem nous contruison le reseau correspondant et isolns ses composantes connexes.
% descirption du plan d'explericence.

%\textbf{Two layers: } population density (threshold $\theta_P$) and network characteristics (threshold $\theta_N$) taken among \{Number of edges, Number of vertices, Cyclomatic number $\mu$, Euclidian efficiency $v$ \}; percolated with a radius $r_0$

% banos2012towards

%\textbf{Rationale: } \textit{two locations will be in relation if they are close, have a high population density and given network characteristics.}

% note : euclperf should be minimized whereas other maximized : issue ?

%\textbf{Implementation: } construction of a single layer spatial network given the condition on the two layers and distances, from the 5km resolution indicators spatial field; extraction of connected components.

In practice, the analysis is implemented using R and the igraph package. The network is constructed by superposing the population density layer with the network layer.

%\textbf{Experience plan: } grid sampling for $r_0,\theta_P,\theta_N$ and network variables; additional gravity potential parameters $\gamma,d_0$ (detailed after)

The experience plan is a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the network variable considered.

%$\rightarrow$ 4800 parameter points
%We systematically explore the clusters obtained for 840 parameter configurations.

% + explain openmole -> later in potential developments

We systematically explore the clusters obtained for 4800 parameter configurations.

\subsection{Extracting endogenous mega-city regions}

Maps reveal that most configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}. These are here defined endogenously from the bottom-up.

% les rsultats sont particulierement interessantm puiqaue ceratines valuers de apremtres menest a des configurations connues par lkes geographesm celles des mega-cities regions
% [rq : affirnation arbitraire iuci car pas evidence-based, il faudrait comp systematique etc , le probeleme vcest que meme pas de liste avec consensus sur l'obejct, comme soubent en geo>...]
% du coup "visuellemtn" randstad, rehinrhuhrm, rehinmainm London, Paris, Manchester Leeds,

% note : forgot r_0 in figs
%\includegraphics[width=\textwidth,height=0.9\textheight]{figures/totalPop4183694_00056402_ecount850_radius8000.png}

%Different endogenous morphologies

% TODO for the paper : check effect of adding network threshold : methodo info on multilayer percol in itself

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}
  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
  \centering
  \label{fig:exclusters}
  \caption{Two examples of obtained clusters for different parameter values.}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

\subsection{Pareto fronts for sustainability}

We show therein that different population, network and distance thresholds yield different performances in terms of sustainability, exhibiting a Pareto front. This suggests policies in terms of regional integration to increase the sustainability of mega-city regions.

%\textit{Superposing Pareto front for observed population and emissions, on all clusters.}
%\includegraphics[height=0.8\textheight]{figures/}

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/full_effective_pareto.png}}
  \centering
  \label{fig:paretos}
  \caption{Point clouds of region-level indicators, namely population and emissions, for different parametrizations.}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

% -> full fronts in reserve slide

% Pareto fronts for aggregated indicators}{
%\textit{Aggregated sustainability indicators suggest some configurations are more Pareto efficient (high $\gamma$ regime, activities with high added value).}

%\includegraphics[height=0.8\textheight]{figures/aggreg_paretos_radiuspopthq.png}

%%%%%%%%%%%%%
\subsection{Linking urban morphology and sustainability}

%# Cumulative Proportion  0.7296 0.9650 0.99761 1.00000
%                PC1       PC2         PC3         PC4
%moran   -0.3088585 0.9493848 -0.04444327  0.03605266
%avgdist  0.5417362 0.1415668 -0.82239570 -0.10072759
%entropy  0.5108424 0.2140447  0.45847647 -0.69499942
%slope    0.5917502 0.1811415  0.33390034  0.71100630

When considering the aggregated indicators for a parametrization of endogenous city regions, one can relate them to morphological indicators, computed by~\cite{}, that we average on areas. This establishes a link between urban morphology and sustainibility. A principal component analysis on considered points yield 96\% of variance with two components, and 73\% explained by the first component alone. The first component relates to a level of monocentricity ($PC1 = -0.3\cdot I + 0.54 \cdot \bar{d} + 0.51\cdot \varepsilon + 0.59 \cdot \alpha$). As shown in Fig.~\ref{fig:emissions-pc1}, there seems to exist an optimal intermediate level of monocentricity regarding emissions only, except for long-range and low-hierarchy interactions. However, when considering both emissions and economic indicators, urban form then acts as a compromise variable, moving points on Pareto fronts, as shown in Fig.~\ref{fig:paretos-relative}. In some case, highly monocentric areas can be a good compromise, whereas the intermediate optimal for emissions may yield highly inefficient areas.

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_pc1-emissions.png}}
  \centering
  \label{fig:emissions-pc1}
  \caption{Aggregated values of normalized potential emissions, as a function of the first morphological principal component (PC1), for varying values of parameters $d_G$ and $\gamma_G$. As PC1 is mainly linked to monocentricity, there seems to exist an optimal intermediate level of monocentricity for emissions alone.}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%Morphological trade-offs

%\textit{No ``optimal'' cities, different forms yield different compromises in terms of relative indicators.}

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale.png}}
  \centering
  \label{fig:paretos-relative}
  \caption{Relative potential emissions against relative potential economic unefficiency (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows). Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

\section{Discussion}

Further work can consist in the use of calibration heuristics \citep{reuillon2013openmole} to find in a more robust way optimal parameter values.

% -> depending on results

%Grid sampling to explore regions rapidly limited

%$\rightarrow$ towards the use of genetic algorithms on grid, made smooth with the OpenMOLE software \url{https://next.openmole.org/}
%\includegraphics[height=0.35\textheight]{figures/openmole.png}
%\footnotesize \textit{OpenMOLE: (i) embed any model as a black box; (ii) transparent access to main High Performance Computing environments; (iii) model exploration and calibration methods.}
%\textbf{Apply to the summer school !} \url{https://exmodelo.org/}

%\textbf{Implications}

%$\rightarrow$ Multi-dimensionality of urban systems and a link between form and function captured through multilayer percolation.

%$\rightarrow$ Possible transfer to policy-making recommandations: Pareto-optimal configuration can be used for the planning of regional transportation networks, policies for subsidies, etc.

%\textbf{Developments}

% detail method to extrapolate parameters of gravity flows

% interesting result : compare these emission with the synthetic ? (then just a refinment ?) -> yes, better. show new Pareto fronts with this fixed parameters for only emissions, and varying for economic ?

%$\rightarrow$ Systematic calibration of parameters to unveil more exhaustive Pareto fronts.

The use of calibration algorithms % TODO finish

Extrapolating transportation flows with a spatially explicit gravity and flow model can allow to compare these with actual transportation flows in the emission database, and yield a possible calibration. These extrapolated parameters could then be used within the economic and emissions potentials.

An other potential development would imply crossing our endogenous definitions of urban regions with socio-economic databases, and compute indicators implied in other dimensions of sustainability, for example related to socio-economic inequalities, spatial distribution of accessibilities, activities with different scaling exponents.
% ex : at least different scaling exponent.

\section{Conclusion}

%$\rightarrow$ Empirical and theoretical research directed towards concrete policy-making applications. \textbf{Need for more data-driven approaches.}
%First step towards systematic benchmarks and multi-modeling. \textbf{Need for more systematic model exploration.} % model coupling / benchmarking

%$\rightarrow$ Towards multi-scalar approaches ? \textbf{Need for more integrated models.}
 %Model integration and multi-scalarity ? \textbf{Need for more integrated models.}

%$\rightarrow$ Multidimensionality of urban systems ? \textbf{Need for more interdisciplinarity.}

%\textbf{Related works}

%Raimbault, J. (2018). Calibration of a density-based model of urban morphogenesis. PloS one, 13(9), e0203516.

%Raimbault, J. (2018). An Urban Morphogenesis Model Capturing Interactions between Networks and Territories. \textit{Forthcoming in Mathematics or Urban Morphogenesis.} arXiv:1805.05195.

%Raimbault, J. (2018). Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution des r{\'e}seaux de transport et des territoires (Doctoral dissertation, Université Paris 7 Denis Diderot). \url{https://halshs.archives-ouvertes.fr/tel-01857741}
%\textbf{Open repository} at \texttt{https://github.com/JusteRaimbault/UrbanMorphology} (code, data and results)%\\\smallskip
% no grid used
%\textbf{Acknowledgments}: thanks to the \textit{EGI} for access to the infrastructure.

\bibliographystyle{jimis-en}
\bibliography{biblio}

\appendix\footnotesize

\section{Appendix 1: supplementary sensitivity analyses}

%\sframe{Results: effective emissions}{

%\textit{Effective emissions exhibit a supralinear scaling of population}
%\includegraphics[height=0.83\textheight]{figures/aggreg_effective.png}

%\sframe{Results: all clusters Pareto fronts}{

%\textit{Variation of Pareto front patterns when potential parameter $\gamma,d_0$ vary.}

%\includegraphics[width=0.9\textwidth]{figures/full_paretos.png}

%\sframe{Results: an optimal morphology}{

%\textit{More monocentric areas are more optimal in terms of relative emissions and efficiency ?}

%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relefficiency.png}
%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relemissions.png}

% TODO : curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.

%\section{Acknowledgment}
%

%\section{Biography}
%Here, feel free to add short biographies of the authors.

\end{document}

%%%%%%%%
%% -- TEMPLATES

%\begin{table}
%  \newcolumntype{+}{>{\global\let\currentrowstyle\relax}}
%  \newcolumntype{^}{>{\currentrowstyle}}
%  \newcommand{\rowstyle}[1]{\gdef\currentrowstyle{#1}%
%    #1\ignorespaces
%  }
%  \centering
%  \begin{tabular}{+>{\bfseries}l^c^c^c^c}
%    \hline
%    \rowstyle{\bfseries}
%    & Sepal.Length & Sepal.Width & Petal.Length & Petal.Width\\
%    Setosa & 5.006 & 3.428 & 1.462 & 0.246\\
%    Versicolor & 5.936 & 2.77  & 4.26  & 1.326\\
%    Verginica & 6.588 & 2.974 & 5.552 & 2.026\\
%    \hline
%  \end{tabular}
%  \caption{Morbi malesuada diam at magna condimentum.}
%  \label{tab:example}
%\end{table}

%
%\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
%  {\includegraphics[width=11cm]{spider.jpg}}
%  \centering
%  \label{frog}
%  \caption{A spider, Picture from Didier Josselin.}
%  \end{figure*}
%

### 2019-03-16 Models/Percolation/percolation.R
added_diameter_and_size_for_fractal_dimension

@@ -9,18 +9,26 @@ source('percolationFunctions.R')
 # assumes data has been consolidated before
 load('../../Data/consolidated/indics.RData')
 
+#purpose='directSampling'
+purpose='test'
+
 # parameter values
-#quantiles = c(0.85,0.9,0.95)
-popquantiles=c(0.85,0.9,0.95)
-nwquantiles=c(0.0,0.8,0.9,0.95)
-radiuses=c(8000,10000,15000,20000,50000)
-#radiuses=c(8000)
-nwindics= c("ecount","mu","vcount","euclPerf")
-#nwindics= c("ecount")
-gammas=c(0.5,1,1.5,2.0)
-#gammas=c(1)
-decays=c(100,1000,10000,50000,100000)
-#decays=c(1000)
+#popquantiles=c(0.85,0.9,0.95)
+popquantiles=c(0.95)
+
+#nwquantiles=c(0.0,0.8,0.9,0.95)
+nwquantiles=c(0.95)
+
+#radiuses=c(8000,10000,15000,20000,50000)
+radiuses=c(8000,10000)
+
+#nwindics= c("ecount","mu","vcount","euclPerf")
+nwindics= c("ecount")
+
+#gammas=c(0.5,1,1.5,2.0)
+gammas=c(1)
+#decays=c(100,1000,10000,50000,100000)
+decays=c(1000)
 
 params=matrix(0,length(popquantiles)*length(nwquantiles)*length(radiuses)*length(nwindics)*length(gammas)*length(decays),6)
 i=1
@@ -33,14 +41,15 @@ colnames(params)=c("nwcol","popthq","nwthq","radius","gamma","decay")
 params$nwthq=as.numeric(as.character(params$nwthq));params$popthq=as.numeric(as.character(params$popthq));params$radius=as.numeric(as.character(params$radius));params$gamma=as.numeric(as.character(params$gamma));params$decay=as.numeric(as.character(params$decay))
 
 library(doParallel)
-cl <- makeCluster(60,outfile='log')
+#cl <- makeCluster(60,outfile='log')
+cl <- makeCluster(2,outfile='log')
+
 registerDoParallel(cl)
 res <- foreach(i=1:nrow(params)) %dopar% {
 #res=list()
 #for(i in 1:nrow(params)){
     #source('Models/Percolation/percolationFunctions.R')
   source('percolationFunctions.R')
[…]

### 2019-03-16 Models/Percolation/percolationFunctions.R

@@ -122,8 +122,13 @@ graphPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
   g = graph_from_adjacency_matrix(adjmat,mode='undirected')
   comps = components(g)
   k=1
+  # compute cluster diameters here as we have the graph
+  clustdiameters=c();clustsizesnodes=c();clustsizesedges=c()
   for(cnum in which(comps$csize >= minclustsize)){
     cpoints$cluster[comps$membership==cnum]=k
+    currentcomp = induced_subgraph(g,which(comps$membership==cnum))
+    clustdiameters=append(clustdiameters,diameter(currentcomp))
+    clustsizesnodes=append(clustsizesnodes,length(V(currentcomp)));clustsizesedges=append(clustsizesedges,length(E(currentcomp)))
     k=k+1
   }
   
@@ -141,7 +146,13 @@ graphPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
     }
   }
   
-  return(computeIndics(sppoints,gamma,decay))
+  indics = computeIndics(sppoints,gamma,decay)
+  
+  indics[["clustdiameters"]]=clustdiameters
+  indics[["clustsizesnodes"]]=clustsizesnodes
+  indics[["clustsizesedges"]]=clustsizesedges
+  
+  return(indics)
   
 }
 
@@ -184,7 +195,7 @@ computeIndics <- function(sppoints,gamma,decay,dmat=NULL,popcol="totalPop",emiss
   show(paste0('Computing indicators for ',nrow(sppoints),'points ; gamma = ',gamma,' ; decay = ',decay))
   #clusteredpoints=sppoints[!is.na(sppoints$cluster),]
   areas=c();pops=c();morans=c();avgdists=c();entropies=c();slopes=c();
-  efficiencies=c();emissions=c();totalemissions=c()
+  efficiencies=c();emissions=c();totalemissions=c();
   clustnums = unique(sppoints$cluster);clustnums=clustnums[!is.na(clustnums)]
   if(length(clustnums)>0){
   for(i in 1:length(clustnums)){

### 2019-03-16 Models/Percolation/percolation.R
test_ok

@@ -40,6 +40,8 @@ params=as.data.frame(params)
 colnames(params)=c("nwcol","popthq","nwthq","radius","gamma","decay")
 params$nwthq=as.numeric(as.character(params$nwthq));params$popthq=as.numeric(as.character(params$popthq));params$radius=as.numeric(as.character(params$radius));params$gamma=as.numeric(as.character(params$gamma));params$decay=as.numeric(as.character(params$decay))
 
+show(paste0('Number of parameters points = ',nrow(params)))
+
 library(doParallel)
 #cl <- makeCluster(60,outfile='log')
 cl <- makeCluster(2,outfile='log')

### 2019-03-16 Models/Percolation/percolation.R
test_bis

@@ -14,13 +14,13 @@ purpose='test'
 
 # parameter values
 #popquantiles=c(0.85,0.9,0.95)
-popquantiles=c(0.95)
+popquantiles=c(0.85)
 
 #nwquantiles=c(0.0,0.8,0.9,0.95)
-nwquantiles=c(0.95)
+nwquantiles=c(0.0)
 
 #radiuses=c(8000,10000,15000,20000,50000)
-radiuses=c(8000,10000)
+radiuses=c(15000,20000)
 
 #nwindics= c("ecount","mu","vcount","euclPerf")
 nwindics= c("ecount")

### 2019-03-16 Docs/Papers/Percolation/paper.tex
added_diameter_and_size_for_fractal_dimension
.

@@
 % List of k Editors
-\editors{First-name-$1$ Family-name-$1$, First-name-$2$ Family-name-$2$, First-name-$k$ Family-name-$k$...}
+\editors{}
 
@@
 
-\abstract{Territorial structure significantly conditions the sustainability of human settlements.
+\abstract{The spatial distribution of human settlements into territories strongly conditions their sustainability.
 Recent forms of urbanization, in particular integrated mega-city regions, may imply different patterns of economic and transportation flows and thus exhibit various performances regarding different indicators of sustainability.
@@
 
-% Morphologies of networks and territories
-% - intro : LGV ? Avignon : travaux Claire ! - truc avec les Papes ? 
-% https://www.persee.fr/doc/anami_0003-4398_1985_num_97_172_2228_t1_0446_0000_3
-% https://www.persee.fr/doc/mefr_0223-5110_1984_num_96_1_2762
-
-% -> visu OSM : see Metropolitan region ; plus positioning of Avignon ; Q of sustainibility ?
-
 % les formes des établissements humains peuvent se percevoir a differentes echelles et selon differentes dimensions.
@@
 % Cette presentation s'axera ainsi sur la morphologie des reseaux et des territories et sur leurs relations.
-%\includegraphics[width=1.2\textwidth]{figures/roadssouth.png}
-%\textit{Source: OpenStreetMap}
 
-The structure of road networks both translates its past growth dynamics and has a significant impact on the sustainability of territories it irrigates.
+The structure of road networks both contains its past growth dynamics and has a significant impact on the sustainability of territories it irrigates.
 Diverse methods to characterize the structure of spatial networks, and more particularly road networks, have been developed in that context.
 
-% here extracts from Claire's work
-% Lagesse, C., Bordin, P., & Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. -> nice picture Avignon
+\cite{lagesse2015spatial} % definition of the object "way"
 
@@
 
-%\textit{Multiple dimensions to characterize road networks}
-%\includegraphics[width=\textwidth,height=0.7\textheight]{figures/avignon_claire.png}
+A method to characterize topologies of these spatial networks is network percolation.
+Percolation in physics can be understood in a broad sense processes related to the progressive occupation or connection of nodes of a network, and is generally associated to a phase transition with the emergence of a giant cluster at a given connection probability \citep{callaway2000network}.
+Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling}.
+It is also a method applied to the analysis of street networks, for example to extract endogenous urban regions \citep{arcaute2016cities}.
+In spatial statistics, this method can be used to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.
 
-%Lagesse, C., Bordin, P., \& Douady, S. (2015). A spatial multi-scale object to analyze road networks. Network Science, 3(1), 156-181. \cite{lagesse2015spatial}
-
-A method to characterize topologies of these spatial networks is network percolation.
-Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling} and to the analysis of street networks for example to extract endogenous urban regions \citep{arcaute2016cities} or to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.
-
-%Network percolation
-% A method to characterize topologies of these spatial networks is network percolation.
-Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling} and to the analysis of street networks for example to extract endogenous urban regions \citep{arcaute2016cities} or to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.
+\cite{piovani2017urban} % meso scale percolation (within London area) and link with a retail location model
 
@@
 
+\cite{cottineau2018defining}
+
 % 
-
-% parmi ces outils, la percolation a ete applique recemment aux reseaux routiers.
-% en phyique, ce phenomene designe la traversee d'un milieu plus ou omins poreux par un fluide, et par exetnsion les processus pour l'attendre.
-En pariculier, nous l'entendrons au sens d'une occupation progressice des sites / connections des noeus d'un reseau, jusqu'a un etat critique ou l'ensemble de ceux-ci sont connectés (percolation threshold).
-[probailite de perco p_c]
-% en particulier ppour letude des viulles
-% des travaux relaticement anciens ont propose un phenomene de percolation locale correlation pour modeliser l'etalement urbain.
-% plus recemment, et plsu proches des travaux des physiciens, une determination endoganee des regions urbaines a etet proposee par l'equipe du acvs en appliquant une heuristqiue de percolation aux reseaux routiers.
 % ces methodes peuvent egalement servier a effecteur des statistqieus spatiales.
@@
 
-%\textbf{Network percolation: } \textit{progressive occupation/connection of nodes of a network} \cite{callaway2000network}
-
-%Application to the study of cities:
-
-%\begin{itemize}
-%	\item modeling urban growth \cite{makse1998modeling}
-%	\item endeogenous determination of regions \cite{arcaute2016cities}
-%	\item characterization of spatial point patterns \cite{huynh2018characterisation}
-%\end{itemize}
 %\textit{Towards complementary dimensions to condition road network percolation}
@@
 Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.
+However, urban systems are known to be multidimensional.
 
@@
 %\textbf{Research objective : } \textit{Investigate a multi-dimensional percolation of territorial networks taking into account urban morphology and road network topology; endogenous characterization of urban regions.}
-
-We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.
 
@@
 
+The percolation on such an abstract network is a necessary condition in our case to link the different dimensions considered, namely population distribution and local road network properties.
+
+We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.
+
 % Precisons a ppresent les objets geographiques a partir  desqueles nous allons construire le reseau.
@@
 
-%\textit{Territorial indicators computed for Europe by \cite{raimbault2018urban}}
-
-% here add maps to better visualize the spatial fields
-
-%\medskip
-
-%\includegraphics[width=0.49\textwidth]{figures/indics_morpho_discrquantiles.png}
-%\includegraphics[width=0.49\textwidth]{figures/indics_network_en_areasize100_offset50_factor0_5.png}
-
-%\textbf{Population distribution morphology} and \textbf{Network topology} (betweenness, closeness, clustering, efficiency, \ldots) computed on 50km spatial windows (Eurostat density grid and OpenStreetMap)
-
+%%%%%%%%%%%%%%%%
 \subsection{Sustainability indicators}
@@
 
+Using these potential flows follows the logic of \cite{arbabi2019development} which shows a need for improved intra-city-region mobility in England and Wales.
+Considering the regions as entities in which such transportation development policies can more easily been developed, we look at the sustainability of different possible regions if these potential flows were realized.
+Varying the parameters $\gamma$ and $d_0$ allows to control for the economic activity considered (high $\gamma$ values correspond to high added-value activities) and the span of interactions through $d_0$.
+
 \section{Results}
@@
 
-% banos2012towards
+\cite{banos2012towards} network efficiency
 
@@
 
-% note : euclperf should be minimized whereas other maximized : issue ?
-
-%\textbf{Implementation: } construction of a single layer spatial network given the condition on the two layers and distances, from the 5km resolution indicators spatial field; extraction of connected components.
-
-In practice, the analysis is implemented using R and the igraph package. The network is constructed by superposing the population density layer with the network layer.
+In practice, the analysis is implemented using R and the igraph package.
+The network is constructed by superposing the population density layer with the network layer, starting from the 5km resolution spatial fields for morphological and network indicators.
+This network is filtered with the threshold parameters for each layer and with the radius parameter.
+Connected components yield the clusters that we interpret as endogenous regions.
 
@@
 The experience plan is a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the network variable considered.
+We also make $\gamma$ and $d_0$ vary.
 
-%$\rightarrow$ 4800 parameter points
-%We systematically explore the clusters obtained for 840 parameter configurations.
-
-% + explain openmole -> later in potential developments
+% TODO recall def of euclPerf and why ok to be maximized
 
@@
 \subsection{Extracting endogenous mega-city regions}
+
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[ht] 
+%\resizebox{10cm}{7cm}
+  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}
+  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
+  \centering
+  \label{fig:exclusters}
+  \caption{Two examples of obtained clusters for different parameter values.}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
 
@@
 % TODO for the paper : check effect of adding network threshold : methodo info on multilayer percol in itself
-
-%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
-%\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
-  \centering
-  \label{fig:exclusters}
-  \caption{Two examples of obtained clusters for different parameter values.}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%
 
@@
 
-When considering the aggregated indicators for a parametrization of endogenous city regions, one can relate them to morphological indicators, computed by~\cite{}, that we average on areas.
+When considering the aggregated indicators for a parametrization of endogenous city regions, one can relate them to morphological indicators for population density spatial distribution, computed by~\cite{raimbault2018calibration}, that we average on areas.
 This establishes a link between urban morphology and sustainibility.
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_pc1-emissions.png}}
+  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_pc1-emissions_targeted.png}}
   \centering
-  \label{fig:emissions-pc1}
-  \caption{Aggregated values of normalized potential emissions, as a function of the first morphological principal component (PC1), for varying values of parameters $d_G$ and $\gamma_G$.
-As PC1 is mainly linked to monocentricity, there seems to exist an optimal intermediate level of monocentricity for emissions alone.}
+  \caption{Aggregated values of normalized potential emissions, as a function of the first morphological principal component (PC1), for varying values of parameters $d_G$ (rows) and $\gamma_G$ (columns).
+Other intermediate values for these parameters yield similar behaviors.
+As PC1 is mainly linked to monocentricity, there seems to exist an optimal intermediate level of monocentricity for emissions alone.\label{fig:emissions-pc1}}
 \end{figure*}
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale.png}}
+  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
   \centering  
-  \label{fig:paretos-relative}
   \caption{Relative potential emissions against relative potential economic unefficiency (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows).
-Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.}
+Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
 \end{figure*}
@@
 
-\appendix\footnotesize
+%\appendix\footnotesize
 
-\section{Appendix 1: supplementary sensitivity analyses}
+%\section{Appendix 1: supplementary sensitivity analyses}
 

### 2019-03-16 Models/Percolation/percolation.R

@@ -9,26 +9,26 @@ source('percolationFunctions.R')
 # assumes data has been consolidated before
 load('../../Data/consolidated/indics.RData')
 
-#purpose='directSampling'
-purpose='test'
+purpose='directSampling'
+#purpose='test'
 
 # parameter values
-#popquantiles=c(0.85,0.9,0.95)
-popquantiles=c(0.95)
+popquantiles=c(0.85,0.9,0.95)
+#popquantiles=c(0.85)
 
-#nwquantiles=c(0.0,0.8,0.9,0.95)
-nwquantiles=c(0.95)
+nwquantiles=c(0.0,0.8,0.9,0.95)
+#nwquantiles=c(0.0)
 
-#radiuses=c(8000,10000,15000,20000,50000)
-radiuses=c(8000,10000)
+radiuses=c(8000,10000,15000,20000,50000)
+#radiuses=c(20000,50000)
 
-#nwindics= c("ecount","mu","vcount","euclPerf")
-nwindics= c("ecount")
+nwindics= c("ecount","mu","vcount","euclPerf")
+#nwindics= c("ecount")
 
-#gammas=c(0.5,1,1.5,2.0)
-gammas=c(1)
-#decays=c(100,1000,10000,50000,100000)
-decays=c(1000)
+gammas=c(0.5,1,1.5,2.0)
+#gammas=c(1)
+decays=c(100,1000,10000,50000,100000)
+#decays=c(1000)
 
 params=matrix(0,length(popquantiles)*length(nwquantiles)*length(radiuses)*length(nwindics)*length(gammas)*length(decays),6)
 i=1
@@ -65,7 +65,10 @@ res <- foreach(i=1:nrow(params)) %dopar% {
 }
 stopCluster(cl)
 
-save(res,file=paste0('res/',purpose,'_',format(Sys.time(), "%Y%m%d_%H%M%S"),'.RData'))
+tstamp=format(Sys.time(), "%Y%m%d_%H%M%S")
+save(res,file=paste0('res/',purpose,'_',tstamp,'.RData'))
+save(params,file=paste0('res/',purpose,'_params_',tstamp,'.RData'))
+
 
 #############
 

### 2019-03-16 Models/Percolation/percolationFunctions.R

@@ -151,6 +151,7 @@ graphPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
   indics[["clustdiameters"]]=clustdiameters
   indics[["clustsizesnodes"]]=clustsizesnodes
   indics[["clustsizesedges"]]=clustsizesedges
+  show(names(indics))
   
   return(indics)
   

### 2019-03-16 Models/Percolation/percolationResults.R

@@ -1,9 +1,18 @@
 
+setwd(paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Models/Percolation'))
+
+# indics
+load('../../Data/consolidated/indics.RData')
+
 # load results from percolation
 
-resstamp = 'directSampling_20181016_024956'
-load(paste0('res/',resstamp,'.RData'))
-resdir = paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/',resstamp,'/');dir.create(resdir)
+resstamp = '20181016_024956'
+purpose='directSampling'
+#resstamp = 'test_20190316_154614'
+load(paste0('res/',purpose,'_',resstamp,'.RData'))
+load(paste0('res/',purpose,'_params_',resstamp,'.RData'))
+resdir = paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/',purpose,'_',resstamp,'/');dir.create(resdir)
+
 
 source(paste0(Sys.getenv('CS_HOME'),'/Organisation/Models/Utils/R/plots.R'))
 
@@ -102,7 +111,9 @@ for(gamma in gammas){for (decay in decays){
 
 
 g=ggplot(sres,aes(x=emissions,y=efficiency,col=pop/totalPop))
-g+geom_point()+facet_grid(decay~gamma)+xlab("Normalized potential emissions")+ylab("1 - Normalized potential efficiency")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
+g+geom_point()+facet_grid(decay~gamma)+
+  xlab("Normalized potential emissions")+ylab("1 - Normalized potential efficiency")+
+  scale_color_continuous(name="Relative\nPopulation")+stdtheme
 ggsave(file=paste0(resdir,'aggreg_paretos.png'),width=30,height=25,units='cm')
 
 g=ggplot(sres,aes(x=emissions*totalPop/pop,y=efficiency*totalPop/pop,col=pop/totalPop))
@@ -190,6 +201,13 @@ g=ggplot(sres,aes(x=PC1,y=emissions,col=pop/totpop))
 g+geom_point()+facet_grid(decay~gamma)+geom_smooth(method='loess',n=50)+xlab("PC1")+ylab("Normalized potential emissions")+scale_color_continuous(name="Relative\nPopulation")+stdtheme
[…]

### 2019-03-16 Docs/Papers/Percolation/paper.tex

@@
 The first component relates to a level of monocentricity ($PC1 = -0.3\cdot I + 0.54 \cdot \bar{d} + 0.51\cdot \varepsilon + 0.59 \cdot \alpha$).
-As shown in Fig.~\ref{fig:emissions-pc1}, there seems to exist an optimal intermediate level of monocentricity regarding emissions only, except for long-range and low-hierarchy interactions.
-However, when considering both emissions and economic indicators, urban form then acts as a compromise variable, moving points on Pareto fronts, as shown in Fig.~\ref{fig:paretos-relative}.
-In some case, highly monocentric areas can be a good compromise, whereas the intermediate optimal for emissions may yield highly inefficient areas.
 
@@
 
-%Morphological trade-offs
-
-%\textit{No ``optimal'' cities, different forms yield different compromises in terms of relative indicators.}
+As shown in Fig.~\ref{fig:emissions-pc1}, there seems to exist an optimal intermediate level of monocentricity regarding the normalized indicator for emissions only, except for long-range and low-hierarchy interactions.
 
@@
 %%%%%%%%%%%%%%%%%%%%
+
+However, when considering both emissions and economic indicators, urban form then acts as a compromise variable, moving points on Pareto fronts, as shown in Fig.~\ref{fig:paretos-relative}.
+In some case, highly monocentric areas can be a good compromise, whereas the intermediate optimal for emissions may yield highly inefficient areas.
+This unveils morphological trad-offs, confirming that there is no optimal urban form, but different compromises regarding the conflicting indicators.
 

### 2019-03-16 Models/Percolation/percolationResults.R
percol_trans

@@ -6,8 +6,11 @@ load('../../Data/consolidated/indics.RData')
 
 # load results from percolation
 
-resstamp = '20181016_024956'
-purpose='directSampling'
+#resstamp = '20181016_024956'
+#purpose='directSampling'
+resstamp = '20190316_182550'
+purpose='directSampling_radius'
+
 #resstamp = 'test_20190316_154614'
 load(paste0('res/',purpose,'_',resstamp,'.RData'))
 load(paste0('res/',purpose,'_params_',resstamp,'.RData'))
@@ -23,17 +26,42 @@ totalPop = sum(indics$totalPop)
 ## full areas stats
 
 for(k in 1:length(res)){res[[k]]$paramrow=rep(k,length(res[[k]]$pops))}
-full = data.frame(matrix(unlist(lapply(res,function(l){l$areas=NULL;unlist(data.frame(t(data.frame(l))))})),byrow=T,ncol=length(res[[1]])-1))
-colnames(full)<-names(res[[1]])[-1]
+full = data.frame(matrix(unlist(lapply(res,function(l){l$areas=NULL;l$clustdiameters=NULL;l$clustsizesnodes=NULL;l$clustsizesedges=NULL;unlist(data.frame(t(data.frame(l))))})),byrow=T,ncol=length(res[[1]])-4))
+colnames(full)<-names(res[[1]])[c(-1,-10,-11,-12)]
 params$paramrow=1:nrow(params)
 full=left_join(full,params)
 
+# compute relative indicators
 full$relemissions=rep(nrow(full),0)
 full$relefficiency=rep(nrow(full),0)
 for(k in unique(full$paramrow)){
   full$relemissions[full$paramrow==k]=(max(full$emissions[full$paramrow==k])-full$emissions[full$paramrow==k])/(max(full$emissions[full$paramrow==k])-min(full$emissions[full$paramrow==k]))
   full$relefficiency[full$paramrow==k]= 1 - (max(full$efficiencies[full$paramrow==k])-full$efficiencies[full$paramrow==k])/(max(full$efficiencies[full$paramrow==k])-min(full$efficiencies[full$paramrow==k])) 
 }
+
+
+#####
+# percolation transition
+
+toremove=c("pops","morans","avgdists","entropies","slopes","efficiencies","emissions","totalemissions")
[…]

### 2019-03-17 Models/Percolation/percolation.R
radius_manymore_sampling

@@ -17,14 +17,14 @@ purpose='directSampling_radius'
 popquantiles=c(0.85)
 
 #nwquantiles=c(0.0,0.8,0.9,0.95)
-nwquantiles=c(0.0,0.8)
+nwquantiles=c(0.0,0.8,0.95)
 
 #radiuses=c(8000,10000,15000,20000,50000)
-radiuses=seq(8000,50000,by=1000)
+radiuses=seq(8000,100000,by=1000)
 #radiuses=c(20000,50000)
 
-#nwindics= c("ecount","mu","vcount","euclPerf")
-nwindics= c("vcount","ecount")
+nwindics= c("ecount","mu","vcount","euclPerf")
+#nwindics= c("vcount","ecount")
 
 #gammas=c(0.5,1,1.5,2.0)
 gammas=c(1)

### 2019-03-17 Docs/Papers/Percolation/paper.tex

@@
 A method to characterize topologies of these spatial networks is network percolation.
-Percolation in physics can be understood in a broad sense processes related to the progressive occupation or connection of nodes of a network, and is generally associated to a phase transition with the emergence of a giant cluster at a given connection probability \citep{callaway2000network}.
-Such approaches have been applied to the modeling of urban growth \citep{makse1998modeling}.
+Percolation in physics can be understood in a broad sense processes related to the progressive occupation or connection of nodes of a network, and is generally associated to a phase transition with the emergence of a giant cluster at a given connection probability \citep{stauffer2014introduction}.
+Important applications include the quantification of network robustness \citep{callaway2000network}.
+
+Such approaches have been applied to urban systems.
+\cite{makse1998modeling} model urban growth with a local percolation model.
 It is also a method applied to the analysis of street networks, for example to extract endogenous urban regions \citep{arcaute2016cities}.
@@
 % from monodimensional to multidimensional
-
-% similar to 
-%Cottineau, C., Finance, O., Hatna, E., Arcaute, E., & Batty, M.
-(2018).
-Defining urban clusters to detect agglomeration economies.
-Environment and Planning B: Urban Analytics and City Science, 2399808318755146.
-
-\cite{cottineau2018defining}
-
-% 
-% ces methodes peuvent egalement servier a effecteur des statistqieus spatiales.
-% nous porposons ici d;'etendre ces methodes en prenant en compte differentes dimensions urbaines, de la meme maniere que cottineau et al combine densite de population et flux de commuting pour definir les regoins urbaines.
 
@@
 
-This paper addresses such a gap by introducing a multi-dimensional percolation heuristic.
+This paper addresses such a gap by introducing a multi-dimensional percolation heuristic, in order to combine different dimensions of the urban system, the same way that \cite{cottineau2018defining} combines population density and commuting flows to produce multiple definitions of urban areas.
 
-%Multidimensional percolation
-
-% Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.
+The rationale behind combining urban form with network topology measures relies on the capture of the link between urban form and function, assumed as distributed by transportation networks \citep{raimbault2018caracterisation}.
 
@@
 
-%$\rightarrow$ Need to combine morphological and functional dimensions of cities \cite{burger2012form}
-
-%$\rightarrow$ Interactions between networks and territories to capture the link between form and function \cite{raimbault2018caracterisation}; potential application to sustainability of urban systems
-
-% research question
-
 %\textbf{Research objective : } \textit{Investigate a multi-dimensional percolation of territorial networks taking into account urban morphology and road network topology; endogenous characterization of urban regions.}
 
+Our contribution relies on several points: (i) this is to the best of our knowledge the first time a multi-dimensional percolation method is appplied to urban systems; (ii) we furthermore apply it on a significant spatial extent; and (iii) we link the cluster obtained with simple sustainability measures.
+
+The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied, and the indicators used to characterized the sustainability of clusters produced.
+We then describe the results of applying this method to population and network variables for the whole European Union, focusing on the endogenous regions produced and their sustainability properties.
+We finally discuss possible developments and the implications of this methodology to the design of policies.
+
+%%%%%%%%%%%%%%%%%
 \section{Methods}
+%%%%%%%%%%%%%%%%%
 
+%%%%%%%%%%%%%%%%%
 \subsection{Multi-dimensional percolation}
-
-Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold.
-This heuristic is similar to multilayer network percolation \cite{boccaletti2014structure}.
-
-% decrivons a presons de neniere stylisee l'algorithme utilise.
-% le processus de peroclation initrial est deterministe, base sur la distance, met est prochess des random eulidian networks evoque hier par Alain Franc : si deux neouds sont a une distance d'un rayon critique r_0, ceux ci sont connectes par un lien.
-% nous ajoutons alors sur chaque couche du reaseu un seul de peroclation qui determine si le neod peut etre prus en compte dqns le reseau de la couhce.
-% par conjonction des contraintes, nous obtenus un reseau a unique couche dans laquelle simplement els composantes connexes donnent les clusters perfcolés.
 
@@
 
-%\textbf{Parameters: } percolation radius $r_0$, percolation thresholds $\theta_i$ for each layer
+Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold.
+This heuristic is similar to multilayer network percolation \citep{boccaletti2014structure}.
+
+\cite{son2012percolation} % extension of epidemic spreading
+
+\cite{hackett2016bond} bond percolation multiplex networks
+
+% decrivons a presons de neniere stylisee l'algorithme utilise.
+% le processus de peroclation initrial est deterministe, base sur la distance, met est prochess des random eulidian networks evoque hier par Alain Franc : si deux neouds sont a une distance d'un rayon critique r_0, ceux ci sont connectes par un lien.
+% nous ajoutons alors sur chaque couche du reaseu un seul de peroclation qui determine si le neod peut etre prus en compte dqns le reseau de la couhce.
+% par conjonction des contraintes, nous obtenus un reseau a unique couche dans laquelle simplement els composantes connexes donnent les clusters perfcolés.
+
+The method works with an arbitrary number of layers.
+
+The rationale
+
+The parameters implied in this heuristic are the percolation radius $r_0$ and the percolation thresholds $\theta_i$ for each layer $i$.
 
@@
 
+The corresponding relative indicators are defined as
+% TODO relative indiactors / normalized (by pop ?)
+
 Using these potential flows follows the logic of \cite{arbabi2019development} which shows a need for improved intra-city-region mobility in England and Wales.
@@
 In practice, the analysis is implemented using R and the igraph package.
+Source code, data and results are available on the open git repository of the project at \url{https://github.com/JusteRaimbault/UrbanMorphology}.
+
 The network is constructed by superposing the population density layer with the network layer, starting from the 5km resolution spatial fields for morphological and network indicators.
@@
 
-%\textbf{Experience plan: } grid sampling for $r_0,\theta_P,\theta_N$ and network variables; additional gravity potential parameters $\gamma,d_0$ (detailed after)
-
 The experience plan is a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the network variable considered.
@@
 
-% TODO recall def of euclPerf and why ok to be maximized
+We recall that the euclidian performance of the network is in our case $<d_e/d_n>$ where the average is taken on all origin-destination pairs in the network, $d_e$ is the euclidian distance and $d_n$ the network distance.
+Thus, it indeed increases with network performance, in consistence with the use done here through thresholding.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%
 \subsection{Extracting endogenous mega-city regions}
@@
 %\resizebox{10cm}{7cm}
+{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}\\
   {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}
@@
   \centering
-  \label{fig:exclusters}
-  \caption{Two examples of obtained clusters for different parameter values.}
+
+  \caption{Examples of obtained clusters for different parameter values.
+In the third case for example, we obtain the urban regions of West midlands and London in the UK, Randstad merged with Rhein-Rhur and Rhein-Main in Germany, Paris in France, also with capital cities such as Copenhaguen, Stockholm and Helsinki.
+There is no cluster in South Europe in that case, due to the high population density threshold.\label{fig:exclusters}}
 \end{figure*}
@@
 
-Maps reveal that most configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}.
-These are here defined endogenously from the bottom-up.
+We obtain different endogenous morphologies.
 
-% les rsultats sont particulierement interessantm puiqaue ceratines valuers de apremtres menest a des configurations connues par lkes geographesm celles des mega-cities regions
-% [rq : affirnation arbitraire iuci car pas evidence-based, il faudrait comp systematique etc , le probeleme vcest que meme pas de liste avec consensus sur l'obejct, comme soubent en geo>...]
-% du coup "visuellemtn" randstad, rehinrhuhrm, rehinmainm London, Paris, Manchester Leeds,
+Maps reveal that some configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}.
+These are here defined endogenously from the bottom-up and have a priori no reason to coincide with these functional regions.
 
-% note : forgot r_0 in figs
-%\includegraphics[width=\textwidth,height=0.9\textheight]{figures/totalPop4183694_00056402_ecount850_radius8000.png}
+We show some examples in Fig.~\ref{fig:exclusters}.
 
-%Different endogenous morphologies
+%%%%%%%%%%%%%%%%%%%
+\subsection{Percolation transition and fractal dimension}
 
-% TODO for the paper : check effect of adding network threshold : methodo info on multilayer percol in itself
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[ht] 
+%\resizebox{10cm}{7cm}
+  %{\includegraphics[width=\linewidth]{figures/}}
+  \centering
+  \label{fig:percolation}
+  \caption{Percolation transition.}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
 
+In its application to road networks by \cite{arcaute2016cities}, the structure of the national urban system for UK is captured by studying the percolation transition, i.e.
+the variation of the size of the largest cluster as a function of the percolation radius.
+As this signature is tightly linked to historical, cultural and geographical conditions, the application to different urban systems should yield different results.
+We study here this property, for different threshold parameter values.
+The relative size of the largest cluster is plotted in Fig.~\ref{fig:percolation} as a function of the percolation radius.
+
+This aspect furthermore gives methodological information on multilayer percolation. Indeed, comparing the result with $\theta_N = 0$ (single layer percolation) with $\theta_N = 0.8$
+
+%%%%%%%%%%%%%%%%%%%
 \subsection{Pareto fronts for sustainability}
 
-We show therein that different population, network and distance thresholds yield different performances in terms of sustainability, exhibiting a Pareto front.
-This suggests policies in terms of regional integration to increase the sustainability of mega-city regions.
+We show therein that different population, network and distance thresholds yield different performances in terms of sustainability.
 
@@
 
+%%%%%%%%%%%%%%%%%
 \section{Discussion}
 
-Further work can consist in the use of calibration heuristics \citep{reuillon2013openmole} to find in a more robust way optimal parameter values.
+%%%%%%%%%%%%%%%%%
+\subsection{Developments}
+
+Further work can consist in the use of calibration heuristics to find in a more robust way optimal parameter values.
+The OpenMOLE model exploration platform provides a transparent access to genetic algorithms for multi-objective optimization \citep{reuillon2013openmole}.
+The use of such calibration algorithms would allow to unveil the effective form of Pareto fronts, that we may have missed here through the grid sampling.
 
@@
 
-%$\rightarrow$ towards the use of genetic algorithms on grid, made smooth with the OpenMOLE software \url{https://next.openmole.org/}
-%\includegraphics[height=0.35\textheight]{figures/openmole.png}
-%\footnotesize \textit{OpenMOLE: (i) embed any model as a black box; (ii) transparent access to main High Performance Computing environments; (iii) model exploration and calibration methods.}
-%\textbf{Apply to the summer school !} \url{https://exmodelo.org/}
-
 %\textbf{Implications}
 
-%$\rightarrow$ Multi-dimensionality of urban systems and a link between form and function captured through multilayer percolation.
-
-%$\rightarrow$ Possible transfer to policy-making recommandations: Pareto-optimal configuration can be used for the planning of regional transportation networks, policies for subsidies, etc.
-
-%\textbf{Developments}
+%$\rightarrow$
 
@@
 
-%$\rightarrow$ Systematic calibration of parameters to unveil more exhaustive Pareto fronts.
-
-The use of calibration algorithms % TODO finish
-
 Extrapolating transportation flows with a spatially explicit gravity and flow model can allow to compare these with actual transportation flows in the emission database, and yield a possible calibration.
@@
 
+%%%%%%%%%%%%%%%%%
+\subsection{Towards policy applications}
+
+This suggests policies in terms of regional integration to increase the sustainability of mega-city regions.
+
+The way such results can be transferred to policy-making recommandations remains an open question, but Pareto-optimal configurations can be used for the planning of regional transportation networks for example, or to design policies for the distribution of subsidies.
+
+%%%%%%%%%%%%%%%%%
 \section{Conclusion}
+
+In conclusion, our multilayer percolation approache captures in a way the multi-dimensionality of urban systems and a link between form and function.
 
@@
 %$\rightarrow$ Multidimensionality of urban systems ? \textbf{Need for more interdisciplinarity.}
-
-%\textbf{Related works}
-
-%Raimbault, J. (2018). Calibration of a density-based model of urban morphogenesis. PloS one, 13(9), e0203516.
-
-%Raimbault, J. (2018). An Urban Morphogenesis Model Capturing Interactions between Networks and Territories. \textit{Forthcoming in Mathematics or Urban Morphogenesis.} arXiv:1805.05195.
-
-%Raimbault, J.
-(2018).
-Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution des r{\'e}seaux de transport et des territoires (Doctoral dissertation, Université Paris 7 Denis Diderot).
-\url{https://halshs.archives-ouvertes.fr/tel-01857741}
-%\textbf{Open repository} at \texttt{https://github.com/JusteRaimbault/UrbanMorphology} (code, data and results)%\\\smallskip
-% no grid used
-%\textbf{Acknowledgments}: thanks to the \textit{EGI} for access to the infrastructure.
 

### 2019-03-17 Docs/Papers/Percolation/paper.tex
all_results

@@
 % header
-\title{Identifying European Sustainable Mega-city Regions with Multi-dimensional Network Percolation}
+\title{Multi-dimensional Urban Network Percolation}
+% Identifying European Sustainable Mega-city Regions with Multi-dimensional Network Percolation
 % Endogenous urban regions as an optimal compromise between economic integration and greenhouse gases emissions
@@
 
-\abstract{The spatial distribution of human settlements into territories strongly conditions their sustainability.
-Recent forms of urbanization, in particular integrated mega-city regions, may imply different patterns of economic and transportation flows and thus exhibit various performances regarding different indicators of sustainability.
-This paper proposes to reconstruct endogenous urban regions from the bottom-up using network percolation, in a multi-dimensional way to take into account both urban form (spatial distribution of population) and urban function (properties of transportation networks).
-Variable parametrizations allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions).
+\abstract{Network percolation has recently been proposed as a method to characterize the global structure of an urban system form the bottom-up.
+This paper proposes to extend urban network percolation in a multi-dimensional way, to take into account both urban form (spatial distribution of population) and urban functions (here as properties of transportation networks).
+The method is applied to the European urban system to reconstruct endogenous urban regions.
+The variable parametrization allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions).
 This suggests a customizable spatial design of policies to develop sustainable territories.}
 
-\keywords{Road network; multilayer percolation; mega-city region}
+\keywords{Road network; Multi-dimensional percolation; European urban system; Mega-city region}
 
@@
 
-% les formes des établissements humains peuvent se percevoir a differentes echelles et selon differentes dimensions.
-Elle font partie de notre vie courante, a ces mutiples echelles, et pourtant nous sont si etrangeres.
-Afin d'introduire notre sujet, je propose ici une image familière pour bien d'entre vous qui fréquentent cette région.
-% La metropole etendue de la region marseilleise ne se limite pas a l'aspect administratif d'aix marseille, mais englobe bien avignon et la ciotat, voire manosque, des lors que l'on prend un certain recul geographique.
-% il n'aura pas manque aux eoils attentifs que ce recul est d'autant plus une abstraction, puisque nous ne represnetons ici uniquement les reseaux routiers, qui s'averent deja etre un proxy suffisant pour estrapoler une densite d'activités humaines.
-% Cette presentation s'axera ainsi sur la morphologie des reseaux et des territories et sur leurs relations.
+The structure of road networks can be used as a proxy to understand its past growth dynamics, but also has a significant impact on the future sustainability of territories it irrigates.
+Diverse methods to characterize the structure of spatial networks, and more particularly road networks, have been developed in that context, including classical network indicators such as centralities \citep{crucitti2006centrality} but also more elaborated constructions capturing more realistic processes in terms of street network use \citep{lagesse2015spatial}.
+These study are by essence interdisciplinary, or at least imply complementary viewpoints from disciplines as diverse as architecture with space syntax \citep{hillier1976space}, physics with the study of spatial networks \citep{barthelemy2011spatial}, or social science disciplines concerned with space such as geography \citep{ducruet2014spatial}.
 
-The structure of road networks both contains its past growth dynamics and has a significant impact on the sustainability of territories it irrigates.
-Diverse methods to characterize the structure of spatial networks, and more particularly road networks, have been developed in that context.
-
-\cite{lagesse2015spatial} % definition of the object "way"
-
-% La caracterisation des reseaux routiers, par exemples en termes de classification et dindicateurs topoliques, est liee a differents enjeux qui incluent la comprehnsion des dynamiques ppasses par la contingence de ces artefacts, amis qussi des questions futues liees a la soutenabilité des territoires qui'ils irriguent.
-% les developpements recents, issus d'efforts interdispclinaires, incluent par exmeple ce travail de Claire Lagesse applique a la ville que v ous reconnaissaez bien, pour la definition de nouveau object a la pertineence thematique plus grande que les seuls entites des bases de donnes sans ontologie propre.
-(ex perso ballade avignon hier : vraie coherence thematique de l'objet "Way")
-
-A method to characterize topologies of these spatial networks is network percolation.
-Percolation in physics can be understood in a broad sense processes related to the progressive occupation or connection of nodes of a network, and is generally associated to a phase transition with the emergence of a giant cluster at a given connection probability \citep{stauffer2014introduction}.
-Important applications include the quantification of network robustness \citep{callaway2000network}.
-
-Such approaches have been applied to urban systems.
-\cite{makse1998modeling} model urban growth with a local percolation model.
-It is also a method applied to the analysis of street networks, for example to extract endogenous urban regions \citep{arcaute2016cities}.
+A method to characterize topologies of such urban spatial networks is network percolation, initially applied to road networks by \cite{arcaute2016cities}.
+Percolation in physics can be understood in a broad sense as processes related to the progressive occupation or connection of nodes of a network, and is generally associated to a phase transition with the emergence of a giant cluster at a given connection probability \citep{stauffer2014introduction}.
+Important applications include the quantification of network robustness \citep{callaway2000network} or the modeling of epidemic spreading \citep{newman1999scaling}.
+Such approaches have been applied to urban systems not only for the study of networks.
+\cite{makse1998modeling} model urban growth with a local percolation model for site occupancy.
+\cite{arcaute2016cities} focus on the analysis of street networks and extract endogenous urban regions for UK which correlate with socio-economic properties, and provide a definition of urban areas which highly correlates with land-cover data.
+\cite{piovani2017urban} apply road network percolation at the mesoscopic scale of London metropolitan area, in relation with a retail location model.
 In spatial statistics, this method can be used to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.
 
-\cite{piovani2017urban} % meso scale percolation (within London area) and link with a retail location model
+Existing heuristics however generally focus on a single dimension or property of the urban system.
+However, such systems are known to be multidimensional, for example through the superposition of the morphological dimension of networks, and the functional properties of the urban environment \citep{burger2012form}.
+The link between urban form and function remains in particular an open question \citep{batty1994fractal}, but more generally the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \cite{perez2016agent}.
+This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic.
+The method allows to combine different dimensions of the urban system, the same way that \cite{cottineau2018defining} combines population density and commuting flows to produce multiple definitions of urban areas.
 
-% from monodimensional to multidimensional
+Beside these methodological issues, applied tools are needed to quantify the sustainability of recently emerged urban forms.
+In particular, according to \cite{lenechet2017peupler}, the most recent transition of human settlement systems (in the sense of \cite{sanders2017peupler}, i.e.
+a change in the dynamical regime ruling the evolution of the spatial structure of settlements) is the emergence of mega-city regions.
+These have been defined by \cite{hall2006polycentric} as polycentric urban structures highly integrated in terms of flows.
+The transition imply complex processes such as changes in the governance structure, and can not be associated to the stylized transition identified by \cite{louf2013modeling} in a simplistic toy urban model, and therefore has chances to imply more drivers than negative externalities of congestion only.
+To what extent these new urban forms are sustainable, for example in the broad sense of UN development goals \citep{komiyama2006sustainability}, remains an open question.
+In order to test our multi-dimensional percolation method, we propose to apply it to the endogenous characterization of urban regions, and compute stylized sustainability indicators on the constructed regions.
 
-%\textit{Towards complementary dimensions to condition road network percolation}
-%$\rightarrow$ similar to \cite{cottineau2018defining} to define urban areas
-
-% necessity of multi-dimensional percolation
-
-Existing heuristics however generally focus on a single morphological dimension of networks, and leave out the functional properties of urban systems \citep{burger2012form}.
-However, urban systems are known to be multidimensional.
-
-This paper addresses such a gap by introducing a multi-dimensional percolation heuristic, in order to combine different dimensions of the urban system, the same way that \cite{cottineau2018defining} combines population density and commuting flows to produce multiple definitions of urban areas.
-
-The rationale behind combining urban form with network topology measures relies on the capture of the link between urban form and function, assumed as distributed by transportation networks \citep{raimbault2018caracterisation}.
-
-% - why the relation between form and function, aka pop and network in our view is crucial
-% - take into account in percolation
-% - potential application : endogenous sustainable urban entities
-
-% notre approceh reponde a la question ouverte du lien entre ofrme et fonction da`ns les systemes urbaine ;  et par ailleurs se base sur les interatcions entre reseaux te territoires poour fournire un proy de ce lien.
-% nous propospns ainsi d'explorer une heuristique multi-dimensionnelle de la percolation qui prend en compte morplholohie urbaine et yopologie dy reseau routier ; nous l'appliquon pour caracterisere endogenement les regions urbaines.
-% [rq : deux niveau de reseuax dans notree apporche : (mais en. fait comme souvent en geo ; ) le physique, et le abstraot que l'on construite a partir d'entites spatiales.
-
-%\textbf{Research objective : } \textit{Investigate a multi-dimensional percolation of territorial networks taking into account urban morphology and road network topology; endogenous characterization of urban regions.}
-
-Our contribution relies on several points: (i) this is to the best of our knowledge the first time a multi-dimensional percolation method is appplied to urban systems; (ii) we furthermore apply it on a significant spatial extent; and (iii) we link the cluster obtained with simple sustainability measures.
-
+Our contribution relies on several points: (i) this is to the best of our knowledge the first time a multi-dimensional percolation method is applied to urban systems; (ii) we furthermore apply it on the significant spatial extent of all European Union; and (iii) we link the clusters obtained with simple sustainability measures.
 The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied, and the indicators used to characterized the sustainability of clusters produced.
@@
   \centering
-  \label{fig:method}
-  \caption{Schematic representation of the multi-dimensional network percolation heuristic.}
+  \caption{Schematic representation of the multi-dimensional network percolation heuristic.
+We show a stylized configuration with two layers having the same nodes, and the links within each layer are created following the percolation radius $r_0$ and the thresholds for the layer variables.
+The final clusters are the superposition of these.
+The green points give examples of starting points in the case of a propagation heuristic.\label{fig:method}}
 \end{figure*}
@@
 
-Given discrete spatial fields, site percolation is operated between two cells given a threshold parameter for each dimension and a distance threshold.
-This heuristic is similar to multilayer network percolation \citep{boccaletti2014structure}.
+Percolation processes in multilayer networks have been proposed as an extension within simple networks \citep{boccaletti2014structure}.
+A generalization of epidemic spreading can for example be achieved using this framework \citep{son2012percolation}.
+In the case of multilayer networks sharing the same nodes for all layers, often called multiplex networks, bond percolation has also been studied \citep{hackett2016bond}.
 
-\cite{son2012percolation} % extension of epidemic spreading
+In the case of our heuristic, bond percolation is operated between two cells given a distance threshold, and furthermore with a threshold parameter for each layer assuming a node function within each layer.
+The distance-based connection is similar to generative processes for random euclidian networks \citep{penrose1999k}.
 
-\cite{hackett2016bond} bond percolation multiplex networks
+More formally, let assume a set of nodes $V = v_i$ common to all layers, and layers edges $E_j$ taken as empty at the initial state of the algorithm.
+Each node has a value of the considered variables associated to each layer, written $v_{ij}$.
+For each layer, a link $e_{kl} \in E_j$ is created if $d(v_k,v_l) < r_0$ where $d$ is the distance between the nodes (which can be any distance) and $r_0$ the percolation radius, and if $v_{kj} > \theta_j$ and $v_{lj} > \theta_j$ where $\theta_j$ is the threshold for layer $j$.
+The final percolated network edges $E$ is composed by links contained within all layers simultaneously.
+The multi-dimensional percolation clusters are then the connected components of this network $(V,E)$.
+The parameters implied in this heuristic are the percolation radius $r_0$ and the percolation thresholds $\theta_j$ for each layer $j$, allowing a flexible application through parametrization.
 
-% decrivons a presons de neniere stylisee l'algorithme utilise.
-% le processus de peroclation initrial est deterministe, base sur la distance, met est prochess des random eulidian networks evoque hier par Alain Franc : si deux neouds sont a une distance d'un rayon critique r_0, ceux ci sont connectes par un lien.
-% nous ajoutons alors sur chaque couche du reaseu un seul de peroclation qui determine si le neod peut etre prus en compte dqns le reseau de la couhce.
-% par conjonction des contraintes, nous obtenus un reseau a unique couche dans laquelle simplement els composantes connexes donnent les clusters perfcolés.
-
+Note that we do not call our method ``multi-layer percolation'', as it is not strictly multi-layer since nodes are common.
+The term of multi-dimensional percolation is more suited to the use of multiple variables and thresholds.
 The method works with an arbitrary number of layers.
-
-The rationale
-
-The parameters implied in this heuristic are the percolation radius $r_0$ and the percolation thresholds $\theta_i$ for each layer $i$.
+See Fig.~\ref{fig:method} for a schematic representation of the method.
+It can be implemented with a propagation heuristic or directly working on adjacency matrices.
+The rationale behind the conjunction of the thresholding of each layer variable and the distance thresholding relies on the idea that two points will interact is they are close enough, but also if they have a strong enough intensity of the activity or dimension captured by each layer, simultaneously for all layers considered.
+This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.
 
@@
 We apply the heuristic to urban morphology and road network topology measures in Europe.
-More precisely, a grid with resolution 50km of population density morphology indicators and road network topology indicators, has been computed on spatial moving windows for all European Union by \cite{raimbault2018urban}.
+The idea to combine urban form with network topology measures relies on the capture of the link between urban form and function as already mentioned, urban functions being assumed as distributed by transportation networks \citep{raimbault2018caracterisation}.
+
+More precisely, a grid of population density morphology indicators and road network topology indicators has been computed on spatial moving windows of width 50km for all European Union by \cite{raimbault2018urban}, with an offset resolution of 5km.
+We use this data to construct a two layers abstract network: a layer which variable is given by population density, and a second layer which variable is given by a network variable.
+Nodes are the center of cells (thus disposed in space on a grid of step 5km).
+We test the variable characterizing the second layer among the following characteristics of the road network within the corresponding window: number of edges, number of vertices, cyclomatic number and euclidian efficiency.
+These measures capture functional properties especially for the two last.
 
 The percolation on such an abstract network is a necessary condition in our case to link the different dimensions considered, namely population distribution and local road network properties.
-
-We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.
-
-% Precisons a ppresent les objets geographiques a partir  desqueles nous allons construire le reseau.
-% nous rappelns que nous cherchons a combiner prorpiete des territoires, que nous approximerons tres simplement par la districbution spatiale de la population, et propriete des reseaux routiers, que nous approximerons par des indicateirs topollogiques locaux.
-% ces differentes caracteristqiues ont ete calucles, dans le cadre de ce travail de modelisation a l'echelle mesocscopique de la coevolution entre reseaux et cilles, sur les feentres glissantes de taille 50km poru l'ensemble de l'europe.
-% nous en donnons ici l'illustration dans le cas de la France
-% [several infos on different regimes etc]
+We have therefore two levels of networks in our approach, namely the physical road network which local properties are taken here as input, and the abstract two layer network on which we do the percolation.
+We will in the following write $\theta_P$ for the threshold parameter of the population layer, and $\theta_N$ for the threshold parameter of the network layer.
+In practice, these parameters will be given in the following as quantile level of the corresponding variable, for an easier interpretation and conception of experience plans.
+The name of the variable considered will be written $v_N$.
 
@@
 
-We use this endogenous definition of regional urban systems produced by the percolation algorithm to evaluate their sustainability, in terms of conflicting objectives of economic integration and greenhouse gases emissions.
+As already detailed, recent forms of urbanization, in particular integrated mega-city regions, may imply different patterns of economic and transportation flows and thus exhibit various performances regarding different indicators of sustainability.
+We propose to use the endogenous definition of regional urban systems produced by the percolation algorithm to evaluate their sustainability, in terms of conflicting objectives of economic integration and greenhouse gases emissions.
 The definition of sustainability, or sustainable development, is by essence multi-dimensional \citep{viguie2012trade}.
@@
 
-The EDGAR database \citep{janssens2017edgar} (version 4.3.2) is used for local estimates of greenhouse gases emissions.
+We use the EDGAR database \citep{janssens2017edgar} (version 4.3.2) for local grid estimates of greenhouse gases emissions.
+We use the latest year available, namely 2012.
+As its resolution is much smaller than our indicator grid, we aggregate the emissions on the closer indicator point for each cell of the emission database.
+Since according to \cite{lashof1990relative} most of the greenhouse effect is caused by $\textrm{CO}_2$, and as in terms of emissions in the database we find that it represents $98.2\%$ in mass proportion of all gases, we only consider it.
 
-% explain indicators : convex hull.
-
-Applying a gravity model to each region, we estimate abstract transportation flows within each and extrapolate emissions by coupling with the Edgar emission database \citep{janssens2017edgar} and economic activities with a scaling law of population.
-More precisely, greenhouse gases emissions derived from economic and transportation flows are estimated with the following expression
+Applying a gravity model to each region, we estimate abstract transportation flows within each and use these to extrapolate emissions from the actual local emission from the Edgar database, and economic activities with a scaling law of population.
+More precisely, following \cite{raimbault2018indirect}, a potential flow between two points $i$ and $j$ can be estimated with the following expression
 
 \begin{equation}
-\phi_{ij} = \left(\frac{v_i v_j}{(\sum_k v_k)^2}\right)^\gamma \cdot \exp\left(\frac{-d_{ij}}{d_0}\right)
+\phi_{ij}^{(k)} = \left(\frac{v^{(k)}_i v^{(k)}_j}{(\sum_l v_l)^2}\right)^\gamma \cdot \exp\left(\frac{-d_{ij}}{d_0}\right)
 \end{equation}
 
-where $v_k$ are either effective local GHG emissions or population.
-Indeed, the economic activity follows relatively well scaling laws of populations \cite{bettencourt2007growth}, the exponent being dependant on the activity and the definition of areas on which it is estimated~\citep{cottineau2017diverse}.
+where $v^{(k)}_i$ are either population or effective local GHG emissions (indexed by $k = 1,2$ respectively), $d_{ij}$ the distance between the two points, $d_0$ a distance decay parameter, and $\gamma$ a scaling parameter.
+Indeed, the economic activity follows relatively well scaling laws of populations \citep{bettencourt2007growth}, the exponent being dependant on the activity and the definition of areas on which it is estimated~\citep{cottineau2017diverse}.
 
-The sum of all flows within the geographical span of the cluster (that we approximate as the convex Hull envelope of its points), allows us to approximate the cumulated potential emissions and economic activity.
-
-The corresponding relative indicators are defined as
-% TODO relative indiactors / normalized (by pop ?)
+The sum of all flows of points within the geographical span of the cluster (that we approximate as the convex Hull envelope of its points), allows us to approximate the cumulated potential emissions and economic activity.
+Writing clusters $K_c$ as this set of points, we define the total economic flow by $E_c = \sum_{i,j \in K_c} \phi_{ij}^{(1)}$ and the total emissions due to flows by $G_c = \sum_{i,j \in C_c} \phi_{ij}^{(2)}$.
+This allows to define a relative economic inefficiency by $e_c = 1 - \frac{\max_c E_c - E_c}{\max_c E_c - \min_c E_c}$ and relative potential emissions by $g_c = \frac{\max_c G_c - G_c}{\max_c G_c - \min_c G_c}$.
+Both indicators should be minimized for sustainability.
+Normalized indicators $\tilde{e}_c,\tilde{g}_c$ are defined in a similar way, but the extrema being computed on all other possible urban configurations with the same $\gamma,d_0$ values.
 
@@
 
-%Network construction
-
-%We percolate the population density layer with a network characteristic layer, that we test among number of edges, number of vertices, cyclomatic number and euclidian efficiency, which capture functional properties especially for the two last.
-
-% a partir de ces deux couches nous considerosn d'un [art la densite de population, d'autre part des proxus pour la densite ou perdormance du reseau ; avce les paramteres associes.
-% la logique geographique derriere cela est liee a la premiere loi de Tobler : deux lieus proches etc  ; ajoute a la logique de la masse (loi de gravite), mais selon different aspects du reseau (car difficile de savoir quelle dimenion a priori jouera le role "equivalent" d'une densite de population : autant laisser libre et explorer plutot que de fixer arbitrairemenbt - maxime d'un computational scientist converti, bien sur.
-% en pratqiuem nous contruison le reseau correspondant et isolns ses composantes connexes.
-% descirption du plan d'explericence.
-
-%\textbf{Two layers: } population density (threshold $\theta_P$) and network characteristics (threshold $\theta_N$) taken among \{Number of edges, Number of vertices, Cyclomatic number $\mu$, Euclidian efficiency $v$ \}; percolated with a radius $r_0$
-
-\cite{banos2012towards} network efficiency
-
-%\textbf{Rationale: } \textit{two locations will be in relation if they are close, have a high population density and given network characteristics.}
-
 In practice, the analysis is implemented using R and the igraph package.
 Source code, data and results are available on the open git repository of the project at \url{https://github.com/JusteRaimbault/UrbanMorphology}.
-
 The network is constructed by superposing the population density layer with the network layer, starting from the 5km resolution spatial fields for morphological and network indicators.
@@
 
+We recall that the euclidian performance of the network \citep{banos2012towards} is in our case $<d_e/d_n>$ where the average is taken on all origin-destination pairs in the network, $d_e$ is the euclidian distance and $d_n$ the network distance.
+Thus, it indeed increases with network performance, in consistence with the use done here through thresholding.
+
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[ht] 
+%\resizebox{10cm}{7cm}
+  {\includegraphics[width=0.49\linewidth]{figures/abssize_nodes.png}}
+  {\includegraphics[width=0.49\linewidth]{figures/relsize_nodes.png}}
+  \centering
+  \caption{Percolation transition.
+On the left, we plot the size of the largest cluster in each configuration in terms of nodes, as a function of the percolation radius $r_0$.
+Color gives the other percolation parameters.
+On the right, the plot is similar but with the size relative to the size of the largest cluster obtained with the maximal radius in each configuration.\label{fig:percolation}}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[ht] 
+%\resizebox{10cm}{7cm}
+  {\includegraphics[width=\linewidth]{figures/fractaldimension.png}}
+  \centering
+  \caption{Percolation transition.\label{fig:fractaldim}}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%
+\subsection{Percolation transition and fractal dimension}
+
+In its application to road networks by \cite{arcaute2016cities}, the structure of the national urban system for UK is captured by studying the percolation transition, i.e.
+the variation of the size of the largest cluster as a function of the percolation radius.
+As this signature is tightly linked to historical, cultural and geographical conditions, the application to different urban systems should yield different results.
+We study here this property, for different threshold parameter values.
+We make the radius vary betweem 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all network variables, and three network thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.
+
+The absolute and relative sizes of the largest cluster are plotted in Fig.~\ref{fig:percolation} as a function of the percolation radius.
+This aspect first gives methodological information on multilayer percolation.
+Indeed, comparing the result with $\theta_N = 0$ (single layer percolation) with positive values of $\theta_N$ shows a significantly different behavior.
+As expected, absolute size are much smaller, but when looking
+
+% ! largest cluster size in terms of edges makes no sense with the implementation we took - nodes only
+%      nwcol         popthq         nwthq          radius         nwparam             fractdim     
+% ecount  :186   Min.   :0.85   Min.   :0.00   Min.   :  8000   Length:1023        Min.   :0.9239  
+% euclPerf:279   1st Qu.:0.85   1st Qu.:0.00   1st Qu.: 31000   Class :character   1st Qu.:1.2639  
+% mu      :279   Median :0.85   Median :0.80   Median : 54000   Mode  :character   Median :1.3306  
+% vcount  :279   Mean   :0.85   Mean   :0.55   Mean   : 54000                      Mean   :1.3225  
+ %               3rd Qu.:0.85   3rd Qu.:0.95   3rd Qu.: 77000                      3rd Qu.:1.3909  
+ %               Max.   :0.85   Max.   :0.95   Max.   :100000                      Max.   :1.6058                                                                                    NA's   :337     
+%   fractdimsd     fractdimadjrsquared   fractrelsd    
+% Min.   :0.0544   Min.   :0.5111      Min.   :0.0395  
+% 1st Qu.:0.0966   1st Qu.:0.7016      1st Qu.:0.0700  
+% Median :0.1328   Median :0.7467      Median :0.1043  
+% Mean   :0.1317   Mean   :0.7456      Mean   :0.1009  
+% 3rd Qu.:0.1578   3rd Qu.:0.7981      3rd Qu.:0.1232  
+% Max.   :0.2381   Max.   :0.8642      Max.   :0.1958  
+% NA's   :337      NA's   :337         NA's   :337
+
+We study also the evolution of the fractal dimension of clusters as a function of $r_0$.
+We do not plot the standard error $\sigma$ of fractal dimensions for visibility purposes, but their relative value given by $\alpha / \sigma\left[\alpha\right]$ is in average 0.10 and in maximum 0.196 on all points.
+Regarding the variability as a function of the percolation radius $r_0$,
+
+%  fractdimsignif
+%# A tibble: 11 x 5
+%# Groups:   nwcol, popthq, nwthq [?]
+%   nwcol    popthq nwthq nwparam       signif
+%   <fct>     <dbl> <dbl> <chr>          <dbl>
+% 1 ecount     0.85  0    ecount0      -0.105 
+% 2 ecount     0.85  0.8  ecount0.8    -0.0452
+% 3 euclPerf   0.85  0    euclPerf0    -0.108 
+% 4 euclPerf   0.85  0.8  euclPerf0.8   0.0360
+% 5 euclPerf   0.85  0.95 euclPerf0.95 -0.108 
+% 6 mu         0.85  0    mu0          -0.108 
+% 7 mu         0.85  0.8  mu0.8        -0.108 
+% 8 mu         0.85  0.95 mu0.95        0.375 
+% 9 vcount     0.85  0    vcount0      -0.105 
+%10 vcount     0.85  0.8  vcount0.8    -0.0416
+%11 vcount     0.85  0.95 vcount0.95   -0.0868
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Extracting endogenous mega-city regions}
+
 The experience plan is a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the network variable considered.
 We also make $\gamma$ and $d_0$ vary.
-
-We recall that the euclidian performance of the network is in our case $<d_e/d_n>$ where the average is taken on all origin-destination pairs in the network, $d_e$ is the euclidian distance and $d_n$ the network distance.
-Thus, it indeed increases with network performance, in consistence with the use done here through thresholding.
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Extracting endogenous mega-city regions}
-
-%%%%%%%%%%%%%%%%%%%%
 \begin{figure*}[ht] 
 %\resizebox{10cm}{7cm}
-{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}\\
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}
+{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}
+  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}\\
   {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
+   {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
   \centering
@@
   \caption{Examples of obtained clusters for different parameter values.
-In the third case for example, we obtain the urban regions of West midlands and London in the UK, Randstad merged with Rhein-Rhur and Rhein-Main in Germany, Paris in France, also with capital cities such as Copenhaguen, Stockholm and Helsinki.
+In the top-right case for example ($\theta_P = 0.9$, $\theta_N = 0.8$, variable \texttt{vcount},$r_0 = 8km$), we obtain the urban regions of West midlands and London in the UK, Randstad merged with Rhein-Rhur and Rhein-Main in Germany, Paris in France, also with capital cities such as Copenhaguen, Stockholm and Helsinki.
 There is no cluster in South Europe in that case, due to the high population density threshold.\label{fig:exclusters}}
@@
 We show some examples in Fig.~\ref{fig:exclusters}.
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Percolation transition and fractal dimension}
-
-%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
-%\resizebox{10cm}{7cm}
-  %{\includegraphics[width=\linewidth]{figures/}}
-  \centering
-  \label{fig:percolation}
-  \caption{Percolation transition.}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%
-
-In its application to road networks by \cite{arcaute2016cities}, the structure of the national urban system for UK is captured by studying the percolation transition, i.e.
-the variation of the size of the largest cluster as a function of the percolation radius.
-As this signature is tightly linked to historical, cultural and geographical conditions, the application to different urban systems should yield different results.
-We study here this property, for different threshold parameter values.
-The relative size of the largest cluster is plotted in Fig.~\ref{fig:percolation} as a function of the percolation radius.
-
-This aspect furthermore gives methodological information on multilayer percolation. Indeed, comparing the result with $\theta_N = 0$ (single layer percolation) with $\theta_N = 0.8$
 

### 2019-03-17 Models/Percolation/percolationFunctions.R

@@ -104,7 +104,7 @@ conditionalPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
 graphPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
                              minclustsize=25,
                              xcol="lonmin",ycol="latmin",popcol="totalPop",
-                             withMaps=T,
+                             withMaps=F,
                              resdir=paste0(Sys.getenv('CS_HOME'),'/UrbanMorphology/Results/Percolation/Maps2/')
 ){
 
@@ -127,8 +127,10 @@ graphPercolation <- function(d,radius,popthq,nwcol,nwthq,gamma,decay,
   for(cnum in which(comps$csize >= minclustsize)){
     cpoints$cluster[comps$membership==cnum]=k
     currentcomp = induced_subgraph(g,which(comps$membership==cnum))
+    clustsizesnodes=append(clustsizesnodes,length(V(currentcomp)));
+    clustsizesedges=append(clustsizesedges,length(E(currentcomp)))
+    # note : keeping short edges only will make no sense for number of edges (prop to number of nodes) - for diameter neither ! (would cut the components)
     clustdiameters=append(clustdiameters,diameter(currentcomp))
-    clustsizesnodes=append(clustsizesnodes,length(V(currentcomp)));clustsizesedges=append(clustsizesedges,length(E(currentcomp)))
     k=k+1
   }
   

### 2019-03-17 Models/Percolation/percolationResults.R

@@ -8,7 +8,8 @@ load('../../Data/consolidated/indics.RData')
 
 #resstamp = '20181016_024956'
 #purpose='directSampling'
-resstamp = '20190316_182550'
+#resstamp = '20190316_182550'
+resstamp='20190316_235006'
 purpose='directSampling_radius'
 
 #resstamp = 'test_20190316_154614'
@@ -43,6 +44,19 @@ for(k in unique(full$paramrow)){
 #####
 # percolation transition
 
+#sapply(res[[1]],length)
+#sapply(res,length)
+#sapply(res,function(l){length(l[[1]])})
+#max(sapply(res,function(l){abs(length(l[[1]])-length(l[[10]]))+abs(length(l[[10]])-length(l[[11]]))+abs(length(l[[11]])-length(l[[12]]))+abs(length(l[[12]])-length(l[[13]]))}))
+# which(params$nwcol=='mu'&params$nwthq==0.95&params$radius==75000)
+
+
+# -> need to filter NAs
+# BEFORE the lapply (not same length) - ULTRA DIRTY data structure -> really need to stop using R...
+resfilt = list();paraminds=c();kk=1;for(k in 1:length(res)){if(length(which(is.na(res[[k]][[1]])))==0){resfilt[[kk]]=res[[k]];paraminds=append(paraminds,k);kk=kk+1}}
+res=resfilt
+params=params[paraminds,]
+
 toremove=c("pops","morans","avgdists","entropies","slopes","efficiencies","emissions","totalemissions")
 for(k in 1:length(res)){res[[k]]$paramrow=rep(k,length(res[[k]]$areas))}
 fulltrans = data.frame(matrix(unlist(lapply(res,function(l){for(tr in toremove){l[[tr]]=NULL};unlist(data.frame(t(data.frame(l))))})),byrow=T,ncol=length(res[[1]])-length(toremove)))
@@ -50,17 +64,87 @@ colnames(fulltrans)<-names(res[[1]])[c(-9:-2)]
 params$paramrow=1:nrow(params)
 fulltrans=as.tbl(left_join(fulltrans,params))
 
+dir.create(paste0(resdir,'transition'))
+
 # not normalized
 fulltrans$nwparam = paste0(fulltrans$nwcol,fulltrans$nwthq)
[…]

### 2019-03-17 Docs/Papers/Percolation/arxiv/v1/paper.tex
paper_submitted

--- 2019-03-17 Docs/Papers/Percolation/paper.tex
@@
-\documentclass{jimis-en}
+\documentclass{article}
 \pdfoutput=1 %forces arXiv/HAL to compile with PDFLaTeX
@@
 \usepackage{graphicx}
+\usepackage{natbib}
+\usepackage[margin=2.4cm]{geometry}
+
+\begin{document}
 
@@
 
-\author[1,2,3,*]{Juste RAIMBAULT}
-\affil[1]{UPS CNRS 3611 ISC-PIF, France} 
-\affil[2]{CASA, UCL, UK}
-\affil[3]{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s}
+\author{Juste RAIMBAULT$^{1,2,3,\ast}$\\
+$^1$ UPS CNRS 3611 ISC-PIF, France\\
+$^2$ CASA, UCL, UK\\
+$^3$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\\medskip
+$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
+}
 
-% corresponding author with his/her email
-\corrauthor{juste.raimbault@polytechnique.edu}
-
-%% Information provided by JIMIS:
-%ISSN
-% DOI 
-\doi{10.18713/JIMIS-ddmmyy-v-a}{}
-% reviewing process: first date = submission, second date = acceptance 
-\review{Month-in-letters Day Year}{Month-in-letters Day Year}
-% Volume and Year of the issue
-\publication{N}{YYYY}
-
-%% Information provided by the Editors:
-% Issue Title 
-\issue{Title of the interdisciplinary issue}
-% List of k Editors
-\editors{}
-
-\begin{document}
+\date{}
 
@@
 The variable parametrization allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions).
-This suggests a customizable spatial design of policies to develop sustainable territories.}
-
-\keywords{Road network; Multi-dimensional percolation; European urban system; Mega-city region}
+This suggests a customizable spatial design of policies to develop sustainable territories.\\\medskip
+\textbf{Keywords: } Road network; Multi-dimensional percolation; European urban system; Mega-city region
+}
 
 \section{Introduction}
-
-\strut
-\vspace{-4ex}
 
@@
 \cite{piovani2017urban} apply road network percolation at the mesoscopic scale of London metropolitan area, in relation with a retail location model.
-In spatial statistics, this method can be used to characterize the spatial morphology of point patterns \cite{huynh2018characterisation}.
+In spatial statistics, this method can be used to characterize the spatial morphology of point patterns \citep{huynh2018characterisation}.
 
@@
 However, such systems are known to be multidimensional, for example through the superposition of the morphological dimension of networks, and the functional properties of the urban environment \citep{burger2012form}.
-The link between urban form and function remains in particular an open question \citep{batty1994fractal}, but more generally the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \cite{perez2016agent}.
+The link between urban form and function remains in particular an open question \citep{batty1994fractal}, but more generally the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}.
 This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic.
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/principle.pdf}}
+  {\includegraphics[width=\linewidth]{principle.pdf}}
   \centering
@@
 Nodes are the center of cells (thus disposed in space on a grid of step 5km).
-We test the variable characterizing the second layer among the following characteristics of the road network within the corresponding window: number of edges, number of vertices, cyclomatic number and euclidian efficiency.
+We test the variable characterizing the second layer among the following characteristics of the road network within the corresponding window: number of edges $N_E$, number of vertices $N_V$, cyclomatic number $\mu$ and euclidian efficiency $v_0$.
 These measures capture functional properties especially for the two last.
@@
 In practice, the analysis is implemented using R and the igraph package.
-Source code, data and results are available on the open git repository of the project at \url{https://github.com/JusteRaimbault/UrbanMorphology}.
+Source code, data and results are available on the open git repository of the project at \texttt{https://github.com/JusteRaimbault/UrbanMorphology}.
 The network is constructed by superposing the population density layer with the network layer, starting from the 5km resolution spatial fields for morphological and network indicators.
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.49\linewidth]{figures/abssize_nodes.png}}
-  {\includegraphics[width=0.49\linewidth]{figures/relsize_nodes.png}}
+  {\includegraphics[width=0.49\linewidth]{abssize_nodes.png}}
+  {\includegraphics[width=0.49\linewidth]{relsize_nodes.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/fractaldimension.png}}
+  {\includegraphics[width=\linewidth]{fractaldimension.png}}
   \centering
-  \caption{Percolation transition.\label{fig:fractaldim}}
+  \caption{Fractal dimension.
+We plot for each parametrization given by the curve color the evolution of the fractal dimension $\alpha$ as a function of $r_0$.
+Standard errors are not plotted for readability.\label{fig:fractaldim}}
 \end{figure*}
@@
 Indeed, comparing the result with $\theta_N = 0$ (single layer percolation) with positive values of $\theta_N$ shows a significantly different behavior.
-As expected, absolute size are much smaller, but when looking
+As expected, absolute size are much smaller, but when looking at relative sizes we observe that the abrupt steps typical to percolation transitions have different distributions across the different parametrizations.
+The more regular curve seems to be the standard percolation on population only, whereas at $\theta_N = 0.95$, different network variables produce either very early transitions (for $\mu$ for example) or very late (for $N_V$).
+Also, changing of scale compared to \cite{arcaute2016cities} gives more steps and less abrupts curves in general, confirming the integration of subsystems with different structures in our analysis and the importance of scale in such analysis.
+As the addition of a layer also changes drastically the results, one should stay careful when switching from a mono-dimensional percolation to a multi-dimensional percolation.
 
@@
 We study also the evolution of the fractal dimension of clusters as a function of $r_0$.
-We do not plot the standard error $\sigma$ of fractal dimensions for visibility purposes, but their relative value given by $\alpha / \sigma\left[\alpha\right]$ is in average 0.10 and in maximum 0.196 on all points.
-Regarding the variability as a function of the percolation radius $r_0$,
+Following \cite{arcaute2016cities}, we estimate the fractal dimension of clusters $\alpha$ with a simple OLS regression between cluster size and cluster diameter, namely $\log N_c = k + \alpha \cdot \log \delta_c$ where $N_c$ is the size of cluster $c$ and $\delta_c$ its diameter.
+As a negative result, that could be due to the abstract nature of our network, a clear maximum in the value of the fractal dimension can not be found.
+Either it is located at resolution that our method can not reached due to the minimal 5km limit imposed by the abstraction, or it does not exist when coupling dimensions.
+Determining which assumption is more plausible is out of the scope of this paper.
+We do not plot the standard error $\sigma$ of fractal dimensions for visibility purposes, but their relative value given by $\alpha / \sigma\left[\alpha\right]$ is in average 0.10 and in maximum 0.196 on all points, meaning that these estimations remain however consistent.
+Regarding the variability as a function of the percolation radius $r_0$, studying the difference $(\alpha - \sigma\left[\alpha\right])_M - (\alpha - \sigma\left[\alpha\right])_m$ where the first is taken at maximum for $\alpha$ and the other at minimum, shows that the configuration for $\mu$ and $\theta_N=0.95$ has a clearly significant maximum (difference at 0.38).
+For this coupling, the endogenous structure given by the maximum may be defined.
+Other configurations yield non-significant maximums.
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\subsection{Extracting endogenous mega-city regions}
-
-The experience plan is a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the network variable considered.
-We also make $\gamma$ and $d_0$ vary.
-
-We systematically explore the clusters obtained for 4800 parameter configurations.
-
-%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
+\begin{figure}[h!] 
 %\resizebox{10cm}{7cm}
-{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}\\
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
-   {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
+{\includegraphics[width=0.49\textwidth]{totalPop4183694_00056402_ecount850_radius8000.png}}
+  {\includegraphics[width=0.49\linewidth]{totalPop2219780_36719597_vcount378_radius8000.png}}\\
+  {\includegraphics[width=0.49\linewidth]{totalPop1474347_36891685_vcount595_radius50000.png}}
+   {\includegraphics[width=0.49\linewidth]{totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
   \centering
@@
 There is no cluster in South Europe in that case, due to the high population density threshold.\label{fig:exclusters}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h!] 
+%\resizebox{10cm}{7cm}
+  {\includegraphics[width=0.7\linewidth]{full_effective_pareto.png}}
+  \centering
+  \caption{Point clouds of region-level indicators, namely population and emissions, for different parametrizations, given by the color.
+Each point represent an endogenous urban region.\label{fig:paretos}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%
+\subsection{Extracting endogenous mega-city regions}
+
+We now switch the experience plan to a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the network variable considered, and also make $\gamma$ and $d_0$ vary.
+We systematically explore the clusters obtained for 4800 parameter configurations, such that for all network variables, $\theta_P \in \{ 0.8 ; 0.9 ; 0.95 \}$, $\theta_N \in \{0 ; 0.8 ; 0.95 \}$, $r_0 \in \{ 8 ; 10 ; 15 ; 20 ; 50\}$ km, $\gamma \in \{ 0.5 ; 1 ; 1.5 ; 2\}$, and $d_0 \in \{ 0.1 ; 1 ; 10 ; 50 ; 100\}$ km.
+
+We obtain very different endogenous morphologies for the different parametrizations.
+Maps reveal that some configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}.
+These are here defined endogenously from the bottom-up and have a priori no reason to coincide with these functional regions.
+We show some examples in Fig.~\ref{fig:exclusters}.
+The first map of this figure, obtained for high population and network thresholds ($\theta_P = 0.95$ and $\theta_N = 0.9$), but a low radius $r_0 = 8$km and edge count $N_E$ as network variable, include several mega-city regions described by \citep{hall2006polycentric}, namely London metropolitan area, the Randstad in Netherland, the Rhein-Main and Rhein-Ruhr in Germany, Greater Paris in France, Brussels area in Belgium.
+The same parameters with $\theta_N = 0$ yield not exactly the same regions, as confirmed by the transition curves in Fig.~\ref{fig:percolation}, what means that our approach taking into account two dimensions may capture effective processes of mega-city regions, in particular by including the road network which is crucial as these regions are integrated in terms of flows.
+The bottom-left map show an example of large clusters emerging in UK and in the center of Europe, the South remaining largely disconnected.
+Finally, the last map shows the result obtained with a very high radius $r_0 = 54$km, with a giant cluster spanning most of Europe.
+UK is still disconnected and the transition where it connects happens at $r_0 = 55$km.
+This does not necessarily mean that UK should be disconnected from continental Europe, as we considered geographic distances only, hiding the high speed connection of the Channel tunnel.
+
+The behavior of sustainability indicators for different population, network and distance thresholds yield different distributions of performances across clusters within a configuration but also between configurations.
+Before considering the flow-based indicators described above, we can already study basic measures such as population $P_c$ and effective emissions $EM_c$, taken as the sum within the cluster of their values at each point.
+We show in Fig.~\ref{fig:paretos} point clouds of $\log EM_c$ against $1 - \log P_c$ for some configurations.
+Indeed, regarding the population it contains, an area can be more or less efficient in terms of emissions.
+Seeing the population as an objective to be maximized (thus the plotted value to be minimized), we observe a Pareto front for all points (i.e.
+all clusters across all configurations), but also no dominating point for each configuration.
+Some clusters are therefore optimal compromises in the Pareto sense in each configuration, while some are dominated and thus not efficient.
+
+% not really interesting + messy figure -> do not show this, morphology is more interesting
+%\includegraphics[height=0.8\textheight]{figures/aggreg_paretos_radiuspopthq.png}
+
+%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[!ht] 
+%\resizebox{10cm}{7cm}
+  {\includegraphics[width=\linewidth]{aggreg_morpho_pc1-emissions_targeted.png}}
+  \centering
+  \caption{Aggregated values of normalized potential emissions $\sum_c \tilde{g}_c$, as a function of the first morphological principal component (PC1), for varying values of parameters $d_G$ (rows) and $\gamma_G$ (columns).
+Other intermediate values for these parameters yield similar behaviors.
+As PC1 is mainly linked to monocentricity, there seems to exist an optimal intermediate level of monocentricity for emissions alone.
+Color level give the share of population within the considered clusters in comparison to all European population.\label{fig:emissions-pc1}}
 \end{figure*}
@@
 
-We obtain different endogenous morphologies.
-
-Maps reveal that some configurations resemble the actual distribution of European mega-city regions, which are functionally integrated polycentric urban areas \citep{hall2006polycentric}.
-These are here defined endogenously from the bottom-up and have a priori no reason to coincide with these functional regions.
-
-We show some examples in Fig.~\ref{fig:exclusters}.
-
-%%%%%%%%%%%%%%%%%%%
-\subsection{Pareto fronts for sustainability}
-
-We show therein that different population, network and distance thresholds yield different performances in terms of sustainability.
-
-%\textit{Superposing Pareto front for observed population and emissions, on all clusters.}
-%\includegraphics[height=0.8\textheight]{figures/}
-
 %%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
+\begin{figure*}[!ht] 
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/full_effective_pareto.png}}
+  {\includegraphics[width=\linewidth]{aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
   \centering  
-  \label{fig:paretos}
-  \caption{Point clouds of region-level indicators, namely population and emissions, for different parametrizations.}
+  \caption{Relative potential emissions $\sum_c g_c$ against relative potential economic unefficiency $\sum_c e_c$ (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows).
+Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
 \end{figure*}
 %%%%%%%%%%%%%%%%%%%%
-
-% -> full fronts in reserve slide
-
-% Pareto fronts for aggregated indicators}{
-%\textit{Aggregated sustainability indicators suggest some configurations are more Pareto efficient (high $\gamma$ regime, activities with high added value).}
-
-%\includegraphics[height=0.8\textheight]{figures/aggreg_paretos_radiuspopthq.png}
 
@@
 
-When considering the aggregated indicators for a parametrization of endogenous city regions, one can relate them to morphological indicators for population density spatial distribution, computed by~\cite{raimbault2018calibration}, that we average on areas.
+We now consider the sustainibility indicators, aggregated for a configuration on all clusters.
+For a given parametrization of endogenous city regions, one can relate them to morphological indicators for population density spatial distribution, computed by~\cite{raimbault2018calibration}, that we average on clusters.
 This establishes a link between urban morphology and sustainibility.
 A principal component analysis on considered points yield 96\% of variance with two components, and 73\% explained by the first component alone.
-The first component relates to a level of monocentricity ($PC1 = -0.3\cdot I + 0.54 \cdot \bar{d} + 0.51\cdot \varepsilon + 0.59 \cdot \alpha$).
+The first component relates to a level of monocentricity ($PC1 = -0.3\cdot I + 0.54 \cdot \bar{d} + 0.51\cdot \varepsilon + 0.59 \cdot h$ where $I$ is Moran index, $\bar{d}$ average distance, $\epsilon$ entropy, and $h$ level of hierarchy).
 
-%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
-%\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_pc1-emissions_targeted.png}}
-  \centering
-  \caption{Aggregated values of normalized potential emissions, as a function of the first morphological principal component (PC1), for varying values of parameters $d_G$ (rows) and $\gamma_G$ (columns).
-Other intermediate values for these parameters yield similar behaviors.
-As PC1 is mainly linked to monocentricity, there seems to exist an optimal intermediate level of monocentricity for emissions alone.\label{fig:emissions-pc1}}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%
+We show in Fig.~\ref{fig:emissions-pc1} the value of $\sum_c \tilde{g}_c$ as a function of the first morphological principal component, for extreme values of $\gamma$ and $d_0$.
+There seems to exist an optimal intermediate value for PC1 regarding the minimization of normalized indicator for emissions only.
+This would correspond to an intermediate level of monocentricity, meaning that urban areas which are too polycentric and spread would emit more, but also areas that are too much monocentric.
+This behavior does not occur for long-range $d_0 = 100$km and low-hierarchy $\gamma=0.5$ interactions.
+The mostly monocentric but emitting configurations capture most of population (given by the level of color), whereas the intermediate configuration capture around half of the population, what means that these low-emissions potential urban regions can cover a significant part of European population.
 
-As shown in Fig.~\ref{fig:emissions-pc1}, there seems to exist an optimal intermediate level of monocentricity regarding the normalized indicator for emissions only, except for long-range and low-hierarchy interactions.
-
-%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
-%\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
-  \centering  
-  \caption{Relative potential emissions against relative potential economic unefficiency (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows).
-Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
-\end{figure*}
-%%%%%%%%%%%%%%%%%%%%
-
-However, when considering both emissions and economic indicators, urban form then acts as a compromise variable, moving points on Pareto fronts, as shown in Fig.~\ref{fig:paretos-relative}.
-In some case, highly monocentric areas can be a good compromise, whereas the intermediate optimal for emissions may yield highly inefficient areas.
-This unveils morphological trad-offs, confirming that there is no optimal urban form, but different compromises regarding the conflicting indicators.
+However, when considering both emissions and economic indicators, urban form then acts as a compromise variable.
+We show in Fig.~\ref{fig:paretos-relative} the point clouds of $\sum_c g_c$ against $\sum_c e_c$, which produce clear Pareto fronts, which shape varies with $\gamma$ and $d_0$.
+As the color level gives the value of PC1, we can see the points on the different fronts with very different morphological properties.
+In some case, highly monocentric areas (yellow points) can be a good compromise, whereas the intermediate optimal for emissions shown before may yield highly inefficient areas (dominated green points).
+For example, considering the fronts for $\gamma = 2$ which have both very similar shape, the points with the lowest emissions are on the top-left of the front and correspond to the optimal unveiled in Fig.~\ref{fig:emissions-pc1}.
+These have however a very low economic efficiency (high inefficiency) and small improvements can be done with the points below, before switching to a totally different urban form with a high value of PC1 (yellow points, highly monocentric).
+Increasing more the economic efficiency is then at the price of much more emissions, with more polycentric areas.
+This analysis therefore unveils morphological trade-offs, confirming that there is no optimal urban form, but different compromises regarding the conflicting sustainability indicators.
 
@@
 
-Further work can consist in the use of calibration heuristics to find in a more robust way optimal parameter values.
+Further work may consist in the use of calibration heuristics to find in a more robust way optimal parameter values.
 The OpenMOLE model exploration platform provides a transparent access to genetic algorithms for multi-objective optimization \citep{reuillon2013openmole}.
@@
 
-% -> depending on results
+An other development would consist in extrapolating transportation flows with a spatially explicit gravity and transportation flow model as a kind of simplified four step model \citep{mcnally2000four}.
+It could then be adjusted on actual transportation flows emissions database which are also available in the Edgar database.
+The corresponding gravity parameters could then be used within the economic and emissions potentials, and the sustainability patterns produced compared with the hypothetical ones we produced here.
 
-%Grid sampling to explore regions rapidly limited
-
-%\textbf{Implications}
-
-%$\rightarrow$
-
-% detail method to extrapolate parameters of gravity flows
-
-% interesting result : compare these emission with the synthetic ?
-(then just a refinment ?) -> yes, better.
-show new Pareto fronts with this fixed parameters for only emissions, and varying for economic ?
-
-Extrapolating transportation flows with a spatially explicit gravity and flow model can allow to compare these with actual transportation flows in the emission database, and yield a possible calibration.
-These extrapolated parameters could then be used within the economic and emissions potentials.
-
-An other potential development would imply crossing our endogenous definitions of urban regions with socio-economic databases, and compute indicators implied in other dimensions of sustainability, for example related to socio-economic inequalities, spatial distribution of accessibilities, activities with different scaling exponents.
-% ex : at least different scaling exponent.
+Finally, an important development would imply crossing our endogenous definitions of urban regions with socio-economic databases, and compute indicators implied in other dimensions of sustainability, for example related to socio-economic inequalities, spatial distribution of accessibilities, or activities with different scaling exponents.
+This includes the mitigation of spatial inequalities and segregation \citep{tammaru2015multi}, which are an important dimension of sustainibility.
 
@@
 
-This suggests policies in terms of regional integration to increase the sustainability of mega-city regions.
-
-The way such results can be transferred to policy-making recommandations remains an open question, but Pareto-optimal configurations can be used for the planning of regional transportation networks for example, or to design policies for the distribution of subsidies.
+Our work suggests the possibility to design policies in terms of regional integration to increase the sustainability of mega-city regions.
+The way such results could actually be transferred to policy-making recommandations remains an open question, but Pareto-optimal configurations can be used for the planning of regional transportation networks for example, or to design policies for the distribution of subsidies.
+Indeed, privileging some infrastructure developments but also collaborations between urban centers can be seen as an aspect of a small scale planning, or territorial strategy.
+As we integrated potential flows that would result from such development, and consider their economic and emissions consequences, and did it in an endogenous way, we suggest that evidence-based strategies for territorial development at the European level could be inspired by this work.
+This would naturally imply a more thorough data integration, model calibration and operationalization.
 
@@
 
-In conclusion, our multilayer percolation approache captures in a way the multi-dimensionality of urban systems and a link between form and function.
+In conclusion, our multilayer percolation approach captures in a way the multi-dimensionality of urban systems and a link between form and function in urban system.
+Its application to the issue of sustainable mega-city regions shows its potentialities.
+This work also illustrates the importance of following data-driven paradigms even when developing, as what is understood of the behavior of the heuristic is through its application to real data and issues.
 
-%$\rightarrow$ Empirical and theoretical research directed towards concrete policy-making applications. \textbf{Need for more data-driven approaches.}
-%First step towards systematic benchmarks and multi-modeling. \textbf{Need for more systematic model exploration.} % model coupling / benchmarking
+\begin{thebibliography}{}
 
-%$\rightarrow$ Towards multi-scalar approaches ? \textbf{Need for more integrated models.}
- %Model integration and multi-scalarity ? \textbf{Need for more integrated models.}
+\bibitem[\protect\citeauthoryear{Arbabi, Mayfield, and McCann}{Arbabi
+  \textit{et~al.}}{2019}]{arbabi2019development}
+Arbabi H., Mayfield M., McCann P. (2019).
+\newblock On the development logic of city-regions: inter-versus intra-city
+  mobility in england and wales.
+\newblock {\em Spatial Economic Analysis\/}, 1--20.
 
-%$\rightarrow$ Multidimensionality of urban systems ? \textbf{Need for more interdisciplinarity.}
+\bibitem[\protect\citeauthoryear{Arcaute, Molinero, Hatna, Murcio, Vargas-Ruiz,
+  Masucci, and Batty}{Arcaute \textit{et~al.}}{2016}]{arcaute2016cities}
+Arcaute E., Molinero C., Hatna E., Murcio R., Vargas-Ruiz C., Masucci A.~P.,
+  Batty M. (2016).
+\newblock Cities and regions in britain through hierarchical percolation.
+\newblock {\em Royal Society open science\/}~{\em 3\/}(4), 150691.
 
-\bibliographystyle{jimis-en}
-\bibliography{biblio}
+\bibitem[\protect\citeauthoryear{Banos and Genre-Grandpierre}{Banos and
+  Genre-Grandpierre}{2012}]{banos2012towards}
+Banos A., Genre-Grandpierre C. (2012).
+\newblock Towards new metrics for urban road networks: Some preliminary
+  evidence from agent-based simulations.
+\newblock In {\em Agent-based models of geographical systems}, pp.\  627--641.
+  Springer.
+
+\bibitem[\protect\citeauthoryear{Barth{\'e}lemy}{Barth{\'e}lemy}{2011}]{barthelemy2011spatial}
+Barth{\'e}lemy M. (2011).
+\newblock Spatial networks.
+\newblock {\em Physics Reports\/}~{\em 499\/}(1-3), 1--101.
+
+\bibitem[\protect\citeauthoryear{Batty and Longley}{Batty and
+  Longley}{1994}]{batty1994fractal}
+Batty M., Longley P.~A. (1994).
+\newblock {\em Fractal cities: a geometry of form and function}.
+\newblock Academic press.
+
+\bibitem[\protect\citeauthoryear{Bettencourt, Lobo, Helbing, K{\"u}hnert, and
+  West}{Bettencourt \textit{et~al.}}{2007}]{bettencourt2007growth}
+Bettencourt L.~M., Lobo J., Helbing D., K{\"u}hnert C., West G.~B. (2007).
+\newblock Growth, innovation, scaling, and the pace of life in cities.
+\newblock {\em Proceedings of the national academy of sciences\/}~{\em
+  104\/}(17), 7301--7306.
+
+\bibitem[\protect\citeauthoryear{Boccaletti, Bianconi, Criado, Del~Genio,
+  G{\'o}mez-Gardenes, Romance, Sendina-Nadal, Wang, and Zanin}{Boccaletti
+  \textit{et~al.}}{2014}]{boccaletti2014structure}
+Boccaletti S., Bianconi G., Criado R., Del~Genio C.~I., G{\'o}mez-Gardenes J.,
+  Romance M., Sendina-Nadal I., Wang Z., Zanin M. (2014).
+\newblock The structure and dynamics of multilayer networks.
+\newblock {\em Physics Reports\/}~{\em 544\/}(1), 1--122.
+
+\bibitem[\protect\citeauthoryear{Burger and Meijers}{Burger and
+  Meijers}{2012}]{burger2012form}
+Burger M., Meijers E. (2012).
+\newblock Form follows function? linking morphological and functional
+  polycentricity.
+\newblock {\em Urban studies\/}~{\em 49\/}(5), 1127--1149.
+
+\bibitem[\protect\citeauthoryear{Callaway, Newman, Strogatz, and
+  Watts}{Callaway \textit{et~al.}}{2000}]{callaway2000network}
+Callaway D.~S., Newman M.~E., Strogatz S.~H., Watts D.~J. (2000).
+\newblock Network robustness and fragility: Percolation on random graphs.
+\newblock {\em Physical review letters\/}~{\em 85\/}(25), 5468.
+
+\bibitem[\protect\citeauthoryear{Cottineau, Finance, Hatna, Arcaute, and
+  Batty}{Cottineau \textit{et~al.}}{2018}]{cottineau2018defining}
+Cottineau C., Finance O., Hatna E., Arcaute E., Batty M. (2018).
+\newblock Defining urban clusters to detect agglomeration economies.
+\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
+  2399808318755146.
+
+\bibitem[\protect\citeauthoryear{Cottineau, Hatna, Arcaute, and
+  Batty}{Cottineau \textit{et~al.}}{2017}]{cottineau2017diverse}
+Cottineau C., Hatna E., Arcaute E., Batty M. (2017).
+\newblock Diverse cities or the systematic paradox of urban scaling laws.
+\newblock {\em Computers, environment and urban systems\/}~{\em 63}, 80--94.
+
+\bibitem[\protect\citeauthoryear{Crucitti, Latora, and Porta}{Crucitti
+  \textit{et~al.}}{2006}]{crucitti2006centrality}
+Crucitti P., Latora V., Porta S. (2006).
+\newblock Centrality measures in spatial networks of urban streets.
+\newblock {\em Physical Review E\/}~{\em 73\/}(3), 036125.
+
+\bibitem[\protect\citeauthoryear{Ducruet and Beauguitte}{Ducruet and
+  Beauguitte}{2014}]{ducruet2014spatial}
+Ducruet C., Beauguitte L. (2014).
+\newblock Spatial science and network science: Review and outcomes of a complex
+  relationship.
+\newblock {\em Networks and Spatial Economics\/}~{\em 14\/}(3-4), 297--316.
+
+\bibitem[\protect\citeauthoryear{Hackett, Cellai, G{\'o}mez, Arenas, and
+  Gleeson}{Hackett \textit{et~al.}}{2016}]{hackett2016bond}
+Hackett A., Cellai D., G{\'o}mez S., Arenas A., Gleeson J.~P. (2016).
+\newblock Bond percolation on multiplex networks.
+\newblock {\em Physical Review X\/}~{\em 6\/}(2), 021002.
+
+\bibitem[\protect\citeauthoryear{Hall and Pain}{Hall and
+  Pain}{2006}]{hall2006polycentric}
+Hall P.~G., Pain K. (2006).
+\newblock {\em The polycentric metropolis: learning from mega-city regions in
+  Europe}.
+\newblock Routledge.
+
+\bibitem[\protect\citeauthoryear{Hillier, Leaman, Stansall, and
+  Bedford}{Hillier \textit{et~al.}}{1976}]{hillier1976space}
+Hillier B., Leaman A., Stansall P., Bedford M. (1976).
+\newblock Space syntax.
+\newblock {\em Environment and Planning B: Planning and design\/}~{\em 3\/}(2),
+  147--185.
+
+\bibitem[\protect\citeauthoryear{Huynh, Makarov, Legara, Monterola, and
+  Chew}{Huynh \textit{et~al.}}{2018}]{huynh2018characterisation}
+Huynh H.~N., Makarov E., Legara E.~F., Monterola C., Chew L.~Y. (2018).
+\newblock Characterisation and comparison of spatial patterns in urban systems:
+  A case study of us cities.
+\newblock {\em Journal of computational science\/}~{\em 24}, 34--43.
+
+\bibitem[\protect\citeauthoryear{Janssens-Maenhout, Crippa, Guizzardi, Muntean,
+  Schaaf, Dentener, Bergamaschi, Pagliari, Olivier, Peters,
+  \textit{et~al.}}{Janssens-Maenhout \textit{et~al.}}{2017}]{janssens2017edgar}
+Janssens-Maenhout G., Crippa M., Guizzardi D., Muntean M., Schaaf E., Dentener
+  F., Bergamaschi P., Pagliari V., Olivier J., Peters J., \textit{et~al.}
+  (2017).
+\newblock Edgar v4. 3.2 global atlas of the three major greenhouse gas
+  emissions for the period 1970--2012.
+\newblock {\em Earth Syst. Sci. Data Discuss\/}.
+
+\bibitem[\protect\citeauthoryear{Komiyama and Takeuchi}{Komiyama and
+  Takeuchi}{2006}]{komiyama2006sustainability}
+Komiyama H., Takeuchi K. (2006).
+\newblock {\em Sustainability science: building a new discipline}.
+
+\bibitem[\protect\citeauthoryear{Lagesse, Bordin, and Douady}{Lagesse
+  \textit{et~al.}}{2015}]{lagesse2015spatial}
+Lagesse C., Bordin P., Douady S. (2015).
+\newblock A spatial multi-scale object to analyze road networks.
+\newblock {\em Network Science\/}~{\em 3\/}(1), 156--181.
+
+\bibitem[\protect\citeauthoryear{Lashof and Ahuja}{Lashof and
+  Ahuja}{1990}]{lashof1990relative}
+Lashof D.~A., Ahuja D.~R. (1990).
+\newblock Relative contributions of greenhouse gas emissions to global warming.
+\newblock {\em Nature\/}~{\em 344\/}(6266), 529.
+
+\bibitem[\protect\citeauthoryear{Le~N{\'e}chet}{Le~N{\'e}chet}{2017}]{lenechet2017peupler}
+Le~N{\'e}chet F. (2017).
+\newblock De l'{\'e}talement urbain aux r{\'e}gions m{\'e}tropolitaines
+  polycentriques : formes de fonctionnement et formes de gouvernance.
+\newblock In {\em {Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
+  m{\'e}tropoles}}. Presses Universitaires Francois Rabelais.
+
+\bibitem[\protect\citeauthoryear{Louf and Barthelemy}{Louf and
+  Barthelemy}{2013}]{louf2013modeling}
+Louf R., Barthelemy M. (2013).
+\newblock Modeling the polycentric transition of cities.
+\newblock {\em Physical review letters\/}~{\em 111\/}(19), 198702.
+
+\bibitem[\protect\citeauthoryear{Makse, Andrade, Batty, Havlin, Stanley,
+  \textit{et~al.}}{Makse \textit{et~al.}}{1998}]{makse1998modeling}
+Makse H.~A., Andrade J.~S., Batty M., Havlin S., Stanley H.~E., \textit{et~al.}
+  (1998).
+\newblock Modeling urban growth patterns with correlated percolation.
+\newblock {\em Physical Review E\/}~{\em 58\/}(6), 7054.
+
+\bibitem[\protect\citeauthoryear{McNally}{McNally}{2000}]{mcnally2000four}
+McNally M.~G. (2000).
+\newblock The four step model.
+
+\bibitem[\protect\citeauthoryear{Newman and Watts}{Newman and
+  Watts}{1999}]{newman1999scaling}
+Newman M.~E., Watts D.~J. (1999).
+\newblock Scaling and percolation in the small-world network model.
+\newblock {\em Physical review E\/}~{\em 60\/}(6), 7332.
+
+\bibitem[\protect\citeauthoryear{Penrose}{Penrose}{1999}]{penrose1999k}
+Penrose M.~D. (1999).
+\newblock On k-connectivity for a geometric random graph.
+\newblock {\em Random Structures \& Algorithms\/}~{\em 15\/}(2), 145--164.
+
+\bibitem[\protect\citeauthoryear{Perez, Banos, and Pettit}{Perez
+  \textit{et~al.}}{2016}]{perez2016agent}
+Perez P., Banos A., Pettit C. (2016).
+\newblock Agent-based modelling for urban planning current limitations and
+  future trends.
+\newblock In {\em International Workshop on Agent Based Modelling of Urban
+  Systems}, pp.\  60--69. Springer.
+
+\bibitem[\protect\citeauthoryear{Piovani, Molinero, and Wilson}{Piovani
+  \textit{et~al.}}{2017}]{piovani2017urban}
+Piovani D., Molinero C., Wilson A. (2017).
+\newblock Urban retail location: insights from percolation theory and spatial
+  interaction modeling.
+\newblock {\em PloS one\/}~{\em 12\/}(10), e0185787.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018a}]{raimbault2018calibration}
+Raimbault J. (2018a).
+\newblock Calibration of a density-based model of urban morphogenesis.
+\newblock {\em PloS one\/}~{\em 13\/}(9), e0203516.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018b}]{raimbault2018caracterisation}
+Raimbault J. (2018b).
+\newblock {\em Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution
+  des r{\'e}seaux de transport et des territoires}.
+\newblock Ph.\ D. thesis, Universit{\'e} Paris 7 Denis Diderot.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018c}]{raimbault2018indirect}
+Raimbault J. (2018c).
+\newblock Indirect evidence of network effects in a system of cities.
+\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
+  2399808318774335.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2019}]{raimbault2018urban}
+Raimbault J. (2019).
+\newblock An urban morphogenesis model capturing interactions between networks
+  and territories.
+\newblock In {\em The Mathematics of Urban Morphology, D'acci L., ed., Springer
+  Birkhauser Mathematics}.
+
+\bibitem[\protect\citeauthoryear{Reuillon, Leclaire, and
+  Rey-Coyrehourcq}{Reuillon \textit{et~al.}}{2013}]{reuillon2013openmole}
+Reuillon R., Leclaire M., Rey-Coyrehourcq S. (2013).
+\newblock Openmole, a workflow engine specifically tailored for the distributed
+  exploration of simulation models.
+\newblock {\em Future Generation Computer Systems\/}~{\em 29\/}(8), 1981--1990.
+
+\bibitem[\protect\citeauthoryear{Sanders}{Sanders}{2017}]{sanders2017peupler}
+Sanders L. (2017).
+\newblock {\em Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
+  m{\'e}tropoles.}
+\newblock Presses Universitaires Francois Rabelais.
+
+\bibitem[\protect\citeauthoryear{Son, Bizhani, Christensen, Grassberger, and
+  Paczuski}{Son \textit{et~al.}}{2012}]{son2012percolation}
+Son S.-W., Bizhani G., Christensen C., Grassberger P., Paczuski M. (2012).
+\newblock Percolation theory on interdependent networks based on epidemic
+  spreading.
+\newblock {\em EPL (Europhysics Letters)\/}~{\em 97\/}(1), 16006.
+
+\bibitem[\protect\citeauthoryear{Stauffer and Aharony}{Stauffer and
+  Aharony}{2014}]{stauffer2014introduction}
+Stauffer D., Aharony A. (2014).
+\newblock {\em Introduction to percolation theory}.
+\newblock Taylor \& Francis.
+
+\bibitem[\protect\citeauthoryear{Tammaru, Marci{\'n}czak, Van~Ham, and
+  Musterd}{Tammaru \textit{et~al.}}{2015}]{tammaru2015multi}
+Tammaru T., Marci{\'n}czak S., Van~Ham M., Musterd S. (2015).
+\newblock A multi-factor approach to understanding socio-economic segregation
+  in european capital cities.
+\newblock In {\em Socio-economic segregation in European capital cities}, pp.\
+  25--53. Routledge.
+
+\bibitem[\protect\citeauthoryear{Tobler}{Tobler}{2004}]{tobler2004first}
+Tobler W. (2004).
+\newblock On the first law of geography: A reply.
+\newblock {\em Annals of the Association of American Geographers\/}~{\em
+  94\/}(2), 304--310.
+
+\bibitem[\protect\citeauthoryear{Vigui{\'e} and Hallegatte}{Vigui{\'e} and
+  Hallegatte}{2012}]{viguie2012trade}
+Vigui{\'e} V., Hallegatte S. (2012).
+\newblock Trade-offs and synergies in urban climate policies.
+\newblock {\em Nature Climate Change\/}~{\em 2\/}(5), 334.
+
+\end{thebibliography}
 

### 2019-03-17 Docs/Papers/Percolation/paper/paper.tex
paper_submitted

--- 2019-03-17 Docs/Papers/Percolation/arxiv/v1/paper.tex
@@
-\documentclass{article}
+\documentclass{jimis-en}
 \pdfoutput=1 %forces arXiv/HAL to compile with PDFLaTeX
@@
 \usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[margin=2.4cm]{geometry}
-
-\begin{document}
 
@@
 
-\author{Juste RAIMBAULT$^{1,2,3,\ast}$\\
-$^1$ UPS CNRS 3611 ISC-PIF, France\\
-$^2$ CASA, UCL, UK\\
-$^3$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\\medskip
-$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
-}
+\author[1,2,3,*]{Juste RAIMBAULT}
+\affil[1]{UPS CNRS 3611 ISC-PIF, France} 
+\affil[2]{CASA, UCL, UK}
+\affil[3]{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s}
 
-\date{}
+% corresponding author with his/her email
+\corrauthor{juste.raimbault@polytechnique.edu}
+
+%% Information provided by JIMIS:
+%ISSN
+% DOI 
+\doi{10.18713/JIMIS-ddmmyy-v-a}{}
+% reviewing process: first date = submission, second date = acceptance 
+\review{Month-in-letters Day Year}{Month-in-letters Day Year}
+% Volume and Year of the issue
+\publication{N}{YYYY}
+
+%% Information provided by the Editors:
+% Issue Title 
+\issue{Title of the interdisciplinary issue}
+% List of k Editors
+\editors{}
+
+\begin{document}
 
@@
 The variable parametrization allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions).
-This suggests a customizable spatial design of policies to develop sustainable territories.\\\medskip
-\textbf{Keywords: } Road network; Multi-dimensional percolation; European urban system; Mega-city region
-}
+This suggests a customizable spatial design of policies to develop sustainable territories.}
+
+\keywords{Road network; Multi-dimensional percolation; European urban system; Mega-city region}
 
 \section{Introduction}
+
+\strut
+\vspace{-4ex}
 
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{principle.pdf}}
+  {\includegraphics[width=\linewidth]{figures/principle.pdf}}
   \centering
@@
 In practice, the analysis is implemented using R and the igraph package.
-Source code, data and results are available on the open git repository of the project at \texttt{https://github.com/JusteRaimbault/UrbanMorphology}.
+Source code, data and results are available on the open git repository of the project at \url{https://github.com/JusteRaimbault/UrbanMorphology}.
 The network is constructed by superposing the population density layer with the network layer, starting from the 5km resolution spatial fields for morphological and network indicators.
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.49\linewidth]{abssize_nodes.png}}
-  {\includegraphics[width=0.49\linewidth]{relsize_nodes.png}}
+  {\includegraphics[width=0.49\linewidth]{figures/abssize_nodes.png}}
+  {\includegraphics[width=0.49\linewidth]{figures/relsize_nodes.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{fractaldimension.png}}
+  {\includegraphics[width=\linewidth]{figures/fractaldimension.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-{\includegraphics[width=0.49\textwidth]{totalPop4183694_00056402_ecount850_radius8000.png}}
-  {\includegraphics[width=0.49\linewidth]{totalPop2219780_36719597_vcount378_radius8000.png}}\\
-  {\includegraphics[width=0.49\linewidth]{totalPop1474347_36891685_vcount595_radius50000.png}}
-   {\includegraphics[width=0.49\linewidth]{totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
+{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}
+  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}\\
+  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
+   {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.7\linewidth]{full_effective_pareto.png}}
+  {\includegraphics[width=0.7\linewidth]{figures/full_effective_pareto.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{aggreg_morpho_pc1-emissions_targeted.png}}
+  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_pc1-emissions_targeted.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
+  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
   \centering  
@@
 
-\begin{thebibliography}{}
-
-\bibitem[\protect\citeauthoryear{Arbabi, Mayfield, and McCann}{Arbabi
-  \textit{et~al.}}{2019}]{arbabi2019development}
-Arbabi H., Mayfield M., McCann P. (2019).
-\newblock On the development logic of city-regions: inter-versus intra-city
-  mobility in england and wales.
-\newblock {\em Spatial Economic Analysis\/}, 1--20.
-
-\bibitem[\protect\citeauthoryear{Arcaute, Molinero, Hatna, Murcio, Vargas-Ruiz,
-  Masucci, and Batty}{Arcaute \textit{et~al.}}{2016}]{arcaute2016cities}
-Arcaute E., Molinero C., Hatna E., Murcio R., Vargas-Ruiz C., Masucci A.~P.,
-  Batty M. (2016).
-\newblock Cities and regions in britain through hierarchical percolation.
-\newblock {\em Royal Society open science\/}~{\em 3\/}(4), 150691.
-
-\bibitem[\protect\citeauthoryear{Banos and Genre-Grandpierre}{Banos and
-  Genre-Grandpierre}{2012}]{banos2012towards}
-Banos A., Genre-Grandpierre C. (2012).
-\newblock Towards new metrics for urban road networks: Some preliminary
-  evidence from agent-based simulations.
-\newblock In {\em Agent-based models of geographical systems}, pp.\  627--641.
-  Springer.
-
-\bibitem[\protect\citeauthoryear{Barth{\'e}lemy}{Barth{\'e}lemy}{2011}]{barthelemy2011spatial}
-Barth{\'e}lemy M. (2011).
-\newblock Spatial networks.
-\newblock {\em Physics Reports\/}~{\em 499\/}(1-3), 1--101.
-
-\bibitem[\protect\citeauthoryear{Batty and Longley}{Batty and
-  Longley}{1994}]{batty1994fractal}
-Batty M., Longley P.~A. (1994).
-\newblock {\em Fractal cities: a geometry of form and function}.
-\newblock Academic press.
-
-\bibitem[\protect\citeauthoryear{Bettencourt, Lobo, Helbing, K{\"u}hnert, and
-  West}{Bettencourt \textit{et~al.}}{2007}]{bettencourt2007growth}
-Bettencourt L.~M., Lobo J., Helbing D., K{\"u}hnert C., West G.~B. (2007).
-\newblock Growth, innovation, scaling, and the pace of life in cities.
-\newblock {\em Proceedings of the national academy of sciences\/}~{\em
-  104\/}(17), 7301--7306.
-
-\bibitem[\protect\citeauthoryear{Boccaletti, Bianconi, Criado, Del~Genio,
-  G{\'o}mez-Gardenes, Romance, Sendina-Nadal, Wang, and Zanin}{Boccaletti
-  \textit{et~al.}}{2014}]{boccaletti2014structure}
-Boccaletti S., Bianconi G., Criado R., Del~Genio C.~I., G{\'o}mez-Gardenes J.,
-  Romance M., Sendina-Nadal I., Wang Z., Zanin M. (2014).
-\newblock The structure and dynamics of multilayer networks.
-\newblock {\em Physics Reports\/}~{\em 544\/}(1), 1--122.
-
-\bibitem[\protect\citeauthoryear{Burger and Meijers}{Burger and
-  Meijers}{2012}]{burger2012form}
-Burger M., Meijers E. (2012).
-\newblock Form follows function? linking morphological and functional
-  polycentricity.
-\newblock {\em Urban studies\/}~{\em 49\/}(5), 1127--1149.
-
-\bibitem[\protect\citeauthoryear{Callaway, Newman, Strogatz, and
-  Watts}{Callaway \textit{et~al.}}{2000}]{callaway2000network}
-Callaway D.~S., Newman M.~E., Strogatz S.~H., Watts D.~J. (2000).
-\newblock Network robustness and fragility: Percolation on random graphs.
-\newblock {\em Physical review letters\/}~{\em 85\/}(25), 5468.
-
-\bibitem[\protect\citeauthoryear{Cottineau, Finance, Hatna, Arcaute, and
-  Batty}{Cottineau \textit{et~al.}}{2018}]{cottineau2018defining}
-Cottineau C., Finance O., Hatna E., Arcaute E., Batty M. (2018).
-\newblock Defining urban clusters to detect agglomeration economies.
-\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
-  2399808318755146.
-
-\bibitem[\protect\citeauthoryear{Cottineau, Hatna, Arcaute, and
-  Batty}{Cottineau \textit{et~al.}}{2017}]{cottineau2017diverse}
-Cottineau C., Hatna E., Arcaute E., Batty M. (2017).
-\newblock Diverse cities or the systematic paradox of urban scaling laws.
-\newblock {\em Computers, environment and urban systems\/}~{\em 63}, 80--94.
-
-\bibitem[\protect\citeauthoryear{Crucitti, Latora, and Porta}{Crucitti
-  \textit{et~al.}}{2006}]{crucitti2006centrality}
-Crucitti P., Latora V., Porta S. (2006).
-\newblock Centrality measures in spatial networks of urban streets.
-\newblock {\em Physical Review E\/}~{\em 73\/}(3), 036125.
-
-\bibitem[\protect\citeauthoryear{Ducruet and Beauguitte}{Ducruet and
-  Beauguitte}{2014}]{ducruet2014spatial}
-Ducruet C., Beauguitte L. (2014).
-\newblock Spatial science and network science: Review and outcomes of a complex
-  relationship.
-\newblock {\em Networks and Spatial Economics\/}~{\em 14\/}(3-4), 297--316.
-
-\bibitem[\protect\citeauthoryear{Hackett, Cellai, G{\'o}mez, Arenas, and
-  Gleeson}{Hackett \textit{et~al.}}{2016}]{hackett2016bond}
-Hackett A., Cellai D., G{\'o}mez S., Arenas A., Gleeson J.~P. (2016).
-\newblock Bond percolation on multiplex networks.
-\newblock {\em Physical Review X\/}~{\em 6\/}(2), 021002.
-
-\bibitem[\protect\citeauthoryear{Hall and Pain}{Hall and
-  Pain}{2006}]{hall2006polycentric}
-Hall P.~G., Pain K. (2006).
-\newblock {\em The polycentric metropolis: learning from mega-city regions in
-  Europe}.
-\newblock Routledge.
-
-\bibitem[\protect\citeauthoryear{Hillier, Leaman, Stansall, and
-  Bedford}{Hillier \textit{et~al.}}{1976}]{hillier1976space}
-Hillier B., Leaman A., Stansall P., Bedford M. (1976).
-\newblock Space syntax.
-\newblock {\em Environment and Planning B: Planning and design\/}~{\em 3\/}(2),
-  147--185.
-
-\bibitem[\protect\citeauthoryear{Huynh, Makarov, Legara, Monterola, and
-  Chew}{Huynh \textit{et~al.}}{2018}]{huynh2018characterisation}
-Huynh H.~N., Makarov E., Legara E.~F., Monterola C., Chew L.~Y. (2018).
-\newblock Characterisation and comparison of spatial patterns in urban systems:
-  A case study of us cities.
-\newblock {\em Journal of computational science\/}~{\em 24}, 34--43.
-
-\bibitem[\protect\citeauthoryear{Janssens-Maenhout, Crippa, Guizzardi, Muntean,
-  Schaaf, Dentener, Bergamaschi, Pagliari, Olivier, Peters,
-  \textit{et~al.}}{Janssens-Maenhout \textit{et~al.}}{2017}]{janssens2017edgar}
-Janssens-Maenhout G., Crippa M., Guizzardi D., Muntean M., Schaaf E., Dentener
-  F., Bergamaschi P., Pagliari V., Olivier J., Peters J., \textit{et~al.}
-  (2017).
-\newblock Edgar v4. 3.2 global atlas of the three major greenhouse gas
-  emissions for the period 1970--2012.
-\newblock {\em Earth Syst. Sci. Data Discuss\/}.
-
-\bibitem[\protect\citeauthoryear{Komiyama and Takeuchi}{Komiyama and
-  Takeuchi}{2006}]{komiyama2006sustainability}
-Komiyama H., Takeuchi K. (2006).
-\newblock {\em Sustainability science: building a new discipline}.
-
-\bibitem[\protect\citeauthoryear{Lagesse, Bordin, and Douady}{Lagesse
-  \textit{et~al.}}{2015}]{lagesse2015spatial}
-Lagesse C., Bordin P., Douady S. (2015).
-\newblock A spatial multi-scale object to analyze road networks.
-\newblock {\em Network Science\/}~{\em 3\/}(1), 156--181.
-
-\bibitem[\protect\citeauthoryear{Lashof and Ahuja}{Lashof and
-  Ahuja}{1990}]{lashof1990relative}
-Lashof D.~A., Ahuja D.~R. (1990).
-\newblock Relative contributions of greenhouse gas emissions to global warming.
-\newblock {\em Nature\/}~{\em 344\/}(6266), 529.
-
-\bibitem[\protect\citeauthoryear{Le~N{\'e}chet}{Le~N{\'e}chet}{2017}]{lenechet2017peupler}
-Le~N{\'e}chet F. (2017).
-\newblock De l'{\'e}talement urbain aux r{\'e}gions m{\'e}tropolitaines
-  polycentriques : formes de fonctionnement et formes de gouvernance.
-\newblock In {\em {Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
-  m{\'e}tropoles}}. Presses Universitaires Francois Rabelais.
-
-\bibitem[\protect\citeauthoryear{Louf and Barthelemy}{Louf and
-  Barthelemy}{2013}]{louf2013modeling}
-Louf R., Barthelemy M. (2013).
-\newblock Modeling the polycentric transition of cities.
-\newblock {\em Physical review letters\/}~{\em 111\/}(19), 198702.
-
-\bibitem[\protect\citeauthoryear{Makse, Andrade, Batty, Havlin, Stanley,
-  \textit{et~al.}}{Makse \textit{et~al.}}{1998}]{makse1998modeling}
-Makse H.~A., Andrade J.~S., Batty M., Havlin S., Stanley H.~E., \textit{et~al.}
-  (1998).
-\newblock Modeling urban growth patterns with correlated percolation.
-\newblock {\em Physical Review E\/}~{\em 58\/}(6), 7054.
-
-\bibitem[\protect\citeauthoryear{McNally}{McNally}{2000}]{mcnally2000four}
-McNally M.~G. (2000).
-\newblock The four step model.
-
-\bibitem[\protect\citeauthoryear{Newman and Watts}{Newman and
-  Watts}{1999}]{newman1999scaling}
-Newman M.~E., Watts D.~J. (1999).
-\newblock Scaling and percolation in the small-world network model.
-\newblock {\em Physical review E\/}~{\em 60\/}(6), 7332.
-
-\bibitem[\protect\citeauthoryear{Penrose}{Penrose}{1999}]{penrose1999k}
-Penrose M.~D. (1999).
-\newblock On k-connectivity for a geometric random graph.
-\newblock {\em Random Structures \& Algorithms\/}~{\em 15\/}(2), 145--164.
-
-\bibitem[\protect\citeauthoryear{Perez, Banos, and Pettit}{Perez
-  \textit{et~al.}}{2016}]{perez2016agent}
-Perez P., Banos A., Pettit C. (2016).
-\newblock Agent-based modelling for urban planning current limitations and
-  future trends.
-\newblock In {\em International Workshop on Agent Based Modelling of Urban
-  Systems}, pp.\  60--69. Springer.
-
-\bibitem[\protect\citeauthoryear{Piovani, Molinero, and Wilson}{Piovani
-  \textit{et~al.}}{2017}]{piovani2017urban}
-Piovani D., Molinero C., Wilson A. (2017).
-\newblock Urban retail location: insights from percolation theory and spatial
-  interaction modeling.
-\newblock {\em PloS one\/}~{\em 12\/}(10), e0185787.
-
-\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018a}]{raimbault2018calibration}
-Raimbault J. (2018a).
-\newblock Calibration of a density-based model of urban morphogenesis.
-\newblock {\em PloS one\/}~{\em 13\/}(9), e0203516.
-
-\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018b}]{raimbault2018caracterisation}
-Raimbault J. (2018b).
-\newblock {\em Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution
-  des r{\'e}seaux de transport et des territoires}.
-\newblock Ph.\ D. thesis, Universit{\'e} Paris 7 Denis Diderot.
-
-\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018c}]{raimbault2018indirect}
-Raimbault J. (2018c).
-\newblock Indirect evidence of network effects in a system of cities.
-\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
-  2399808318774335.
-
-\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2019}]{raimbault2018urban}
-Raimbault J. (2019).
-\newblock An urban morphogenesis model capturing interactions between networks
-  and territories.
-\newblock In {\em The Mathematics of Urban Morphology, D'acci L., ed., Springer
-  Birkhauser Mathematics}.
-
-\bibitem[\protect\citeauthoryear{Reuillon, Leclaire, and
-  Rey-Coyrehourcq}{Reuillon \textit{et~al.}}{2013}]{reuillon2013openmole}
-Reuillon R., Leclaire M., Rey-Coyrehourcq S. (2013).
-\newblock Openmole, a workflow engine specifically tailored for the distributed
-  exploration of simulation models.
-\newblock {\em Future Generation Computer Systems\/}~{\em 29\/}(8), 1981--1990.
-
-\bibitem[\protect\citeauthoryear{Sanders}{Sanders}{2017}]{sanders2017peupler}
-Sanders L. (2017).
-\newblock {\em Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
-  m{\'e}tropoles.}
-\newblock Presses Universitaires Francois Rabelais.
-
-\bibitem[\protect\citeauthoryear{Son, Bizhani, Christensen, Grassberger, and
-  Paczuski}{Son \textit{et~al.}}{2012}]{son2012percolation}
-Son S.-W., Bizhani G., Christensen C., Grassberger P., Paczuski M. (2012).
-\newblock Percolation theory on interdependent networks based on epidemic
-  spreading.
-\newblock {\em EPL (Europhysics Letters)\/}~{\em 97\/}(1), 16006.
-
-\bibitem[\protect\citeauthoryear{Stauffer and Aharony}{Stauffer and
-  Aharony}{2014}]{stauffer2014introduction}
-Stauffer D., Aharony A. (2014).
-\newblock {\em Introduction to percolation theory}.
-\newblock Taylor \& Francis.
-
-\bibitem[\protect\citeauthoryear{Tammaru, Marci{\'n}czak, Van~Ham, and
-  Musterd}{Tammaru \textit{et~al.}}{2015}]{tammaru2015multi}
-Tammaru T., Marci{\'n}czak S., Van~Ham M., Musterd S. (2015).
-\newblock A multi-factor approach to understanding socio-economic segregation
-  in european capital cities.
-\newblock In {\em Socio-economic segregation in European capital cities}, pp.\
-  25--53. Routledge.
-
-\bibitem[\protect\citeauthoryear{Tobler}{Tobler}{2004}]{tobler2004first}
-Tobler W. (2004).
-\newblock On the first law of geography: A reply.
-\newblock {\em Annals of the Association of American Geographers\/}~{\em
-  94\/}(2), 304--310.
-
-\bibitem[\protect\citeauthoryear{Vigui{\'e} and Hallegatte}{Vigui{\'e} and
-  Hallegatte}{2012}]{viguie2012trade}
-Vigui{\'e} V., Hallegatte S. (2012).
-\newblock Trade-offs and synergies in urban climate policies.
-\newblock {\em Nature Climate Change\/}~{\em 2\/}(5), 334.
-
-\end{thebibliography}
+\bibliographystyle{jimis-en}
+\bibliography{biblio}
 

### 2019-09-11 Docs/Papers/Percolation/paper/Response_to_Reviewers.tex
todo_revision_jimis

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\documentclass[11pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

\usepackage[document]{ragged2e}
% pour justifier

% moderncv themes
\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
\renewcommand{\familydefault}{\rmdefault} 
%\nopagenumbers{} 
\usepackage[utf8]{inputenc} 
\usepackage[scale=0.9]{geometry}

\def \draft {1}
\usepackage{xparse}
\usepackage{ifthen}
\DeclareDocumentCommand{\comment}{o m o o o o}
{\ifthenelse{\draft=1}{
  \IfValueT{#1}{
      \textcolor{red}{\textbf{C (#1) : }#2}
      \IfValueT{#3}{\textcolor{blue}{\textbf{A1 : }#3}}
      \IfValueT{#4}{\textcolor{green}{\textbf{A2 : }#4}}
      \IfValueT{#5}{\textcolor{red!50!blue}{\textbf{A3 : }#5}}
      \IfValueT{#6}{\textcolor{blue}{\textbf{A4 : }#6}}
    }
    \IfNoValueT{#1}{
      \textcolor{red}{\textbf{C : }#2}
      \IfValueT{#3}{\textcolor{blue}{\textbf{A1 : }#3}}
      \IfValueT{#4}{\textcolor{green}{\textbf{A2 : }#4}}
      \IfValueT{#5}{\textcolor{red!50!blue}{\textbf{A3 : }#5}}
      \IfValueT{#6}{\textcolor{blue}{\textbf{A4 : }#6}}
    }
 }{}
}

\firstname{}
\lastname{}
\begin{document}

% recipient data
\recipient{Editor JASSS}{}
\date{\today}
\opening{Dear Editor,}
\closing{Yours faithfully,\\
Juste Raimbault%\\
%Université Paris 7 - UMR CNRS 8504 Géographie-cités
}
         % use an optional argument to use a string other than "Enclosure", or redefine \enclname

%\makelettertitle

\justify

\textbf{Response to reviewers}

\bigskip

\bigskip

\textbf{Reviewer 1:}

\begin{enumerate}

%I have carefully read the manuscript written by Raimbault entitled as “Multi-dimensional Urban Network Percolation”, which was submitted to JIMIS for publication. This paper addresses a multi-dimensional network representation of urban population density and transportation networks to study their correlations and percolation properties over Europe. Beyond the observations of the percolation transition by varying the percolation radius, the author addresses the organisation of poly- and multicetered mega city regions from a sustainability point of view.
%Considering that I am not an expert of urban development, but having background in network science, as far as I can judge the methodology, statistical data analysis, argumentation, and synthesis presented in this manuscript are correct and the language also reaches the scientific standard. However, I have a few comments what I would like to suggest to the author to address as it would largely improve the readability of the manuscript. 
%Comments:

\item \textit{Although the focus of this work is about network percolation, the definition of multidimensional (multilayer) network is poorly explained. I understood that in both layers nodes might be spatial grid bins, and that links in one layer might be defined via population density and in the other layer might be defined by transportation. But the manuscript does not explain clearly how the networks were constructed, i.e. the definition of nodes, and links and the matching between different layers. Do you use the gravity model to define links between locations? Do you use a disk percolation kind of approach? This should be defined in Section 2.2 but the description there is very cryptic.}

$\rightarrow$ The methodological section describing the percolation heuristic was reorganized and rewritten to be clearer. In particular, the general part for any network was made explicitly separate from the construction of the gridded network and the construction of nodes, and also from the application to the empirical dataset of population density and network indicators. The construction of links within each layer, done through the distance between nodes (corresponding to disk percolation) and the values of the variable corresponding to the layer, was also clarified.

\medskip

\item \textit{At the same time I could not understand what a network variable is, also mentioned in 2.2. It is a too general term as there are many variables of a network.}

$\rightarrow$ This was also clarified with the description of the heuristic and network construction. In particular, ``network variable'' are network indicators computed on the road network on a spatial window centered around the node. This term was preferentially used.

\medskip

\item \textit{Considering that JIMIS is an interdisciplinary journal the author assumes too much about the reader. Terms like cyclomatic number, euclidian efficiency, or Pareto front should be defined in the manuscript.}

$\rightarrow$ Definition of terms were added.

\medskip

\item \textit{The author misses to describe in details how he obtained and pre-processed the data he uses to construct the network and other aspects of the study. Although he shares the data this is an important question from the point of credibility, methodology and reproducibility.}

$\rightarrow$ Data preprocessing and network construction were described with more details.

\medskip

\item \textit{Typos \ldots}

$\rightarrow$ Typos were corrected
%- Page 1 Section 1: “These study” -> “This study”
%- Page 4 Section 1: “interacts is they are” -> “interacts if they are”

\end{enumerate}

\bigskip

\textbf{Reviewer 2:}

\medskip

\begin{enumerate}
%This paper proposes to apply multilayer percolation to the European urban system. It applies a known method to a new field (urban systems). This idea is quite interesting and has a great potential.
\item \textit{Nevertheless, to me, the main drawback of this paper is that the research question is unclear. What is leading the methodology? The aim of mobilized indicators (economic performance and greenhouse gas emissions) is also not straightforward.}

$\rightarrow$ The research question of the paper was formulated in a clearer way in the introduction. It consists in two main axis: (i) from the methodological viewpoint, how can urban network percolation be extended to multiple dimensions; and (ii) from the thematic viewpoint, how can endogenous mega urban regions be characterized which such a method, and how can it provide insights into their sustainability. These two axis are separated, but still tightly related as the second depends on the new methodology, and the first is illustrated through its concrete application in the second. The choice of sustainability indicators was also better motivated.

\medskip

\item \textit{Several key concepts and main points of the methodology are mentioned as the text goes by, without having been introduced earlier. Thus, a great number of formulas and parameters are gathered, without clearly explaining the articulation between all of them, and the purpose they serve.}

$\rightarrow$ The content of the methodology was better presented and contextualized in the introduction, and formulas and parameters were described with more details.

\medskip

\item \textit{The discussion and conclusion are very expeditious and reinforce the idea of a methodology without a research question to lean on.}

$\rightarrow$ The discussion was extended and made more precise, with a stronger relation with the research question.

\medskip

%Details:
%– INTRODUCTION –

\item \textit{Other quantitative methods able to bring information linked to the question leading the work, and how the method exposed here can bring new elements, should be more developed. This part needs a more substantial structure and presentation.}

$\rightarrow$ The introduction was reorganized and more literature was added, both for the methodological part and the thematic part of the research question.

\medskip

\item \textit{The author should make clearer what is taken into account in the urban system. Spatial distribution of population and properties of transportation network are mentioned but need to be specified (which one, defined how?).}

$\rightarrow$ The dimension of the urban system taken into account were better specified in the introduction.

\medskip

%– METHODS – 
%- Multidimensional percolation -

\item \textit{The mathematical aspect of the methodology is a very interesting part of the work. But the only fact of applying a new methodology does not make a research result to me: it is necessary to show how brought information is significant to analyze an urban system.}

$\rightarrow$ The discussion of results was reworked regarding the thematic research question.

\medskip

\item \textit{The figure 1 is unclear; its legend is confused; different concepts are mixed.}

$\rightarrow$ Legend of figure 1 was made clearer.

\medskip

\item \textit{The third paragraph is much clearer than the two first. It may come first. Furthermore, the percolation on one single layer has to be defined before presenting the one on several layers.}

$\rightarrow$ This section was reorganized to be clearer.

\medskip

\item \textit{English seems very ``French'' in the last paragraph.}

$\rightarrow$ This paragraph, and more generally the paper, was edited for English soundness.

\medskip

\item \textit{Empirical data: Here we understand that nodes are centers of cells, and that each layer corresponds to (1) population density (2) network variable (Nv / Ne / cyclomatic number / euclidian efficiency) This clarification should come earlier in the paper, in the methodological part.}

$\rightarrow$ The description of methods and data was reorganized to make this clearer.

\medskip

\item \textit{Besides, several questions need explanations at this point : What is this efficiency chosen? Why those indicators? And still\ldots What is the research question?}

$\rightarrow$ The choice of network indicators was further discussed, in relation with the research question.

\medskip

\item \textit{Sustainability indicators: The ``messy'' way to present formulas (even not really explained sometimes) make their relevance for the development questionable.}

$\rightarrow$ More details and explanations were given for these indicators.

\medskip

\item \textit{``recent forms of urbanization, in particular integrated megacity regions, may imply different patterns of economic and transportation flows and thus exhibit various performances regarding different indicators of sustainability.'' This is one of the strong hypothesis of the paper. It needs to be introduced earlier.}

$\rightarrow$ This assumption, linked to the research question on the sustainability of megacity regions, was reformulated and integrated in the rewritten introduction.

\medskip

\item \textit{Sustainability and greenhouse gas emissions have also to be introduced before: why is the focus here?}

$\rightarrow$ The choice of these dimensions of sustainability is to give an illustration with two paradigmatic opposed dimensions, which can furthermore be approximated indirectly from gravity models as we propose. The explanation of this choice was added in the introduction.

\medskip

%– RESULTS – 
%- Implementation -
%The choice to offer online results is great!

\item \textit{``Thus, it indeed increases with network performance, in consistence with the use done here through thresholding.'' This part is not very clear to me.}

$\rightarrow$ This paragraph was clarified.

\medskip

%- Percolation transition and fractal dimension -

\item \textit{Fractals are introduced just here! It seems to be part of the methodology\ldots Idem for the standard deviation, and the difference $(\alpha - \sigma \left[\alpha\right])_M - (\alpha - \sigma\left[\alpha\right])$: it is not clear why those formulas are added here and what they support.}

$\rightarrow$ The study of fractal dimension is not essential in the methodology, but comes as an auxiliary result to validate the multi-dimensional approach regarding the original papers studying urban percolation. This was specified, and the formulas were better explained.

\medskip

\item \textit{A summary table would have been great here.}

$\rightarrow$

\medskip

\item \textit{Extracting endogenous megacity regions: Why is this configuration working? The impression is that the result comes as the work is felt the way along.}

$\rightarrow$

\medskip

\item \textit{How do we know that it is not just an effect of the population density layer?}

$\rightarrow$

\medskip

\item \textit{``we observe a Pareto front for all points\ldots'': What do we learn about urban system from that?}

$\rightarrow$

\medskip

\item \textit{Linking urban morphology and sustainability: Very confused part.}

$\rightarrow$ This part was clarified by adding more context literature and more explanations.

\medskip

%– DISCUSSION & CONCLUSION –

\item \textit{The discussion and conclusion need to be developed. Before announcing developments (4.1), the reader has to understand what this specific research work brings (what are the main results, linked to the research question, and their limits). The conclusion is very broad (“link between form and function in urban system”; “shows its potentialities”). It needs to be more specific.}

$\rightarrow$ The discussion was extended in relation with answers to the research question.

%From my point of view, this paper could be very interesting with a clarification of the research question and a more structured presentation of methodology and indicators.

\medskip

\item \textit{I am not a native English speaker, but it seems that there are some English points to adjust. Some sentences are too long to be clear.}

$\rightarrow$ English was reworked throughout the paper to make it more readable and understandable.

\end{enumerate}

\end{document}

%% end of file `template.tex'.

### 2019-09-11 Docs/Papers/Percolation/paper/paper.tex
todo_revision_jimis

@@
 
+% TODO English was reworked throughout the paper to make it more readable and understandable.
+
 \section{Introduction}
@@
 
+\subsection{Towards multidimensional urban network percolation}
+
 The structure of road networks can be used as a proxy to understand its past growth dynamics, but also has a significant impact on the future sustainability of territories it irrigates.
 Diverse methods to characterize the structure of spatial networks, and more particularly road networks, have been developed in that context, including classical network indicators such as centralities \citep{crucitti2006centrality} but also more elaborated constructions capturing more realistic processes in terms of street network use \citep{lagesse2015spatial}.
-These study are by essence interdisciplinary, or at least imply complementary viewpoints from disciplines as diverse as architecture with space syntax \citep{hillier1976space}, physics with the study of spatial networks \citep{barthelemy2011spatial}, or social science disciplines concerned with space such as geography \citep{ducruet2014spatial}.
+Such studies of urban networks are by essence interdisciplinary, or at least imply complementary viewpoints from disciplines as diverse as architecture with space syntax \citep{hillier1976space}, physics with the study of spatial networks \citep{barthelemy2011spatial}, or social science disciplines concerned with space such as geography \citep{ducruet2014spatial}.
 
@@
 Important applications include the quantification of network robustness \citep{callaway2000network} or the modeling of epidemic spreading \citep{newman1999scaling}.
+
 Such approaches have been applied to urban systems not only for the study of networks.
@@
 \cite{piovani2017urban} apply road network percolation at the mesoscopic scale of London metropolitan area, in relation with a retail location model.
+At a larger scale, the paradigm of percolation transition has been applied to the study of urban traffic dynamics \citep{Li669,Zeng23}.
 In spatial statistics, this method can be used to characterize the spatial morphology of point patterns \citep{huynh2018characterisation}.
@@
 
-Beside these methodological issues, applied tools are needed to quantify the sustainability of recently emerged urban forms.
+\subsection{Sustainability of mega-urban regions}
+
+Beside these methodological issues of characterizing urban networks and more particularly their endogenous hierarchical structure, some related applied research issues can be considered.
+Indeed, quantitative tools are needed to evaluate the sustainability of recently emerged urban forms.
 In particular, according to \cite{lenechet2017peupler}, the most recent transition of human settlement systems (in the sense of \cite{sanders2017peupler}, i.e.
@@
 The transition imply complex processes such as changes in the governance structure, and can not be associated to the stylized transition identified by \cite{louf2013modeling} in a simplistic toy urban model, and therefore has chances to imply more drivers than negative externalities of congestion only.
+
 To what extent these new urban forms are sustainable, for example in the broad sense of UN development goals \citep{komiyama2006sustainability}, remains an open question.
+Indeed, these integrated mega-city regions may for example imply different patterns of economic and transportation flows and thus exhibit various performances regarding different indicators of sustainability.
+
+Case studies of targeted megacities have for example focused on the links between urban form and mobility or resources management \citep{sorensen2010megacities}.
+A significant literature has focused on the sustainability of mega urban regions with a qualitative approach, such as \cite{laquian2011planning} which establish a typology of these and corresponding governance structures in the Asian context.
+In the case of Europe, \cite{marull2013emerging} use econometric analysis to see the economic and ecological advantage of integrated urban regions.
+\cite{feng2018spatiotemporal} introduce a method to measure the level of polycentricity of an urban mega-region.
+\cite{su2017china} propose a framework to evaluate the performance of urban mega-regions, regarding economic, environmental, social and spatial dimensions.
+
+Beside these targeted studies, a remaining open issue, to the best of our knowledge never tackled at this small geographical scale, is the endogenous characterization of such urban regions.
+The delineation of such geographical systems is taken as exogenous, and their performance and sustainability is then evaluated.
+
+\subsection{Proposed approach}
+
+% research question
+% TODO clear formulation of research question : include territory - network link
+
+This work proposes to partly tackle the two above research questions by linking them.
+More precisely, it first from a methodological viewpoint investigates how urban network percolation can be generalized to multiple dimensions, and secondly explores the endogenous characterization of mega-city regions using such a multi-dimensional percolation.
+%  It consists in two main axis: (i) from the methodological viewpoint, how can urban network percolation be extended to multiple dimensions; and (ii) from the thematic viewpoint, how can endogenous mega urban regions be characterized which such a method, and how can it provide insights into their sustainability.
+These two axis are separated, but still tightly related as the second depends on the new methodology, and the first is illustrated through its concrete application in the second.
+
 In order to test our multi-dimensional percolation method, we propose to apply it to the endogenous characterization of urban regions, and compute stylized sustainability indicators on the constructed regions.
+
+We apply the heuristic to urban morphology and road network topology measures in Europe.
+The idea to combine urban form with network topology measures relies on the capture of the link between urban form and function as already mentioned, urban functions being assumed as distributed by transportation networks \citep{raimbault2018caracterisation}.
+
+% TODO dimension of the urban system taken into account were better specified in the introduction
 
@@
 We finally discuss possible developments and the implications of this methodology to the design of policies.
+
+% TODO add more details on all aspects of methodology
+
+% TODO more details in the text on formulas and parameters
 
@@
 \subsection{Multi-dimensional percolation}
+
+% TODO rewrite legend of fig 1
 
@@
 
+\subsubsection{Proposed heuristic}
+
 Percolation processes in multilayer networks have been proposed as an extension within simple networks \citep{boccaletti2014structure}.
 A generalization of epidemic spreading can for example be achieved using this framework \citep{son2012percolation}.
-In the case of multilayer networks sharing the same nodes for all layers, often called multiplex networks, bond percolation has also been studied \citep{hackett2016bond}.
-
+In the case of multilayer networks sharing the same nodes for all layers, often called multiplex networks \citep{aleta2018multilayer}, bond percolation has also been studied \citep{hackett2016bond}.
 In the case of our heuristic, bond percolation is operated between two cells given a distance threshold, and furthermore with a threshold parameter for each layer assuming a node function within each layer.
@@
 
-More formally, let assume a set of nodes $V = v_i$ common to all layers, and layers edges $E_j$ taken as empty at the initial state of the algorithm.
+Note that we do not call our method ``multi-layer percolation'', as it is not strictly multi-layer since nodes are common.
+The term of multi-dimensional percolation is more suited to the use of multiple variables and thresholds.
+The method works with an arbitrary number of layers.
+See Fig.~\ref{fig:method} for a schematic representation of the method.
+It can be implemented with a propagation heuristic or directly working on adjacency matrices.
+The rationale behind the conjunction of the thresholding of each layer variable and the distance thresholding relies on the idea that two points will interact if they are close enough, but also if they have a strong enough intensity of the activity or dimension captured by each layer, simultaneously for all layers considered.
+This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.
+
+\subsubsection{Formal description}
+
+More formally, let assume a set of nodes $V = v_i$ common to all layers of the network, and layers edges $E_j$ for layer $j$ taken as empty at the initial state of the algorithm.
 Each node has a value of the considered variables associated to each layer, written $v_{ij}$.
@@
 
-Note that we do not call our method ``multi-layer percolation'', as it is not strictly multi-layer since nodes are common.
-The term of multi-dimensional percolation is more suited to the use of multiple variables and thresholds.
-The method works with an arbitrary number of layers.
-See Fig.~\ref{fig:method} for a schematic representation of the method.
-It can be implemented with a propagation heuristic or directly working on adjacency matrices.
-The rationale behind the conjunction of the thresholding of each layer variable and the distance thresholding relies on the idea that two points will interact is they are close enough, but also if they have a strong enough intensity of the activity or dimension captured by each layer, simultaneously for all layers considered.
-This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.
+\subsubsection{Application to gridded networks}
+
+The previous generic method must be applied to a consistent urban network, in the sense of multiple variables associated to spatial nodes.
+To each variable then corresponds a network layer.
+We propose to work on gridded networks, namely nodes regularly spaced in two dimensions.
+In practice, these nodes will be the center of raster cells.
+We consider two layers, one defined by population density and the other by road network structure indicators computed at the same resolution.
+% TODO French-like paragraph
 
@@
 
-We apply the heuristic to urban morphology and road network topology measures in Europe.
-The idea to combine urban form with network topology measures relies on the capture of the link between urban form and function as already mentioned, urban functions being assumed as distributed by transportation networks \citep{raimbault2018caracterisation}.
+We detail now how the empirical layers were computed and the network constructed.
+A grid of population density morphology indicators has been computed on spatial moving windows of width 50km for all European Union by \cite{raimbault2018calibration}, with an offset resolution of 5km.
+From this study we get the aggregated population, producing a raster grid of population with a resolution of 5km.
 
-More precisely, a grid of population density morphology indicators and road network topology indicators has been computed on spatial moving windows of width 50km for all European Union by \cite{raimbault2018urban}, with an offset resolution of 5km.
-We use this data to construct a two layers abstract network: a layer which variable is given by population density, and a second layer which variable is given by a network variable.
+Furthermore, road network topology indicators were computed at a similar resolution by \cite{raimbault2018urban}.
+In practice, (i) the full OpenStreetMap road network for Europe was simplified at the minimal resolution of 100m, keeping the topological properties and link attributes (including maximal speed for example); (ii) for each cell of the population morphology raster, the road network within the same spatial window of width 50km was retrieved; (iii) network structure indicators (summary statistics, centralities, etc.) were computed on this local network, providing a network indicators raster at the same resolution than population.
+
+We use this data to construct a two layers abstract network: a layer which variable is given by population density, and a second layer which variable is given by a road network indicator.
 Nodes are the center of cells (thus disposed in space on a grid of step 5km).
-We test the variable characterizing the second layer among the following characteristics of the road network within the corresponding window: number of edges $N_E$, number of vertices $N_V$, cyclomatic number $\mu$ and euclidian efficiency $v_0$.
+We test several possible networks by varying the road indicator taken into account for the second layer.
+In particular, we test the following indicators:
+\begin{enumerate}
+	\item number of edges $N_E$;
+	\item number of vertices $N_V$;
+	\item cyclomatic number $\mu$ which is defined by $\mu = N_E - N_V + c$ where $c$ is the number of connected components of the graph; this measure captures the number of distinct cycles in the graph
+	\item euclidian efficiency $v_0$, defined by \cite{banos2012towards}
+\end{enumerate}
+
 These measures capture functional properties especially for the two last.
+
+% TODO The choice of network indicators was further discussed, in relation with the research question.
+
+% TODO remove the term network variable ? network indicators
 
@@
 
-As already detailed, recent forms of urbanization, in particular integrated mega-city regions, may imply different patterns of economic and transportation flows and thus exhibit various performances regarding different indicators of sustainability.
+As already detailed, urban regions may perform more or less well regarding different dimensions of sustainability.
 We propose to use the endogenous definition of regional urban systems produced by the percolation algorithm to evaluate their sustainability, in terms of conflicting objectives of economic integration and greenhouse gases emissions.
@@
 We work here with two stylized indicators for two conflicting dimensions, as a proof-of-concept.
+
+% TODO The choice of these dimensions of sustainability is to give an illustration with two paradigmatic opposed dimensions, which can furthermore be approximated indirectly from gravity models as we propose.
+The explanation of this choice was added in the introduction.
 
@@
 The sum of all flows of points within the geographical span of the cluster (that we approximate as the convex Hull envelope of its points), allows us to approximate the cumulated potential emissions and economic activity.
-Writing clusters $K_c$ as this set of points, we define the total economic flow by $E_c = \sum_{i,j \in K_c} \phi_{ij}^{(1)}$ and the total emissions due to flows by $G_c = \sum_{i,j \in C_c} \phi_{ij}^{(2)}$.
-This allows to define a relative economic inefficiency by $e_c = 1 - \frac{\max_c E_c - E_c}{\max_c E_c - \min_c E_c}$ and relative potential emissions by $g_c = \frac{\max_c G_c - G_c}{\max_c G_c - \min_c G_c}$.
+Writing clusters $K_c$ as this set of points, we define the total economic flow by
+
+\begin{equation}
+E_c = \sum_{i,j \in K_c} \phi_{ij}^{(1)}
+\end{equation}
+
+and the total emissions due to flows by
+
+\begin{equation}
+G_c = \sum_{i,j \in C_c} \phi_{ij}^{(2)}
+\end{equation}
+
+This allows defining a relative economic inefficiency by
+
+\begin{equation}
+e_c = 1 - \frac{\max_c E_c - E_c}{\max_c E_c - \min_c E_c}
+\end{equation}
+
+and relative potential emissions by
+
+\begin{equation}
+g_c = \frac{\max_c G_c - G_c}{\max_c G_c - \min_c G_c}
+\end{equation}
+
 Both indicators should be minimized for sustainability.
@@
 Thus, it indeed increases with network performance, in consistence with the use done here through thresholding.
+% TODO clarify
 
@@
 
+% TODO The study of fractal dimension is not essential in the methodology, but comes as an auxiliary result to validate the multi-dimensional approach regarding the original papers studying urban percolation.
+This was specified, and the formulas were better explained.
+
+This study of percolation transition and fractal dimension shows that our multi-dimensional percolation heuristic remains relevant, as results analog to the one-dimensional approach can be obtained.
+
+% TODO table ?
 %  fractdimsignif
@@
 Indeed, regarding the population it contains, an area can be more or less efficient in terms of emissions.
-Seeing the population as an objective to be maximized (thus the plotted value to be minimized), we observe a Pareto front for all points (i.e.
-all clusters across all configurations), but also no dominating point for each configuration.
+Seeing the population as an objective to be maximized (thus the plotted value to be minimized) while the emissions must be minimized, we observe a Pareto front for all points (i.e.
+all clusters across all configurations).
+Given different dimensions to be minimized, a Pareto front consists of the points which are not Pareto-dominated by any other point, i.e.
+that there exists no other point performing best on all objectives.
+In practice, this yields optimization compromises in the context of multi-objective optimization, when no aggregation of the dimensions is possible or desirable.
+In Fig.~\ref{fig:paretos}, the front is the lower bound of the point cloud.
+We also find no dominating point for each configuration.
 Some clusters are therefore optimal compromises in the Pareto sense in each configuration, while some are dominated and thus not efficient.
-
-% not really interesting + messy figure -> do not show this, morphology is more interesting
-%\includegraphics[height=0.8\textheight]{figures/aggreg_paretos_radiuspopthq.png}
 
@@
 \subsection{Linking urban morphology and sustainability}
+
+% TODO This part was clarified by adding more context literature and more explanations.
 
@@
 \section{Discussion}
+
+% TODO extend with relation to research questions
+
+% TODO But the only fact of applying a new methodology does not make a research result to me: it is necessary to show how brought information is significant to analyze an urban system
+
+% TODO The discussion was extended in relation with answers to the research question.
 
@@
 
-% TODO : curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.
+%  curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.
 

### 2019-09-11 Models/MorphologyPackage/morphology.R

@@ -91,6 +91,8 @@ summaryPopulation<-function(m=NULL){
   )
 }
 
+# TODO add effective density
+
 
 
 #moran index 

### 2019-09-11 Docs/Papers/Percolation/paper/Response_to_Reviewers.tex
paper_v2

@@
 
-$\rightarrow$
+$\rightarrow$ We chose not to add a summary table for fractal dimensions, as it is not the main purpose of the study and their variation is furthermore not significant.
 
@@
 
-$\rightarrow$
+$\rightarrow$ This work is exploratory, thus 4800 parameter configuration were explored.
+The ones shown in Fig.
+4 are close to actual functional regions for some, but very far for others.
+This suggest that our parametrizable approach is flexible to consider varying definitions of urban regions, and that in some cases it coincides with the purely functional definition of (Hall and Pain, 2016), i.e.
+that some thresholds of population and road network performance intensity capture actual functional linkages.
+There is however no way to explain this emerging property, as our method constructs clusters from the bottom-up.
+This was added in the discussion.
 
@@
 
-$\rightarrow$
+$\rightarrow$ The shape of percolation transition curves in Fig.
+2 show that the addition of the road network layer adds some information, as the curve for $\theta_N = 0$ (population only) can be compared with other values.
+This was made more explicit.
 
@@
 
-$\rightarrow$
+$\rightarrow$ The presence of Pareto fronts confirms that urban systems consist in general of compromises between multiple objectives. This was clarified.
 

### 2019-09-11 Docs/Papers/Percolation/paper/paper.tex
paper_v2

@@
 
-\abstract{Network percolation has recently been proposed as a method to characterize the global structure of an urban system form the bottom-up.
+\abstract{Network percolation has recently been proposed as a method to characterize the hierarchical structure of an urban system form the bottom-up.
 This paper proposes to extend urban network percolation in a multi-dimensional way, to take into account both urban form (spatial distribution of population) and urban functions (here as properties of transportation networks).
@@
 \keywords{Road network; Multi-dimensional percolation; European urban system; Mega-city region}
-
-% TODO English was reworked throughout the paper to make it more readable and understandable.
 
@@
 The structure of road networks can be used as a proxy to understand its past growth dynamics, but also has a significant impact on the future sustainability of territories it irrigates.
-Diverse methods to characterize the structure of spatial networks, and more particularly road networks, have been developed in that context, including classical network indicators such as centralities \citep{crucitti2006centrality} but also more elaborated constructions capturing more realistic processes in terms of street network use \citep{lagesse2015spatial}.
-Such studies of urban networks are by essence interdisciplinary, or at least imply complementary viewpoints from disciplines as diverse as architecture with space syntax \citep{hillier1976space}, physics with the study of spatial networks \citep{barthelemy2011spatial}, or social science disciplines concerned with space such as geography \citep{ducruet2014spatial}.
+Diverse methods to characterize the structure of spatial networks, and more particularly road networks, have been developed in that context.
+They include classical network indicators such as centralities \citep{crucitti2006centrality} but also more elaborated constructions capturing more realistic processes in terms of street network use \citep{lagesse2015spatial}.
+Such studies of urban networks are by essence interdisciplinary, or at least imply complementary viewpoints from diverse disciplines.
+These for example include architecture with space syntax \citep{hillier1976space}, physics with the study of spatial networks \citep{barthelemy2011spatial}, or social science disciplines concerned with space such as geography \citep{ducruet2014spatial}.
 
-A method to characterize topologies of such urban spatial networks is network percolation, initially applied to road networks by \cite{arcaute2016cities}.
-Percolation in physics can be understood in a broad sense as processes related to the progressive occupation or connection of nodes of a network, and is generally associated to a phase transition with the emergence of a giant cluster at a given connection probability \citep{stauffer2014introduction}.
-Important applications include the quantification of network robustness \citep{callaway2000network} or the modeling of epidemic spreading \citep{newman1999scaling}.
+A method to characterize the hierarchical structure of such urban spatial networks is network percolation, initially applied to urban road networks by \cite{arcaute2016cities}.
+Percolation in physics can be understood in a broad sense as processes related to the progressive occupation or connection of nodes of a network.
+It is generally associated to a phase transition with the emergence of a giant cluster at a given probability of connection \citep{stauffer2014introduction}.
+Practical applications in different fields include the quantification of network robustness \citep{callaway2000network} or the modeling of epidemic spreading \citep{newman1999scaling}.
 
-Such approaches have been applied to urban systems not only for the study of networks.
+Such approaches have been applied to urban systems with other applications than the study of networks.
 \cite{makse1998modeling} model urban growth with a local percolation model for site occupancy.
@@
 
-Existing heuristics however generally focus on a single dimension or property of the urban system.
-However, such systems are known to be multidimensional, for example through the superposition of the morphological dimension of networks, and the functional properties of the urban environment \citep{burger2012form}.
-The link between urban form and function remains in particular an open question \citep{batty1994fractal}, but more generally the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}.
+Existing heuristics generally focus on a single dimension or property of the urban system.
+However, such systems are known to be highly multidimensional.
+For example, the morphological dimension of networks is complementary to the functional properties of the urban environment \citep{burger2012form}.
+The link between urban form and function remains in particular an open question \citep{batty1994fractal}.
+More generally, the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}.
 This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic.
-The method allows to combine different dimensions of the urban system, the same way that \cite{cottineau2018defining} combines population density and commuting flows to produce multiple definitions of urban areas.
+The method allows combining different dimensions of the urban system, the same way that \cite{cottineau2018defining} link population density with commuting flows to produce multiple definitions of urban areas.
 
@@
 These have been defined by \cite{hall2006polycentric} as polycentric urban structures highly integrated in terms of flows.
-The transition imply complex processes such as changes in the governance structure, and can not be associated to the stylized transition identified by \cite{louf2013modeling} in a simplistic toy urban model, and therefore has chances to imply more drivers than negative externalities of congestion only.
+The transition imply complex processes such as changes in the governance structure, and can not be associated to the stylized transition identified by \cite{louf2013modeling} in a toy urban model including negative externalities of congestion only.
 
@@
 Beside these targeted studies, a remaining open issue, to the best of our knowledge never tackled at this small geographical scale, is the endogenous characterization of such urban regions.
-The delineation of such geographical systems is taken as exogenous, and their performance and sustainability is then evaluated.
+The delineation of these geographical systems is often taken as exogenous, and their performance and sustainability is then evaluated.
+Several approaches can be taken to define such systems, including integration through transportation networks, continuity of night lights, and economic productivity thresholds~\citep{lang2005beyond,florida2008rise}.
+In these existing studies, parameters to define such regions remain fixed, and an endogenous hierarchical structure is possibly ignored.
+The second component of our research question will thus be the endogenous characterization of urban regions and how this can be applied to study their sustainability.
 
@@
 
-% research question
-% TODO clear formulation of research question : include territory - network link
+This work proposes to partly tackle the two above research questions by linking them.
+More precisely, it first investigates from a methodological viewpoint how urban network percolation can be generalized to multiple dimensions, and secondly explores the endogenous characterization of mega-city regions using such a multi-dimensional percolation, and how this can be used to quantify the sustainability of these systems.
+These two axis are tightly linked.
+Indeed, on the one hand street network percolation has initially been proposed to identify endogenous entities in territorial systems, and on the other hand mega-urban regions are characterized simultaneously by morphological dimensions (continuity of the built environment) and functional dimensions (high integration of flows).
 
-This work proposes to partly tackle the two above research questions by linking them.
-More precisely, it first from a methodological viewpoint investigates how urban network percolation can be generalized to multiple dimensions, and secondly explores the endogenous characterization of mega-city regions using such a multi-dimensional percolation.
-%  It consists in two main axis: (i) from the methodological viewpoint, how can urban network percolation be extended to multiple dimensions; and (ii) from the thematic viewpoint, how can endogenous mega urban regions be characterized which such a method, and how can it provide insights into their sustainability.
-These two axis are separated, but still tightly related as the second depends on the new methodology, and the first is illustrated through its concrete application in the second.
+We therefore consider percolation on two dimensions of the urban systems, one linked to urban form which is the spatial distribution of population, and one linked to transportation networks, through structural indicators quantifying local road networks.
+We apply in particular the heuristic to urban morphology and road network topology measures in Europe.
+The idea to combine urban form with network topology measures relies on the capture of the link between urban form and function as already mentioned, urban functions being assumed as distributed by transportation networks \citep{raimbault2018caracterisation}.
+In general, the interactions between transportation networks and territories have been shown to be central processes in urban dynamics \citep{espacegeo2014effets}.
 
-In order to test our multi-dimensional percolation method, we propose to apply it to the endogenous characterization of urban regions, and compute stylized sustainability indicators on the constructed regions.
+We thus introduce in this paper a multi-dimensional percolation method, and apply it to the endogenous characterization of urban regions, to finally compute stylized sustainability indicators on the constructed regions.
+These sustainability measures are extrapolated from gravity flows, and are used as a proof-of-concept of how this work can be applied towards policy-making.
+As detailed below, we use paradigmatic opposed dimensions of sustainability which are greenhouse gases emissions and economic productivity.
 
-We apply the heuristic to urban morphology and road network topology measures in Europe.
-The idea to combine urban form with network topology measures relies on the capture of the link between urban form and function as already mentioned, urban functions being assumed as distributed by transportation networks \citep{raimbault2018caracterisation}.
+The originality of our contribution relies thus on several points: (i) this is to the best of our knowledge the first time a multi-dimensional percolation method is applied to urban systems; (ii) we furthermore apply it on the significant spatial extent of all European Union; and (iii) we link the clusters obtained with simple sustainability measures.
 
-% TODO dimension of the urban system taken into account were better specified in the introduction
-
-Our contribution relies on several points: (i) this is to the best of our knowledge the first time a multi-dimensional percolation method is applied to urban systems; (ii) we furthermore apply it on the significant spatial extent of all European Union; and (iii) we link the clusters obtained with simple sustainability measures.
-The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied, and the indicators used to characterized the sustainability of clusters produced.
-We then describe the results of applying this method to population and network variables for the whole European Union, focusing on the endogenous regions produced and their sustainability properties.
+The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied and how the network is constructed, and the indicators used to characterized the sustainability of clusters produced.
+We then describe the results of applying this method to population and road network indicators for the whole European Union, focusing on the endogenous regions produced and their sustainability properties.
 We finally discuss possible developments and the implications of this methodology to the design of policies.
-
-% TODO add more details on all aspects of methodology
-
-% TODO more details in the text on formulas and parameters
 
@@
 
-% TODO rewrite legend of fig 1
-
 %%%%%%%%%%%%%%%%%%%%
@@
   \caption{Schematic representation of the multi-dimensional network percolation heuristic.
-We show a stylized configuration with two layers having the same nodes, and the links within each layer are created following the percolation radius $r_0$ and the thresholds for the layer variables.
-The final clusters are the superposition of these.
-The green points give examples of starting points in the case of a propagation heuristic.\label{fig:method}}
+Two layers are considered here, with similar nodes but different links in each.
+To each layer is associated a node variable.
+In each layer, links are created according to a distance threshold (percolation radius) and a variable value threshold.
+The final clusters (bottom in color) are obtained by considering the links present in each layer.\label{fig:method}}
 \end{figure*}
@@
 
-Note that we do not call our method ``multi-layer percolation'', as it is not strictly multi-layer since nodes are common.
-The term of multi-dimensional percolation is more suited to the use of multiple variables and thresholds.
-The method works with an arbitrary number of layers.
-See Fig.~\ref{fig:method} for a schematic representation of the method.
+We do not call our method ``multi-layer percolation'', since nodes are common between layers.
+We show in Fig.~\ref{fig:method} for a schematic representation of the method.
 It can be implemented with a propagation heuristic or directly working on adjacency matrices.
-The rationale behind the conjunction of the thresholding of each layer variable and the distance thresholding relies on the idea that two points will interact if they are close enough, but also if they have a strong enough intensity of the activity or dimension captured by each layer, simultaneously for all layers considered.
+The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a large enough proximity.
+The second component is a strong enough intensity of the urban dimension captured by each layer, simultaneously for all layers considered.
 This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.
@@
 More formally, let assume a set of nodes $V = v_i$ common to all layers of the network, and layers edges $E_j$ for layer $j$ taken as empty at the initial state of the algorithm.
-Each node has a value of the considered variables associated to each layer, written $v_{ij}$.
+Each node has a value of the considered variables associated to each layer, written as a matrix $v_{ij}$.
+The algorithm works as follows:
+\begin{enumerate}
+	\item Percolation is first done within each layer.
 For each layer, a link $e_{kl} \in E_j$ is created if $d(v_k,v_l) < r_0$ where $d$ is the distance between the nodes (which can be any distance) and $r_0$ the percolation radius, and if $v_{kj} > \theta_j$ and $v_{lj} > \theta_j$ where $\theta_j$ is the threshold for layer $j$.
-The final percolated network edges $E$ is composed by links contained within all layers simultaneously.
+	\item Layers are combined, by computing the final percolated network edges $E$ as the links contained within all layers simultaneously.
 The multi-dimensional percolation clusters are then the connected components of this network $(V,E)$.
-The parameters implied in this heuristic are the percolation radius $r_0$ and the percolation thresholds $\theta_j$ for each layer $j$, allowing a flexible application through parametrization.
+\end{enumerate}
+
+The parameters implied in this heuristic are the percolation radius $r_0$ and the percolation thresholds $\vec{\theta} = \theta_j$ for each layer $j$.
+Varying these parameters allows considering different levels of percolation.
 
@@
 In practice, these nodes will be the center of raster cells.
-We consider two layers, one defined by population density and the other by road network structure indicators computed at the same resolution.
-% TODO French-like paragraph
+We will consider two layers, one defined by population density and the other by road network structure indicators computed at the same resolution.
 
-\subsection{Empirical data}
+\subsection{Empirical data and network construction}
 
@@
 	\item cyclomatic number $\mu$ which is defined by $\mu = N_E - N_V + c$ where $c$ is the number of connected components of the graph; this measure captures the number of distinct cycles in the graph
-	\item euclidian efficiency $v_0$, defined by \cite{banos2012towards}
+	\item euclidian efficiency $v_0$, defined by \cite{banos2012towards}, as the average rate between network distance and euclidian distance for all couples of links
 \end{enumerate}
 
-These measures capture functional properties especially for the two last.
-
-% TODO The choice of network indicators was further discussed, in relation with the research question.
-
-% TODO remove the term network variable ? network indicators
+The choice of these measures is aimed at capturing basic aspects of network structure, and functional properties especially for the two last.
+Indeed, euclidian efficiency measures how the network is performant to link nodes, while the number of cycles is linked to redundancy of paths and in a way to robustness.
+These choice are arbitrary, but several aspects of transportation networks are still captured by these indicators.
+An increase of each is related to a more performant network regarding different dimensions, what is relevant for our application.
+A systematic study with other indicators such as centralities or accessibilities is out of the scope of this paper.
 
@@
 We have therefore two levels of networks in our approach, namely the physical road network which local properties are taken here as input, and the abstract two layer network on which we do the percolation.
+
 We will in the following write $\theta_P$ for the threshold parameter of the population layer, and $\theta_N$ for the threshold parameter of the network layer.
 In practice, these parameters will be given in the following as quantile level of the corresponding variable, for an easier interpretation and conception of experience plans.
-The name of the variable considered will be written $v_N$.
+The name of the road network indicator considered will be written $v_N$.
 
@@
 Its characterization as quantitative indicators is even more subject to numerous degrees of freedom.
-We work here with two stylized indicators for two conflicting dimensions, as a proof-of-concept.
-
-% TODO The choice of these dimensions of sustainability is to give an illustration with two paradigmatic opposed dimensions, which can furthermore be approximated indirectly from gravity models as we propose.
-The explanation of this choice was added in the introduction.
+We work here with these two stylized indicators for two conflicting dimensions, as a proof-of-concept.
+These dimensions can furthermore be approximated indirectly from gravity models as we will describe below.
+By introducing other datasets, our work could be extended to more realistic indicators and other dimensions.
 
@@
 
-Applying a gravity model to each region, we estimate abstract transportation flows within each and use these to extrapolate emissions from the actual local emission from the Edgar database, and economic activities with a scaling law of population.
+Applying a gravity model to each region, we estimate abstract transportation flows within each with a gravity model.
 More precisely, following \cite{raimbault2018indirect}, a potential flow between two points $i$ and $j$ can be estimated with the following expression
@@
 
-where $v^{(k)}_i$ are either population or effective local GHG emissions (indexed by $k = 1,2$ respectively), $d_{ij}$ the distance between the two points, $d_0$ a distance decay parameter, and $\gamma$ a scaling parameter.
+where $v^{(k)}_i$ are either population or effective local GHG emissions computed with the EDGAR database (indexed by $k = 1,2$ respectively), $d_{ij}$ the distance between the two points, $d_0$ a distance decay parameter, and $\gamma$ a scaling parameter.
 Indeed, the economic activity follows relatively well scaling laws of populations \citep{bettencourt2007growth}, the exponent being dependant on the activity and the definition of areas on which it is estimated~\citep{cottineau2017diverse}.
+The distance decay captures the geographical span of potential interactions.
+These two parameters $\gamma,d_0$ are left free and varying them allows considering stylized configurations, such as long or short span interactions, or infra- or supra-linear scaling activities.
+Finally, considering the flow with the population variable ($k=1$) provides a proxy for economic flows, while with the GHG emissions ($k=2$) it provides a proxy for emissions in relation with this economic activity.
+Indeed, effective emissions are linked to local emissions and transportation emissions linked to the intensity of flows.
 
-The sum of all flows of points within the geographical span of the cluster (that we approximate as the convex Hull envelope of its points), allows us to approximate the cumulated potential emissions and economic activity.
-Writing clusters $K_c$ as this set of points, we define the total economic flow by
+We then consider the sum of all these flows of points within the geographical span of a given cluster of nodes in our network.
+These clusters are obtained with the percolation method described above, and are numbered by $1 \leq c \leq C$.
+For the sake of simplicity, we approximate the corresponding geographical area as the convex Hull envelope of the points in the cluster, that we write $K_c$.
+By cumulating the flows, we therefore define the total economic flow as the sum of economic flows by
 
@@
 
-This allows defining a relative economic inefficiency by
+These indicators have no interpretable unit and need to be renormalized. This allows defining a relative economic inefficiency by
 
@@
 
-and relative potential emissions by
+where $\max_c E_c$ (resp. $\min_c E_c$) is the maximal (resp. minimal) value of $E_c$ across all clusters. We define in a similar way the relative potential emissions by
 
@@
 
+where $\max_c G_c$ (resp. $\min_c G_c$) is the maximal (resp. minimal) value of $G_c$ across all clusters.
+
 Both indicators should be minimized for sustainability.
-Normalized indicators $\tilde{e}_c,\tilde{g}_c$ are defined in a similar way, but the extrema being computed on all other possible urban configurations with the same $\gamma,d_0$ values.
+Their value is dependant on the number of clusters and their extent, i.e.
+the geographical surface they cover.
+To be able to compare values across different clusterings (corresponding to different parameter values for the percolation heuristic), we finally define normalized indicators $\tilde{e}_c,\tilde{g}_c$ in a similar way, but the extrema being computed on all other possible urban configurations with the same $\gamma,d_0$ values.
 
@@
 Varying the parameters $\gamma$ and $d_0$ allows to control for the economic activity considered (high $\gamma$ values correspond to high added-value activities) and the span of interactions through $d_0$.
+
+For descriptive purposes, we also consider summary measures of clusters, as the population $P_c$ and effective emissions $EM_c$ taken as the sum of population (resp. emissions) of the points in $K_c$.
 
@@
 Connected components yield the clusters that we interpret as endogenous regions.
-
-We recall that the euclidian performance of the network \citep{banos2012towards} is in our case $<d_e/d_n>$ where the average is taken on all origin-destination pairs in the network, $d_e$ is the euclidian distance and $d_n$ the network distance.
-Thus, it indeed increases with network performance, in consistence with the use done here through thresholding.
-% TODO clarify
 
@@
 We study here this property, for different threshold parameter values.
-We make the radius vary betweem 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all network variables, and three network thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.
+We make the radius vary betweem 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.
 
@@
 As expected, absolute size are much smaller, but when looking at relative sizes we observe that the abrupt steps typical to percolation transitions have different distributions across the different parametrizations.
-The more regular curve seems to be the standard percolation on population only, whereas at $\theta_N = 0.95$, different network variables produce either very early transitions (for $\mu$ for example) or very late (for $N_V$).
+This result is particularly interesting regarding the first axis of our research question, as it shows that the structure of clusters obtained is not only due to the population layer, and that the multi-dimensional percolation captures a complementary signal.
+
+The more regular curve seems to be the standard percolation on population only, whereas at $\theta_N = 0.95$, different road network indicators produce either very early transitions (for $\mu$ for example) or very late (for $N_V$).
 Also, changing of scale compared to \cite{arcaute2016cities} gives more steps and less abrupts curves in general, confirming the integration of subsystems with different structures in our analysis and the importance of scale in such analysis.
@@
 
-We study also the evolution of the fractal dimension of clusters as a function of $r_0$.
-Following \cite{arcaute2016cities}, we estimate the fractal dimension of clusters $\alpha$ with a simple OLS regression between cluster size and cluster diameter, namely $\log N_c = k + \alpha \cdot \log \delta_c$ where $N_c$ is the size of cluster $c$ and $\delta_c$ its diameter.
-As a negative result, that could be due to the abstract nature of our network, a clear maximum in the value of the fractal dimension can not be found.
-Either it is located at resolution that our method can not reached due to the minimal 5km limit imposed by the abstraction, or it does not exist when coupling dimensions.
+We study also the evolution of the fractal dimension of clusters as a function of $r_0$, to verify how the initial percolation approach is robust to multi-dimensionality.
+Following \cite{arcaute2016cities}, we estimate the fractal dimension $\alpha$ of clusters with a simple OLS regression between cluster size and cluster diameter, namely
+
+\begin{equation}
+\log N_c = k + \alpha \cdot \log \delta_c
+\end{equation}
+
+where $N_c$ is the size of cluster $c$ and $\delta_c$ its diameter.
+These estimations are shown in Fig.~\ref{fig:fractaldim}.
+As a negative result, which could be due to the abstract nature of our network, a clear maximum in the value of the fractal dimension can not be found.
+Either it is located at a resolution that our method can not reach due to the minimal 5km limit imposed by the abstraction in the network construction, or it does not exist when coupling dimensions.
 Determining which assumption is more plausible is out of the scope of this paper.
-We do not plot the standard error $\sigma$ of fractal dimensions for visibility purposes, but their relative value given by $\alpha / \sigma\left[\alpha\right]$ is in average 0.10 and in maximum 0.196 on all points, meaning that these estimations remain however consistent.
-Regarding the variability as a function of the percolation radius $r_0$, studying the difference $(\alpha - \sigma\left[\alpha\right])_M - (\alpha - \sigma\left[\alpha\right])_m$ where the first is taken at maximum for $\alpha$ and the other at minimum, shows that the configuration for $\mu$ and $\theta_N=0.95$ has a clearly significant maximum (difference at 0.38).
+
+We do not plot the standard error $\sigma$ of fractal dimensions (obtained as the estimation error in the OLS) for visibility purposes, but their relative value given by $\alpha / \sigma\left[\alpha\right]$ is in average 0.10 and in maximum 0.196 on all points, meaning that these estimations remain however consistent.
+
+Regarding the variability of fractal dimension as a function of the percolation radius $r_0$, we study the possible existence of a significant maximum when $r_0$ varies.
+We therefore simply consider the difference $(\alpha - \sigma\left[\alpha\right])_M - (\alpha - \sigma\left[\alpha\right])_m$ where the first is taken at the maximum value of $\alpha$ and the other at its minimum value.
+This intuitively corresponds to checking if confidence intervals do not overlap between the maximum and the minimum of the curve.
+We find that the configuration for $\mu$ and $\theta_N=0.95$ has a clearly significant maximum (difference at 0.38).
 For this coupling, the endogenous structure given by the maximum may be defined.
-Other configurations yield non-significant maximums.
+Other configurations yield non-significant maximums (negative values of the difference).
 
-% TODO The study of fractal dimension is not essential in the methodology, but comes as an auxiliary result to validate the multi-dimensional approach regarding the original papers studying urban percolation.
-This was specified, and the formulas were better explained.
+This study of percolation transition and fractal dimension thus shows that our multi-dimensional percolation heuristic remains relevant, as results analog but qualitatively different to the one-dimensional approach can be obtained.
 
-This study of percolation transition and fractal dimension shows that our multi-dimensional percolation heuristic remains relevant, as results analog to the one-dimensional approach can be obtained.
-
-% TODO table ?
 %  fractdimsignif
@@
 
-We now switch the experience plan to a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the network variable considered, and also make $\gamma$ and $d_0$ vary.
-We systematically explore the clusters obtained for 4800 parameter configurations, such that for all network variables, $\theta_P \in \{ 0.8 ; 0.9 ; 0.95 \}$, $\theta_N \in \{0 ; 0.8 ; 0.95 \}$, $r_0 \in \{ 8 ; 10 ; 15 ; 20 ; 50\}$ km, $\gamma \in \{ 0.5 ; 1 ; 1.5 ; 2\}$, and $d_0 \in \{ 0.1 ; 1 ; 10 ; 50 ; 100\}$ km.
+We now switch the experience plan to a full grid, for parameters $r_0$, $\theta_P$, $\theta_N$ and the road network indicator considered, and also make $\gamma$ and $d_0$ vary.
+We systematically explore the clusters obtained for 4800 parameter configurations, such that for all road network indicator, $\theta_P \in \{ 0.8 ; 0.9 ; 0.95 \}$, $\theta_N \in \{0 ; 0.8 ; 0.95 \}$, $r_0 \in \{ 8 ; 10 ; 15 ; 20 ; 50\}$ km, $\gamma \in \{ 0.5 ; 1 ; 1.5 ; 2\}$, and $d_0 \in \{ 0.1 ; 1 ; 10 ; 50 ; 100\}$ km.
 
@@
 We show some examples in Fig.~\ref{fig:exclusters}.
-The first map of this figure, obtained for high population and network thresholds ($\theta_P = 0.95$ and $\theta_N = 0.9$), but a low radius $r_0 = 8$km and edge count $N_E$ as network variable, include several mega-city regions described by \citep{hall2006polycentric}, namely London metropolitan area, the Randstad in Netherland, the Rhein-Main and Rhein-Ruhr in Germany, Greater Paris in France, Brussels area in Belgium.
+The first map of this figure, obtained for high population and network thresholds ($\theta_P = 0.95$ and $\theta_N = 0.9$), but a low radius $r_0 = 8$km and edge count $N_E$ to define the road network layer, include several mega-city regions described by \citep{hall2006polycentric}, namely London metropolitan area, the Randstad in Netherland, the Rhein-Main and Rhein-Ruhr in Germany, Greater Paris in France, Brussels area in Belgium.
 The same parameters with $\theta_N = 0$ yield not exactly the same regions, as confirmed by the transition curves in Fig.~\ref{fig:percolation}, what means that our approach taking into account two dimensions may capture effective processes of mega-city regions, in particular by including the road network which is crucial as these regions are integrated in terms of flows.
@@
 The behavior of sustainability indicators for different population, network and distance thresholds yield different distributions of performances across clusters within a configuration but also between configurations.
-Before considering the flow-based indicators described above, we can already study basic measures such as population $P_c$ and effective emissions $EM_c$, taken as the sum within the cluster of their values at each point.
+Before considering the flow-based indicators described above, we can already study basic summary measures such as population $P_c$ and effective emissions $EM_c$.
 We show in Fig.~\ref{fig:paretos} point clouds of $\log EM_c$ against $1 - \log P_c$ for some configurations.
@@
 In Fig.~\ref{fig:paretos}, the front is the lower bound of the point cloud.
-We also find no dominating point for each configuration.
+We also find no dominating point for each configuration, i.e.
+that considering point clouds of a single color, a Pareto front with more than one point is still present.
 Some clusters are therefore optimal compromises in the Pareto sense in each configuration, while some are dominated and thus not efficient.
+This confirms that urban systems are generally compromises between multiple objectives.
 
@@
 
-% TODO This part was clarified by adding more context literature and more explanations.
-
 %# Cumulative Proportion  0.7296 0.9650 0.99761 1.00000
@@
 
-We now consider the sustainibility indicators, aggregated for a configuration on all clusters.
-For a given parametrization of endogenous city regions, one can relate them to morphological indicators for population density spatial distribution, computed by~\cite{raimbault2018calibration}, that we average on clusters.
+We now consider the sustainability indicators, aggregated for a configuration on all clusters.
+An important question is how these relate with measures of urban form \citep{le2012urban}.
+For a given parametrization of endogenous city regions, one can relate them to morphological indicators for population density spatial distribution.
+Such indicators were computed by~\cite{raimbault2018calibration}, and capture different dimensions of the spatial distribution of population, such as spatial autocorrelation (Moran index), homogeneity (entropy index), hierarchy (rank-size exponent), or level of aggregation (average distance between individuals).
+We average them here on clusters.
 This establishes a link between urban morphology and sustainibility.
@@
 The first component relates to a level of monocentricity ($PC1 = -0.3\cdot I + 0.54 \cdot \bar{d} + 0.51\cdot \varepsilon + 0.59 \cdot h$ where $I$ is Moran index, $\bar{d}$ average distance, $\epsilon$ entropy, and $h$ level of hierarchy).
+In a nutshell, the principal dimension of urban form in differentiating our urban clusters is the level of monocentricity.
+We can relate it with indicators for emissions and economic efficiency.
 
@@
 As the color level gives the value of PC1, we can see the points on the different fronts with very different morphological properties.
-In some case, highly monocentric areas (yellow points) can be a good compromise, whereas the intermediate optimal for emissions shown before may yield highly inefficient areas (dominated green points).
+In some cases, highly monocentric areas (yellow points) can be a good compromise, whereas the intermediate optimal for emissions shown before may yield highly inefficient areas (dominated green points).
 For example, considering the fronts for $\gamma = 2$ which have both very similar shape, the points with the lowest emissions are on the top-left of the front and correspond to the optimal unveiled in Fig.~\ref{fig:emissions-pc1}.
@@
 
-% TODO extend with relation to research questions
+\subsection{From multi-dimensional percolation to the sustainability of mega-city regions}
 
-% TODO But the only fact of applying a new methodology does not make a research result to me: it is necessary to show how brought information is significant to analyze an urban system
+From the methodological point of view, we showed that network percolation can successfully be applied to multidimensional urban networks.
+This requires a consistent overlay within the same nodes of the different dimensions considered.
+The existence of percolation transitions which are different to the unidimensional case confirms that the approach captures complementary information, and that it could be applied to characterize urban systems in a more refined way.
+The non-existence of a significant maximum for the fractal dimension remains to be investigated, since it could be due a bias of our abstract network construction.
+Studies on other dimensions or on non-abstract networks remain to be done to understand how multi-dimensional percolation differs from the unidimensional one.
 
-% TODO The discussion was extended in relation with answers to the research question.
+Regarding the second axis of our research question, we showed that multi-dimensional percolation is a useful tool to extract endogenous mega-urban regions while taking into account complementary aspects of population distribution and performance of transportation networks.
+Varying the parameters of the percolation algorithm provides a comparative view on possible clustering structures for the European urban system, and corresponding performance in terms of stylized sustainability indicators.
+Indeed, this work is exploratory in terms of possible definitions of urban subsystems.
+The fact that some systems obtained coincide with effective functional regions \citep{hall2006polycentric} shows that some thresholds of population and road network performance intensity capture actual functional linkages.
+This correspondance could not be predicted a priori nor explained through simple arguments as our approach reconstructs clusters from the bottom-up.
+Finally, the links between urban form and sustainability indicators made in the last section are also interesting for the management of urban systems, suggesting a certain performance of polycentric systems in particular regarding emissions.
 
@@
 Finally, an important development would imply crossing our endogenous definitions of urban regions with socio-economic databases, and compute indicators implied in other dimensions of sustainability, for example related to socio-economic inequalities, spatial distribution of accessibilities, or activities with different scaling exponents.
-This includes the mitigation of spatial inequalities and segregation \citep{tammaru2015multi}, which are an important dimension of sustainibility.
+This includes the mitigation of spatial inequalities and segregation \citep{tammaru2015multi}, which are an important dimension of sustainability.
 
@@
 In conclusion, our multilayer percolation approach captures in a way the multi-dimensionality of urban systems and a link between form and function in urban system.
-Its application to the issue of sustainable mega-city regions shows its potentialities.
+In particular, in our application on the bilayer case of an abstract network constructed from population density and road network indicators, it is shown to capture a different structure than in the unidimensional case.
+Its application to the issue of sustainable mega-city regions show how its properties of constructing urban clusters from the bottom-up can be used to study sustainability issues.
 This work also illustrates the importance of following data-driven paradigms even when developing, as what is understood of the behavior of the heuristic is through its application to real data and issues.

### 2019-10-24 Docs/Papers/Percolation/paper/paper.tex
final_version_paper
prod_version_paper

@@
 \title{Multi-dimensional Urban Network Percolation}
-% Identifying European Sustainable Mega-city Regions with Multi-dimensional Network Percolation
-% Endogenous urban regions as an optimal compromise between economic integration and greenhouse gases emissions
-% A multi-dimensional percolation approach to characterize sustainable mega-city regions
 
@@
 The method allows combining different dimensions of the urban system, the same way that \cite{cottineau2018defining} link population density with commuting flows to produce multiple definitions of urban areas.
+We can indeed expect significantly different qualitative behaviors when switching from a single-dimension percolation to a multi-dimensional one, since the spatial structure of different urban dimensions are correlated but also largely complementary.
+If this is the case, the method then captures more aspects of the urban system and some of their interrelations.
 
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/principle.pdf}}
+  {\includegraphics[width=\linewidth]{Fig1.png}}
   \centering
@@
 It can be implemented with a propagation heuristic or directly working on adjacency matrices.
-The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a large enough proximity.
+The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a geographical proximity.
 The second component is a strong enough intensity of the urban dimension captured by each layer, simultaneously for all layers considered.
@@
 
-The parameters implied in this heuristic are the percolation radius $r_0$ and the percolation thresholds $\vec{\theta} = \theta_j$ for each layer $j$.
+The parameters implied in this heuristic are the percolation radius $r_0$ and the variable thresholds $\vec{\theta} = \theta_j$ for each layer $j$.
 Varying these parameters allows considering different levels of percolation.
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.49\linewidth]{figures/abssize_nodes.png}}
-  {\includegraphics[width=0.49\linewidth]{figures/relsize_nodes.png}}
+  {\includegraphics[width=\linewidth]{Fig2.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/fractaldimension.png}}
+  {\includegraphics[width=\linewidth]{Fig3.png}}
   \centering
@@
 We study here this property, for different threshold parameter values.
-We make the radius vary betweem 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.
+We make the radius vary between 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.
+These values yield a good precision for the radius which is the most important variable to study transition and estimate fractal dimensions, while this population threshold is enough to provide a broad coverage for large radiuses (as shown by other explorations described below).
+Changing network variables and their threshold allows us to investigate how the transition behavior do change regarding the dimension considered and its intensity.
 
@@
 As the addition of a layer also changes drastically the results, one should stay careful when switching from a mono-dimensional percolation to a multi-dimensional percolation.
-
-% ! largest cluster size in terms of edges makes no sense with the implementation we took - nodes only
-%      nwcol         popthq         nwthq          radius         nwparam             fractdim     
-% ecount  :186   Min.   :0.85   Min.   :0.00   Min.   :  8000   Length:1023        Min.   :0.9239  
-% euclPerf:279   1st Qu.:0.85   1st Qu.:0.00   1st Qu.: 31000   Class :character   1st Qu.:1.2639  
-% mu      :279   Median :0.85   Median :0.80   Median : 54000   Mode  :character   Median :1.3306  
-% vcount  :279   Mean   :0.85   Mean   :0.55   Mean   : 54000                      Mean   :1.3225  
- %               3rd Qu.:0.85   3rd Qu.:0.95   3rd Qu.: 77000                      3rd Qu.:1.3909  
- %               Max.   :0.85   Max.   :0.95   Max.   :100000                      Max.   :1.6058                                                                                    NA's   :337     
-%   fractdimsd     fractdimadjrsquared   fractrelsd    
-% Min.   :0.0544   Min.   :0.5111      Min.   :0.0395  
-% 1st Qu.:0.0966   1st Qu.:0.7016      1st Qu.:0.0700  
-% Median :0.1328   Median :0.7467      Median :0.1043  
-% Mean   :0.1317   Mean   :0.7456      Mean   :0.1009  
-% 3rd Qu.:0.1578   3rd Qu.:0.7981      3rd Qu.:0.1232  
-% Max.   :0.2381   Max.   :0.8642      Max.   :0.1958  
-% NA's   :337      NA's   :337         NA's   :337
 
@@
 
-%  fractdimsignif
-%# A tibble: 11 x 5
-%# Groups:   nwcol, popthq, nwthq [?]
-%   nwcol    popthq nwthq nwparam       signif
-%   <fct>     <dbl> <dbl> <chr>          <dbl>
-% 1 ecount     0.85  0    ecount0      -0.105 
-% 2 ecount     0.85  0.8  ecount0.8    -0.0452
-% 3 euclPerf   0.85  0    euclPerf0    -0.108 
-% 4 euclPerf   0.85  0.8  euclPerf0.8   0.0360
-% 5 euclPerf   0.85  0.95 euclPerf0.95 -0.108 
-% 6 mu         0.85  0    mu0          -0.108 
-% 7 mu         0.85  0.8  mu0.8        -0.108 
-% 8 mu         0.85  0.95 mu0.95        0.375 
-% 9 vcount     0.85  0    vcount0      -0.105 
-%10 vcount     0.85  0.8  vcount0.8    -0.0416
-%11 vcount     0.85  0.95 vcount0.95   -0.0868
-
 %%%%%%%%%%%%%%%%%%%%
@@
 %\resizebox{10cm}{7cm}
-{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}\\
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
-   {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
+{\includegraphics[width=\linewidth]{Fig4.png}}
   \centering
-
   \caption{Examples of obtained clusters for different parameter values.
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.7\linewidth]{figures/full_effective_pareto.png}}
+  {\includegraphics[width=0.7\linewidth]{Fig5.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_pc1-emissions_targeted.png}}
+  {\includegraphics[width=\linewidth]{Fig6.png}}
   \centering
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
+  {\includegraphics[width=\linewidth]{Fig7.png}}
   \centering  
@@
 \subsection{Linking urban morphology and sustainability}
-
-%# Cumulative Proportion  0.7296 0.9650 0.99761 1.00000
-%                PC1       PC2         PC3         PC4
-%moran   -0.3088585 0.9493848 -0.04444327  0.03605266
-%avgdist  0.5417362 0.1415668 -0.82239570 -0.10072759
-%entropy  0.5108424 0.2140447  0.45847647 -0.69499942
-%slope    0.5917502 0.1811415  0.33390034  0.71100630
 
@@
 
-\bibliographystyle{jimis-en}
-\bibliography{biblio}
+\section*{Acknowledgment}
 
-%\appendix\footnotesize
+This work is part of DynamiCity, a FUI project funded by BPI France, Auvergne-Rh{\^o}ne-Alpes region, Ile-de-France region and Lyon metropolis.
 
-%\section{Appendix 1: supplementary sensitivity analyses}
+\begin{thebibliography}{}
 
-%\sframe{Results: effective emissions}{
+\bibitem[\protect\citeauthoryear{Aleta and Moreno}{Aleta and
+  Moreno}{2018}]{aleta2018multilayer}
+Aleta A., Moreno Y. (2018).
+\newblock Multilayer networks in a nutshell.
+\newblock {\em Annual Review of Condensed Matter Physics\/}.
 
-%\textit{Effective emissions exhibit a supralinear scaling of population}
-%\includegraphics[height=0.83\textheight]{figures/aggreg_effective.png}
+\bibitem[\protect\citeauthoryear{Arbabi, Mayfield, and McCann}{Arbabi
+  \textit{et~al.}}{2019}]{arbabi2019development}
+Arbabi H., Mayfield M., McCann P. (2019).
+\newblock On the development logic of city-regions: inter-versus intra-city
+  mobility in england and wales.
+\newblock {\em Spatial Economic Analysis\/}, 1--20.
 
-%\sframe{Results: all clusters Pareto fronts}{
+\bibitem[\protect\citeauthoryear{Arcaute, Molinero, Hatna, Murcio, Vargas-Ruiz,
+  Masucci, and Batty}{Arcaute \textit{et~al.}}{2016}]{arcaute2016cities}
+Arcaute E., Molinero C., Hatna E., Murcio R., Vargas-Ruiz C., Masucci A.~P.,
+  Batty M. (2016).
+\newblock Cities and regions in britain through hierarchical percolation.
+\newblock {\em Royal Society open science\/}~{\em 3\/}(4), 150691.
 
-%\textit{Variation of Pareto front patterns when potential parameter $\gamma,d_0$ vary.}
+\bibitem[\protect\citeauthoryear{Banos and Genre-Grandpierre}{Banos and
+  Genre-Grandpierre}{2012}]{banos2012towards}
+Banos A., Genre-Grandpierre C. (2012).
+\newblock Towards new metrics for urban road networks: Some preliminary
+  evidence from agent-based simulations.
+\newblock In {\em Agent-based models of geographical systems}, pp.\  627--641.
+  Springer.
 
-%\includegraphics[width=0.9\textwidth]{figures/full_paretos.png}
+\bibitem[\protect\citeauthoryear{Barth{\'e}lemy}{Barth{\'e}lemy}{2011}]{barthelemy2011spatial}
+Barth{\'e}lemy M. (2011).
+\newblock Spatial networks.
+\newblock {\em Physics Reports\/}~{\em 499\/}(1-3), 1--101.
 
-%\sframe{Results: an optimal morphology}{
+\bibitem[\protect\citeauthoryear{Batty and Longley}{Batty and
+  Longley}{1994}]{batty1994fractal}
+Batty M., Longley P.~A. (1994).
+\newblock {\em Fractal cities: a geometry of form and function}.
+\newblock Academic press.
 
-%\textit{More monocentric areas are more optimal in terms of relative emissions and efficiency ?}
+\bibitem[\protect\citeauthoryear{Bettencourt, Lobo, Helbing, K{\"u}hnert, and
+  West}{Bettencourt \textit{et~al.}}{2007}]{bettencourt2007growth}
+Bettencourt L.~M., Lobo J., Helbing D., K{\"u}hnert C., West G.~B. (2007).
+\newblock Growth, innovation, scaling, and the pace of life in cities.
+\newblock {\em Proceedings of the national academy of sciences\/}~{\em
+  104\/}(17), 7301--7306.
 
-%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relefficiency.png}
-%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relemissions.png}
+\bibitem[\protect\citeauthoryear{Boccaletti, Bianconi, Criado, Del~Genio,
+  G{\'o}mez-Gardenes, Romance, Sendina-Nadal, Wang, and Zanin}{Boccaletti
+  \textit{et~al.}}{2014}]{boccaletti2014structure}
+Boccaletti S., Bianconi G., Criado R., Del~Genio C.~I., G{\'o}mez-Gardenes J.,
+  Romance M., Sendina-Nadal I., Wang Z., Zanin M. (2014).
+\newblock The structure and dynamics of multilayer networks.
+\newblock {\em Physics Reports\/}~{\em 544\/}(1), 1--122.
 
-%  curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.
+\bibitem[\protect\citeauthoryear{Burger and Meijers}{Burger and
+  Meijers}{2012}]{burger2012form}
+Burger M., Meijers E. (2012).
+\newblock Form follows function? linking morphological and functional
+  polycentricity.
+\newblock {\em Urban studies\/}~{\em 49\/}(5), 1127--1149.
 
-%\section{Acknowledgment}
-%
+\bibitem[\protect\citeauthoryear{Callaway, Newman, Strogatz, and
+  Watts}{Callaway \textit{et~al.}}{2000}]{callaway2000network}
+Callaway D.~S., Newman M.~E., Strogatz S.~H., Watts D.~J. (2000).
+\newblock Network robustness and fragility: Percolation on random graphs.
+\newblock {\em Physical review letters\/}~{\em 85\/}(25), 5468.
 
-%\section{Biography}
-%Here, feel free to add short biographies of the authors.
+\bibitem[\protect\citeauthoryear{Cottineau, Finance, Hatna, Arcaute, and
+  Batty}{Cottineau \textit{et~al.}}{2018}]{cottineau2018defining}
+Cottineau C., Finance O., Hatna E., Arcaute E., Batty M. (2018).
+\newblock Defining urban clusters to detect agglomeration economies.
+\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
+  2399808318755146.
+
+\bibitem[\protect\citeauthoryear{Cottineau, Hatna, Arcaute, and
+  Batty}{Cottineau \textit{et~al.}}{2017}]{cottineau2017diverse}
+Cottineau C., Hatna E., Arcaute E., Batty M. (2017).
+\newblock Diverse cities or the systematic paradox of urban scaling laws.
+\newblock {\em Computers, environment and urban systems\/}~{\em 63}, 80--94.
+
+\bibitem[\protect\citeauthoryear{Crucitti, Latora, and Porta}{Crucitti
+  \textit{et~al.}}{2006}]{crucitti2006centrality}
+Crucitti P., Latora V., Porta S. (2006).
+\newblock Centrality measures in spatial networks of urban streets.
+\newblock {\em Physical Review E\/}~{\em 73\/}(3), 036125.
+
+\bibitem[\protect\citeauthoryear{Ducruet and Beauguitte}{Ducruet and
+  Beauguitte}{2014}]{ducruet2014spatial}
+Ducruet C., Beauguitte L. (2014).
+\newblock Spatial science and network science: Review and outcomes of a complex
+  relationship.
+\newblock {\em Networks and Spatial Economics\/}~{\em 14\/}(3-4), 297--316.
+
+\bibitem[\protect\citeauthoryear{Feng, Wu, Wu, Su, Weng, and Bian}{Feng
+  \textit{et~al.}}{2018}]{feng2018spatiotemporal}
+Feng Y., Wu S., Wu P., Su S., Weng M., Bian M. (2018).
+\newblock Spatiotemporal characterization of megaregional poly-centrality:
+  Evidence for new urban hypotheses and implications for polycentric policies.
+\newblock {\em Land use policy\/}~{\em 77}, 712--731.
+
+\bibitem[\protect\citeauthoryear{Florida, Gulden, and Mellander}{Florida
+  \textit{et~al.}}{2008}]{florida2008rise}
+Florida R., Gulden T., Mellander C. (2008).
+\newblock The rise of the mega-region.
+\newblock {\em Cambridge Journal of Regions, Economy and Society\/}~{\em
+  1\/}(3), 459--476.
+
+\bibitem[\protect\citeauthoryear{Hackett, Cellai, G{\'o}mez, Arenas, and
+  Gleeson}{Hackett \textit{et~al.}}{2016}]{hackett2016bond}
+Hackett A., Cellai D., G{\'o}mez S., Arenas A., Gleeson J.~P. (2016).
+\newblock Bond percolation on multiplex networks.
+\newblock {\em Physical Review X\/}~{\em 6\/}(2), 021002.
+
+\bibitem[\protect\citeauthoryear{Hall and Pain}{Hall and
+  Pain}{2006}]{hall2006polycentric}
+Hall P.~G., Pain K. (2006).
+\newblock {\em The polycentric metropolis: learning from mega-city regions in
+  Europe}.
+\newblock Routledge.
+
+\bibitem[\protect\citeauthoryear{Hillier, Leaman, Stansall, and
+  Bedford}{Hillier \textit{et~al.}}{1976}]{hillier1976space}
+Hillier B., Leaman A., Stansall P., Bedford M. (1976).
+\newblock Space syntax.
+\newblock {\em Environment and Planning B: Planning and design\/}~{\em 3\/}(2),
+  147--185.
+
+\bibitem[\protect\citeauthoryear{Huynh, Makarov, Legara, Monterola, and
+  Chew}{Huynh \textit{et~al.}}{2018}]{huynh2018characterisation}
+Huynh H.~N., Makarov E., Legara E.~F., Monterola C., Chew L.~Y. (2018).
+\newblock Characterisation and comparison of spatial patterns in urban systems:
+  A case study of us cities.
+\newblock {\em Journal of computational science\/}~{\em 24}, 34--43.
+
+\bibitem[\protect\citeauthoryear{Janssens-Maenhout, Crippa, Guizzardi, Muntean,
+  Schaaf, Dentener, Bergamaschi, Pagliari, Olivier, Peters,
+  \textit{et~al.}}{Janssens-Maenhout \textit{et~al.}}{2017}]{janssens2017edgar}
+Janssens-Maenhout G., Crippa M., Guizzardi D., Muntean M., Schaaf E., Dentener
+  F., Bergamaschi P., Pagliari V., Olivier J., Peters J., \textit{et~al.}
+  (2017).
+\newblock Edgar v4. 3.2 global atlas of the three major greenhouse gas
+  emissions for the period 1970--2012.
+\newblock {\em Earth Syst. Sci. Data Discuss\/}.
+
+\bibitem[\protect\citeauthoryear{Komiyama and Takeuchi}{Komiyama and
+  Takeuchi}{2006}]{komiyama2006sustainability}
+Komiyama H., Takeuchi K. (2006).
+\newblock {\em Sustainability science: building a new discipline}.
+
+\bibitem[\protect\citeauthoryear{Lagesse, Bordin, and Douady}{Lagesse
+  \textit{et~al.}}{2015}]{lagesse2015spatial}
+Lagesse C., Bordin P., Douady S. (2015).
+\newblock A spatial multi-scale object to analyze road networks.
+\newblock {\em Network Science\/}~{\em 3\/}(1), 156--181.
+
+\bibitem[\protect\citeauthoryear{Lang and Dhavale}{Lang and
+  Dhavale}{2005}]{lang2005beyond}
+Lang R.~E., Dhavale D. (2005).
+\newblock Beyond megalopolis: Exploring america’s new “megapolitan”
+  geography.
+
+\bibitem[\protect\citeauthoryear{Laquian}{Laquian}{2011}]{laquian2011planning}
+Laquian A. (2011).
+\newblock The planning and governance of asia’s mega-urban regions.
+\newblock {\em Population Distribution, Urbanization, Internal Migration and
+  Development: An International Perspective, United Nations Department of
+  Economic and Social Affairs, Population Division, New York: United
+  Nations\/}, 302--322.
+
+\bibitem[\protect\citeauthoryear{Lashof and Ahuja}{Lashof and
+  Ahuja}{1990}]{lashof1990relative}
+Lashof D.~A., Ahuja D.~R. (1990).
+\newblock Relative contributions of greenhouse gas emissions to global warming.
+\newblock {\em Nature\/}~{\em 344\/}(6266), 529.
+
+\bibitem[\protect\citeauthoryear{Le~N{\'e}chet}{Le~N{\'e}chet}{2012}]{le2012urban}
+Le~N{\'e}chet F. (2012).
+\newblock Urban spatial structure, daily mobility and energy consumption: a
+  study of 34 european cities.
+\newblock {\em Cybergeo: European Journal of Geography\/}.
+
+\bibitem[\protect\citeauthoryear{Le~N{\'e}chet}{Le~N{\'e}chet}{2017}]{lenechet2017peupler}
+Le~N{\'e}chet F. (2017).
+\newblock De l'{\'e}talement urbain aux r{\'e}gions m{\'e}tropolitaines
+  polycentriques : formes de fonctionnement et formes de gouvernance.
+\newblock In {\em {Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
+  m{\'e}tropoles}}. Presses Universitaires Francois Rabelais.
+
+\bibitem[\protect\citeauthoryear{Li, Fu, Wang, Lu, Berezin, Stanley, and
+  Havlin}{Li \textit{et~al.}}{2015}]{Li669}
+Li D., Fu B., Wang Y., Lu G., Berezin Y., Stanley H.~E., Havlin S. (2015).
+\newblock Percolation transition in dynamical traffic network with evolving
+  critical bottlenecks.
+\newblock {\em Proceedings of the National Academy of Sciences\/}~{\em
+  112\/}(3), 669--672.
+\newblock URL: \url{https://www.pnas.org/content/112/3/669}, \href
+  {http://arxiv.org/abs/https://www.pnas.org/content/112/3/669.full.pdf}
+  {arXiv:https://www.pnas.org/content/112/3/669.full.pdf}, \href
+  {http://dx.doi.org/10.1073/pnas.1419185112} {doi:10.1073/pnas.1419185112}.
+
+\bibitem[\protect\citeauthoryear{Louf and Barthelemy}{Louf and
+  Barthelemy}{2013}]{louf2013modeling}
+Louf R., Barthelemy M. (2013).
+\newblock Modeling the polycentric transition of cities.
+\newblock {\em Physical review letters\/}~{\em 111\/}(19), 198702.
+
+\bibitem[\protect\citeauthoryear{Makse, Andrade, Batty, Havlin, Stanley,
+  \textit{et~al.}}{Makse \textit{et~al.}}{1998}]{makse1998modeling}
+Makse H.~A., Andrade J.~S., Batty M., Havlin S., Stanley H.~E., \textit{et~al.}
+  (1998).
+\newblock Modeling urban growth patterns with correlated percolation.
+\newblock {\em Physical Review E\/}~{\em 58\/}(6), 7054.
+
+\bibitem[\protect\citeauthoryear{Marull, Galletto, Domene, and
+  Trull{\'e}n}{Marull \textit{et~al.}}{2013}]{marull2013emerging}
+Marull J., Galletto V., Domene E., Trull{\'e}n J. (2013).
+\newblock Emerging megaregions: a new spatial scale to explore urban
+  sustainability.
+\newblock {\em Land Use Policy\/}~{\em 34}, 353--366.
+
+\bibitem[\protect\citeauthoryear{McNally}{McNally}{2000}]{mcnally2000four}
+McNally M.~G. (2000).
+\newblock The four step model.
+
+\bibitem[\protect\citeauthoryear{Newman and Watts}{Newman and
+  Watts}{1999}]{newman1999scaling}
+Newman M.~E., Watts D.~J. (1999).
+\newblock Scaling and percolation in the small-world network model.
+\newblock {\em Physical review E\/}~{\em 60\/}(6), 7332.
+
+\bibitem[\protect\citeauthoryear{Offner, Beaucire, Delaplace, Fr{\'e}mont,
+  Ninot, Bretagnolle, and Pumain}{Offner
+  \textit{et~al.}}{2014}]{espacegeo2014effets}
+Offner J.-M., Beaucire F., Delaplace M., Fr{\'e}mont A., Ninot O., Bretagnolle
+  A., Pumain D. (2014).
+\newblock Les effets structurants des infrastructures de transport.
+\newblock {\em Espace Geographique\/}~(42), p--51.
+
+\bibitem[\protect\citeauthoryear{Penrose}{Penrose}{1999}]{penrose1999k}
+Penrose M.~D. (1999).
+\newblock On k-connectivity for a geometric random graph.
+\newblock {\em Random Structures \& Algorithms\/}~{\em 15\/}(2), 145--164.
+
+\bibitem[\protect\citeauthoryear{Perez, Banos, and Pettit}{Perez
+  \textit{et~al.}}{2016}]{perez2016agent}
+Perez P., Banos A., Pettit C. (2016).
+\newblock Agent-based modelling for urban planning current limitations and
+  future trends.
+\newblock In {\em International Workshop on Agent Based Modelling of Urban
+  Systems}, pp.\  60--69. Springer.
+
+\bibitem[\protect\citeauthoryear{Piovani, Molinero, and Wilson}{Piovani
+  \textit{et~al.}}{2017}]{piovani2017urban}
+Piovani D., Molinero C., Wilson A. (2017).
+\newblock Urban retail location: insights from percolation theory and spatial
+  interaction modeling.
+\newblock {\em PloS one\/}~{\em 12\/}(10), e0185787.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018a}]{raimbault2018calibration}
+Raimbault J. (2018a).
+\newblock Calibration of a density-based model of urban morphogenesis.
+\newblock {\em PloS one\/}~{\em 13\/}(9), e0203516.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018b}]{raimbault2018caracterisation}
+Raimbault J. (2018b).
+\newblock {\em Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution
+  des r{\'e}seaux de transport et des territoires}.
+\newblock Ph.\ D. thesis, Universit{\'e} Paris 7 Denis Diderot.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018c}]{raimbault2018indirect}
+Raimbault J. (2018c).
+\newblock Indirect evidence of network effects in a system of cities.
+\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
+  2399808318774335.
+
+\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2019}]{raimbault2018urban}
+Raimbault J. (2019).
+\newblock An urban morphogenesis model capturing interactions between networks
+  and territories.
+\newblock In {\em The Mathematics of Urban Morphology}, pp.\  383--409.
+  Springer.
+
+\bibitem[\protect\citeauthoryear{Reuillon, Leclaire, and
+  Rey-Coyrehourcq}{Reuillon \textit{et~al.}}{2013}]{reuillon2013openmole}
+Reuillon R., Leclaire M., Rey-Coyrehourcq S. (2013).
+\newblock Openmole, a workflow engine specifically tailored for the distributed
+  exploration of simulation models.
+\newblock {\em Future Generation Computer Systems\/}~{\em 29\/}(8), 1981--1990.
+
+\bibitem[\protect\citeauthoryear{Sanders}{Sanders}{2017}]{sanders2017peupler}
+Sanders L. (2017).
+\newblock {\em Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
+  m{\'e}tropoles.}
+\newblock Presses Universitaires Francois Rabelais.
+
+\bibitem[\protect\citeauthoryear{Son, Bizhani, Christensen, Grassberger, and
+  Paczuski}{Son \textit{et~al.}}{2012}]{son2012percolation}
+Son S.-W., Bizhani G., Christensen C., Grassberger P., Paczuski M. (2012).
+\newblock Percolation theory on interdependent networks based on epidemic
+  spreading.
+\newblock {\em EPL (Europhysics Letters)\/}~{\em 97\/}(1), 16006.
+
+\bibitem[\protect\citeauthoryear{Sorensen and Okata}{Sorensen and
+  Okata}{2010}]{sorensen2010megacities}
+Sorensen A., Okata J. (2010).
+\newblock {\em Megacities: Urban form, governance, and sustainability},
+  Volume~10.
+\newblock Springer Science \& Business Media.
+
+\bibitem[\protect\citeauthoryear{Stauffer and Aharony}{Stauffer and
+  Aharony}{2014}]{stauffer2014introduction}
+Stauffer D., Aharony A. (2014).
+\newblock {\em Introduction to percolation theory}.
+\newblock Taylor \& Francis.
+
+\bibitem[\protect\citeauthoryear{Su, Liu, Xu, Li, Pi, and Weng}{Su
+  \textit{et~al.}}{2017}]{su2017china}
+Su S., Liu Z., Xu Y., Li J., Pi J., Weng M. (2017).
+\newblock China’s megaregion policy: Performance evaluation framework,
+  empirical findings and implications for spatial polycentric governance.
+\newblock {\em Land Use Policy\/}~{\em 63}, 1--19.
+
+\bibitem[\protect\citeauthoryear{Tammaru, Marci{\'n}czak, Van~Ham, and
+  Musterd}{Tammaru \textit{et~al.}}{2015}]{tammaru2015multi}
+Tammaru T., Marci{\'n}czak S., Van~Ham M., Musterd S. (2015).
+\newblock A multi-factor approach to understanding socio-economic segregation
+  in european capital cities.
+\newblock In {\em Socio-economic segregation in European capital cities}, pp.\
+  25--53. Routledge.
+
+\bibitem[\protect\citeauthoryear{Tobler}{Tobler}{2004}]{tobler2004first}
+Tobler W. (2004).
+\newblock On the first law of geography: A reply.
+\newblock {\em Annals of the Association of American Geographers\/}~{\em
+  94\/}(2), 304--310.
+
+\bibitem[\protect\citeauthoryear{Vigui{\'e} and Hallegatte}{Vigui{\'e} and
+  Hallegatte}{2012}]{viguie2012trade}
+Vigui{\'e} V., Hallegatte S. (2012).
+\newblock Trade-offs and synergies in urban climate policies.
+\newblock {\em Nature Climate Change\/}~{\em 2\/}(5), 334.
+
+\bibitem[\protect\citeauthoryear{Zeng, Li, Guo, Gao, Gao, Stanley, and
+  Havlin}{Zeng \textit{et~al.}}{2019}]{Zeng23}
+Zeng G., Li D., Guo S., Gao L., Gao Z., Stanley H.~E., Havlin S. (2019).
+\newblock Switch between critical percolation modes in city traffic dynamics.
+\newblock {\em Proceedings of the National Academy of Sciences\/}~{\em
+  116\/}(1), 23--28.
+
+\end{thebibliography}
 
 \end{document}
-
-%%%%%%%%
-%% -- TEMPLATES
-
-%\begin{table}
-%  \newcolumntype{+}{>{\global\let\currentrowstyle\relax}}
-%  \newcolumntype{^}{>{\currentrowstyle}}
-%  \newcommand{\rowstyle}[1]{\gdef\currentrowstyle{#1}%
-%    #1\ignorespaces
-%  }
-%  \centering
-%  \begin{tabular}{+>{\bfseries}l^c^c^c^c}
-%    \hline
-%    \rowstyle{\bfseries}
-%    & Sepal.Length & Sepal.Width & Petal.Length & Petal.Width\\
-%    Setosa & 5.006 & 3.428 & 1.462 & 0.246\\
-%    Versicolor & 5.936 & 2.77  & 4.26  & 1.326\\
-%    Verginica & 6.588 & 2.974 & 5.552 & 2.026\\
-%    \hline
-%  \end{tabular}
-%  \caption{Morbi malesuada diam at magna condimentum.}
-%  \label{tab:example}
-%\end{table}
-
-%
-%\begin{figure*}[ht] 
-%\resizebox{10cm}{7cm}
-%  {\includegraphics[width=11cm]{spider.jpg}}
-%  \centering
-%  \label{frog}
-%  \caption{A spider, Picture from Didier Josselin.}
-%  \end{figure*}
-%

### 2020-02-13 Docs/Papers/Percolation/paper/paper_v1.tex

--- 2019-03-17 Docs/Papers/Percolation/arxiv/v1/paper.tex
@@
 The variable parametrization allows to consider patterns of optimization for two stylized contradictory sustainability indicators (economic performance and greenhouse gases emissions).
-This suggests a customizable spatial design of policies to develop sustainable territories.\\\medskip
-\textbf{Keywords: } Road network; Multi-dimensional percolation; European urban system; Mega-city region
-}
+This suggests a customizable spatial design of policies to develop sustainable territories.}
+
+\keywords{Road network; Multi-dimensional percolation; European urban system; Mega-city region}
+
+\added{\cite{aleta2018multilayer} multiplex/multilayer}
 
 \section{Introduction}
+
+\strut
+\vspace{-4ex}
+
+\added{traffic : other scale \cite{Li669} \cite{Zeng23} }
 

### 2020-02-13 Docs/Papers/Percolation/submission/Raimbault2019/Raimbault2019.tex

--- 2019-09-11 Docs/Papers/Percolation/paper/paper.tex
@@
 % DOI 
-\doi{10.18713/JIMIS-ddmmyy-v-a}{}
+\doi{10.18713/JIMIS-180719-5-5}{}
 % reviewing process: first date = submission, second date = acceptance 
-\review{Month-in-letters Day Year}{Month-in-letters Day Year}
+\review{March 17, 2019}{October 23, 2019}
 % Volume and Year of the issue
-\publication{N}{YYYY}
+\publication{5}{2019}
 
@@
 % Issue Title 
-\issue{Title of the interdisciplinary issue}
+\issue{Graph and network analysis}
 % List of k Editors
-\editors{}
+\editors{Vincent Labatut}
+
+\pgfplotsset{compat=1.14}
 
@@
 
-\abstract{Network percolation has recently been proposed as a method to characterize the hierarchical structure of an urban system form the bottom-up.
+\abstract{Network percolation has recently been proposed as a method to characterize the hierarchical structure of an urban system from the bottom-up.
 This paper proposes to extend urban network percolation in a multi-dimensional way, to take into account both urban form (spatial distribution of population) and urban functions (here as properties of transportation networks).
@@
 The link between urban form and function remains in particular an open question \citep{batty1994fractal}.
-More generally, the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}.
+More generally, the inclusion of multiple dimensions in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}.
 This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic.
 The method allows combining different dimensions of the urban system, the same way that \cite{cottineau2018defining} link population density with commuting flows to produce multiple definitions of urban areas.
+We can indeed expect significantly different qualitative behaviors when switching from a single-dimension percolation to a multi-dimensional one, since the spatial structure of different urban dimensions are correlated but also largely complementary.
+If this is the case, the method then captures more aspects of the urban system and some of their interrelations.
 
@@
 Case studies of targeted megacities have for example focused on the links between urban form and mobility or resources management \citep{sorensen2010megacities}.
-A significant literature has focused on the sustainability of mega urban regions with a qualitative approach, such as \cite{laquian2011planning} which establish a typology of these and corresponding governance structures in the Asian context.
+A significant literature has focused on the sustainability of mega urban regions with a qualitative approach, such as \cite{laquian2011planning} which establishes a typology of these and corresponding governance structures in the Asian context.
 In the case of Europe, \cite{marull2013emerging} use econometric analysis to see the economic and ecological advantage of integrated urban regions.
@@
 
-The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied and how the network is constructed, and the indicators used to characterized the sustainability of clusters produced.
+The rest of this paper is organized as follows: we first describe the multi-dimensional percolation heuristic, the data and variables to which it is applied and how the network is constructed, and the indicators used to characterize the sustainability of clusters produced.
 We then describe the results of applying this method to population and road network indicators for the whole European Union, focusing on the endogenous regions produced and their sustainability properties.
@@
 %%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
+\begin{figure*}[!ht] 
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/principle.pdf}}
+  {\includegraphics[width=0.75\linewidth]{Fig1.png}}
   \centering
@@
 We do not call our method ``multi-layer percolation'', since nodes are common between layers.
-We show in Fig.~\ref{fig:method} for a schematic representation of the method.
+We show in Figure~\ref{fig:method} for a schematic representation of the method.
 It can be implemented with a propagation heuristic or directly working on adjacency matrices.
-The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a large enough proximity.
+The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a geographical proximity.
 The second component is a strong enough intensity of the urban dimension captured by each layer, simultaneously for all layers considered.
@@
 	\item number of vertices $N_V$;
-	\item cyclomatic number $\mu$ which is defined by $\mu = N_E - N_V + c$ where $c$ is the number of connected components of the graph; this measure captures the number of distinct cycles in the graph
-	\item euclidian efficiency $v_0$, defined by \cite{banos2012towards}, as the average rate between network distance and euclidian distance for all couples of links
+	\item cyclomatic number $\mu$ which is defined by $\mu = N_E - N_V + c$ where $c$ is the number of connected components of the graph; this measure captures the number of distinct cycles in the graph;
+	\item euclidian efficiency $v_0$, defined by \cite{banos2012towards}, as the average rate between network distance and euclidian distance for all couples of links.
 \end{enumerate}
@@
 Indeed, euclidian efficiency measures how the network is performant to link nodes, while the number of cycles is linked to redundancy of paths and in a way to robustness.
-These choice are arbitrary, but several aspects of transportation networks are still captured by these indicators.
+These choices are arbitrary, but several aspects of transportation networks are still captured by these indicators.
 An increase of each is related to a more performant network regarding different dimensions, what is relevant for our application.
@@
 
-Using these potential flows follows the logic of \cite{arbabi2019development} which shows a need for improved intra-city-region mobility in England and Wales.
+Using these potential flows follows the logic of \cite{arbabi2019development} which show a need for improved intra-city-region mobility in England and Wales.
 Considering the regions as entities in which such transportation development policies can more easily been developed, we look at the sustainability of different possible regions if these potential flows were realized.
-Varying the parameters $\gamma$ and $d_0$ allows to control for the economic activity considered (high $\gamma$ values correspond to high added-value activities) and the span of interactions through $d_0$.
+Varying the parameters $\gamma$ and $d_0$ allows controlling for the economic activity considered (high $\gamma$ values correspond to high added-value activities) and the span of interactions through $d_0$.
 
@@
 %%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
+\begin{figure*}[!ht] 
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.49\linewidth]{figures/abssize_nodes.png}}
-  {\includegraphics[width=0.49\linewidth]{figures/relsize_nodes.png}}
+  {\includegraphics[width=\linewidth]{Fig2.png}}
   \centering
@@
 %%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[ht] 
+\begin{figure*}[!ht] 
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/fractaldimension.png}}
+  {\includegraphics[width=\linewidth]{Fig3.png}}
   \centering
@@
 We study here this property, for different threshold parameter values.
-We make the radius vary betweem 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.
+We make the radius vary between 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.
+These values yield a good precision for the radius which is the most important variable to study transition and estimate fractal dimensions, while this population threshold is enough to provide a broad coverage for large radiuses (as shown by other explorations described below).
+Changing network variables and their threshold allows us to investigate how the transition behavior does change regarding the dimension considered and its intensity.
 
-The absolute and relative sizes of the largest cluster are plotted in Fig.~\ref{fig:percolation} as a function of the percolation radius.
+The absolute and relative sizes of the largest cluster are plotted in Figure~\ref{fig:percolation} as a function of the percolation radius.
 This aspect first gives methodological information on multilayer percolation.
@@
 
-% ! largest cluster size in terms of edges makes no sense with the implementation we took - nodes only
-%      nwcol         popthq         nwthq          radius         nwparam             fractdim     
-% ecount  :186   Min.   :0.85   Min.   :0.00   Min.   :  8000   Length:1023        Min.   :0.9239  
-% euclPerf:279   1st Qu.:0.85   1st Qu.:0.00   1st Qu.: 31000   Class :character   1st Qu.:1.2639  
-% mu      :279   Median :0.85   Median :0.80   Median : 54000   Mode  :character   Median :1.3306  
-% vcount  :279   Mean   :0.85   Mean   :0.55   Mean   : 54000                      Mean   :1.3225  
- %               3rd Qu.:0.85   3rd Qu.:0.95   3rd Qu.: 77000                      3rd Qu.:1.3909  
- %               Max.   :0.85   Max.   :0.95   Max.   :100000                      Max.   :1.6058                                                                                    NA's   :337     
-%   fractdimsd     fractdimadjrsquared   fractrelsd    
-% Min.   :0.0544   Min.   :0.5111      Min.   :0.0395  
-% 1st Qu.:0.0966   1st Qu.:0.7016      1st Qu.:0.0700  
-% Median :0.1328   Median :0.7467      Median :0.1043  
-% Mean   :0.1317   Mean   :0.7456      Mean   :0.1009  
-% 3rd Qu.:0.1578   3rd Qu.:0.7981      3rd Qu.:0.1232  
-% Max.   :0.2381   Max.   :0.8642      Max.   :0.1958  
-% NA's   :337      NA's   :337         NA's   :337
-
 We study also the evolution of the fractal dimension of clusters as a function of $r_0$, to verify how the initial percolation approach is robust to multi-dimensionality.
@@
 where $N_c$ is the size of cluster $c$ and $\delta_c$ its diameter.
-These estimations are shown in Fig.~\ref{fig:fractaldim}.
+These estimations are shown in Figure~\ref{fig:fractaldim}.
 As a negative result, which could be due to the abstract nature of our network, a clear maximum in the value of the fractal dimension can not be found.
@@
 
-%  fractdimsignif
-%# A tibble: 11 x 5
-%# Groups:   nwcol, popthq, nwthq [?]
-%   nwcol    popthq nwthq nwparam       signif
-%   <fct>     <dbl> <dbl> <chr>          <dbl>
-% 1 ecount     0.85  0    ecount0      -0.105 
-% 2 ecount     0.85  0.8  ecount0.8    -0.0452
-% 3 euclPerf   0.85  0    euclPerf0    -0.108 
-% 4 euclPerf   0.85  0.8  euclPerf0.8   0.0360
-% 5 euclPerf   0.85  0.95 euclPerf0.95 -0.108 
-% 6 mu         0.85  0    mu0          -0.108 
-% 7 mu         0.85  0.8  mu0.8        -0.108 
-% 8 mu         0.85  0.95 mu0.95        0.375 
-% 9 vcount     0.85  0    vcount0      -0.105 
-%10 vcount     0.85  0.8  vcount0.8    -0.0416
-%11 vcount     0.85  0.95 vcount0.95   -0.0868
-
 %%%%%%%%%%%%%%%%%%%%
-\begin{figure}[h!] 
+\begin{figure}[!ht] 
 %\resizebox{10cm}{7cm}
-{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}\\
-  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
-   {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
+{\includegraphics[width=\linewidth]{Fig4.png}}
   \centering
-
   \caption{Examples of obtained clusters for different parameter values.
@@
 %%%%%%%%%%%%%%%%%%%%
-\begin{figure}[h!] 
+\begin{figure}[!ht] 
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=0.7\linewidth]{figures/full_effective_pareto.png}}
+  {\includegraphics[width=0.7\linewidth]{Fig5.png}}
   \centering
@@
 These are here defined endogenously from the bottom-up and have a priori no reason to coincide with these functional regions.
-We show some examples in Fig.~\ref{fig:exclusters}.
-The first map of this figure, obtained for high population and network thresholds ($\theta_P = 0.95$ and $\theta_N = 0.9$), but a low radius $r_0 = 8$km and edge count $N_E$ to define the road network layer, include several mega-city regions described by \citep{hall2006polycentric}, namely London metropolitan area, the Randstad in Netherland, the Rhein-Main and Rhein-Ruhr in Germany, Greater Paris in France, Brussels area in Belgium.
-The same parameters with $\theta_N = 0$ yield not exactly the same regions, as confirmed by the transition curves in Fig.~\ref{fig:percolation}, what means that our approach taking into account two dimensions may capture effective processes of mega-city regions, in particular by including the road network which is crucial as these regions are integrated in terms of flows.
+We show some examples in Figure~\ref{fig:exclusters}.
+The first map of this figure, obtained for high population and network thresholds ($\theta_P = 0.95$ and $\theta_N = 0.9$), but a low radius $r_0 = 8$km and edge count $N_E$ to define the road network layer, include several mega-city regions described by \cite{hall2006polycentric}, namely London metropolitan area, the Randstad in Netherland, the Rhein-Main and Rhein-Ruhr in Germany, Greater Paris in France, Brussels area in Belgium.
+The same parameters with $\theta_N = 0$ yield not exactly the same regions, as confirmed by the transition curves in Figure~\ref{fig:percolation}, what means that our approach taking into account two dimensions may capture effective processes of mega-city regions, in particular by including the road network which is crucial as these regions are integrated in terms of flows.
 The bottom-left map show an example of large clusters emerging in UK and in the center of Europe, the South remaining largely disconnected.
@@
 Before considering the flow-based indicators described above, we can already study basic summary measures such as population $P_c$ and effective emissions $EM_c$.
-We show in Fig.~\ref{fig:paretos} point clouds of $\log EM_c$ against $1 - \log P_c$ for some configurations.
+We show in Figure~\ref{fig:paretos} point clouds of $\log EM_c$ against $1 - \log P_c$ for some configurations.
 Indeed, regarding the population it contains, an area can be more or less efficient in terms of emissions.
@@
 In practice, this yields optimization compromises in the context of multi-objective optimization, when no aggregation of the dimensions is possible or desirable.
-In Fig.~\ref{fig:paretos}, the front is the lower bound of the point cloud.
+In Figure~\ref{fig:paretos}, the front is the lower bound of the point cloud.
 We also find no dominating point for each configuration, i.e.
@@
 %\resizebox{10cm}{7cm}
-  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
+  {\includegraphics[width=\linewidth]{Fig7.png}}
   \centering  
-  \caption{Relative potential emissions $\sum_c g_c$ against relative potential economic unefficiency $\sum_c e_c$ (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows).
+  \caption{Relative potential emissions $\sum_c g_c$ against relative potential economic inefficiency $\sum_c e_c$ (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows).
 Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
@@
 
-We show in Fig.~\ref{fig:emissions-pc1} the value of $\sum_c \tilde{g}_c$ as a function of the first morphological principal component, for extreme values of $\gamma$ and $d_0$.
-There seems to exist an optimal intermediate value for PC1 regarding the minimization of normalized indicator for emissions only.
+We show in Figure~\ref{fig:emissions-pc1} the value of $\sum_c \tilde{g}_c$ as a function of the first morphological principal component, for extreme values of $\gamma$ and $d_0$.
+There seems to exist an optimal intermediate value for PC1 regarding the minimization of the normalized indicator for emissions only.
 This would correspond to an intermediate level of monocentricity, meaning that urban areas which are too polycentric and spread would emit more, but also areas that are too much monocentric.
 This behavior does not occur for long-range $d_0 = 100$km and low-hierarchy $\gamma=0.5$ interactions.
-The mostly monocentric but emitting configurations capture most of population (given by the level of color), whereas the intermediate configuration capture around half of the population, what means that these low-emissions potential urban regions can cover a significant part of European population.
+The mostly monocentric but emitting configurations capture most of population (given by the level of color), whereas the intermediate configurations capture around half of the population, what means that these low-emissions potential urban regions can cover a significant part of European population.
 
 However, when considering both emissions and economic indicators, urban form then acts as a compromise variable.
-We show in Fig.~\ref{fig:paretos-relative} the point clouds of $\sum_c g_c$ against $\sum_c e_c$, which produce clear Pareto fronts, which shape varies with $\gamma$ and $d_0$.
+We show in Figure~\ref{fig:paretos-relative} the point clouds of $\sum_c g_c$ against $\sum_c e_c$, which produce clear Pareto fronts, which shape varies with $\gamma$ and $d_0$.
 As the color level gives the value of PC1, we can see the points on the different fronts with very different morphological properties.
 In some cases, highly monocentric areas (yellow points) can be a good compromise, whereas the intermediate optimal for emissions shown before may yield highly inefficient areas (dominated green points).
-For example, considering the fronts for $\gamma = 2$ which have both very similar shape, the points with the lowest emissions are on the top-left of the front and correspond to the optimal unveiled in Fig.~\ref{fig:emissions-pc1}.
+For example, considering the fronts for $\gamma = 2$ which have both very similar shape, the points with the lowest emissions are on the top-left of the front and correspond to the optimal unveiled in Figure~\ref{fig:emissions-pc1}.
 These have however a very low economic efficiency (high inefficiency) and small improvements can be done with the points below, before switching to a totally different urban form with a high value of PC1 (yellow points, highly monocentric).
@@
 The OpenMOLE model exploration platform provides a transparent access to genetic algorithms for multi-objective optimization \citep{reuillon2013openmole}.
-The use of such calibration algorithms would allow to unveil the effective form of Pareto fronts, that we may have missed here through the grid sampling.
+The use of such calibration algorithms would allow unveiling the effective form of Pareto fronts, that we may have missed here through the grid sampling.
 
@@
 Indeed, privileging some infrastructure developments but also collaborations between urban centers can be seen as an aspect of a small scale planning, or territorial strategy.
-As we integrated potential flows that would result from such development, and consider their economic and emissions consequences, and did it in an endogenous way, we suggest that evidence-based strategies for territorial development at the European level could be inspired by this work.
+As we integrated potential flows that would result from such development, and considered their economic and emissions consequences, and did it in an endogenous way, we suggest that evidence-based strategies for territorial development at the European level could be inspired by this work.
 This would naturally imply a more thorough data integration, model calibration and operationalization.
@@
 Its application to the issue of sustainable mega-city regions show how its properties of constructing urban clusters from the bottom-up can be used to study sustainability issues.
-This work also illustrates the importance of following data-driven paradigms even when developing, as what is understood of the behavior of the heuristic is through its application to real data and issues.
+This work also illustrates the importance of following data-driven paradigms even when developing new methods, as what is understood of the behavior of the heuristic is through its application to real data and issues.
+
+\section*{Acknowledgment}
+
+This work is part of DynamiCity, a FUI project funded by BPI France, Auvergne-Rh{\^o}ne-Alpes region, Ile-de-France region and Lyon metropolis.
 
 \bibliographystyle{jimis-en}
-\bibliography{biblio}
-
-%\appendix\footnotesize
-
-%\section{Appendix 1: supplementary sensitivity analyses}
-
-%\sframe{Results: effective emissions}{
-
-%\textit{Effective emissions exhibit a supralinear scaling of population}
-%\includegraphics[height=0.83\textheight]{figures/aggreg_effective.png}
-
-%\sframe{Results: all clusters Pareto fronts}{
-
-%\textit{Variation of Pareto front patterns when potential parameter $\gamma,d_0$ vary.}
-
-%\includegraphics[width=0.9\textwidth]{figures/full_paretos.png}
-
-%\sframe{Results: an optimal morphology}{
-
-%\textit{More monocentric areas are more optimal in terms of relative emissions and efficiency ?}
-
-%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relefficiency.png}
-%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relemissions.png}
-
-%  curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.
-
-%\section{Acknowledgment}
-%
-
-%\section{Biography}
-%Here, feel free to add short biographies of the authors.
+\bibliography{Raimbault2019}
 
 \end{document}
-
-%%%%%%%%
-%% -- TEMPLATES
-
-%\begin{table}
-%  \newcolumntype{+}{>{\global\let\currentrowstyle\relax}}
-%  \newcolumntype{^}{>{\currentrowstyle}}
-%  \newcommand{\rowstyle}[1]{\gdef\currentrowstyle{#1}%
-%    #1\ignorespaces
-%  }
-%  \centering
-%  \begin{tabular}{+>{\bfseries}l^c^c^c^c}
-%    \hline
-%    \rowstyle{\bfseries}
-%    & Sepal.Length & Sepal.Width & Petal.Length & Petal.Width\\
-%    Setosa & 5.006 & 3.428 & 1.462 & 0.246\\
-%    Versicolor & 5.936 & 2.77  & 4.26  & 1.326\\
-%    Verginica & 6.588 & 2.974 & 5.552 & 2.026\\
-%    \hline
-%  \end{tabular}
-%  \caption{Morbi malesuada diam at magna condimentum.}
-%  \label{tab:example}
-%\end{table}
-
-%
-%\begin{figure*}[ht] 
-%\resizebox{10cm}{7cm}
-%  {\includegraphics[width=11cm]{spider.jpg}}
-%  \centering
-%  \label{frog}
-%  \caption{A spider, Picture from Didier Josselin.}
-%  \end{figure*}
-%
